### 2018-05-04 zwischenstand.tex
Zwischenstand und Procedure
Das was wir am 2.5 besprochen haben und nochmal der "neue" Ansatz sauber
aufgeschrieben

\documentclass[11pt]{article}
\usepackage{amsmath} % Math
\usepackage{amssymb} % Math symbols
\usepackage[english]{babel} % Language
\usepackage{fancyhdr} % Header
\usepackage[a4paper, total={15cm, 20cm}]{geometry} % Dimensions of the paper and the text area
\usepackage[utf8]{inputenc} % encoding in UTF, needed for umlauts if German
\usepackage{mathtools} % Text above arrows
\usepackage{msc} % Drawing MSCs
\usepackage{multicol} % Multiple columns
\usepackage[explicit]{titlesec} % Automatic section titles
\usepackage{tikz} % Diagrams
\usetikzlibrary{arrows.meta, automata, shapes, matrix}

% Other packages that might be useful in the future
\usepackage{mathptmx,amssymb,amsmath}
%\usepackage{lingmacros}
%\usepackage{tree-dvips}
%\usepackage{ulem}
%\usepackage{amsthm}
%\usepackage{amsbsy}
%\usepackage{textcomp,gensymb}
%\usepackage{graphicx}
%\usepackage{mathtools}

% Custom variant of msc environment:
% - No "msc" keyword, longer partial messages
% - Increased vertical distance between messages
% - Less distance to the frame left and right
% - Less distance between header and processes
% - Less distance between footer and frame
% - Passing all given options down to the msc environment
\newenvironment{cmsc}[1][]{\msc[msc keyword={}, self message width=1.1cm, level height=0.6cm, environment distance=1.2cm, head top distance=0.75cm, foot distance=0.5cm, #1]}{\endmsc}

% No indentation at new paragraphs
\setlength{\parindent}{0pt}

% Distance between columns
\setlength{\columnsep}{1cm}
% Vertical line between columns
\setlength{\columnseprule}{0.5pt}
\def\columnseprulecolor{\color{gray}}

% Settings
\newcommand{\sheetNr}{1}

%% Header
\fancyhf{}
\pagestyle{fancy}
\lhead{The Logic Of Knowledge Bases Exercise Sheet \sheetNr}
\rhead{Timo Bergerbusch: 344408}
\setlength{\headheight}{28pt}

%% Automatic section titles
%\titleformat{\section}{\normalfont\Large\bfseries}{}{0em}{Exercise #1}
%\titleformat{\subsection}{\normalfont\large\bfseries}{}{0em}{#1)}

\begin{document}

\section*{Stand 02.05.2018:}
\subsection*{Schlechte Daten}
\begin{itemize}
	\item keine ID's: kann keine zwei Movements von einer Person aggregieren
	\item[$\Rightarrow$] Einteilung in Gruppen nicht wirklich Personen zuzuordnen
	\item[]
	\item Möglich wäre es Movements basierend auf: \texttt{Alter}, \texttt{Geschlecht}, \texttt{Strata} $\Rightarrow$ ID?
	\item Unterscheidung von zwei zeitgleichen Movements, welche demnach nicht der selben Person gehören können
	\item \underline{Problem}: false Positives: wenn zwei unterschiedliche Personen als eine erkannt werden
	\item Möglicherweise als Datenbank-JOIN?
\end{itemize}

\subsection*{Deliverables}
\begin{itemize}
	\item ein Plotter für alle möglichen Variable-Kombinationen und auch so etwas wie: \texttt{DIST/TIME}
	\item Python Datensatz Verringerung
\end{itemize}

\end{document}

### 2018-05-17 testdata/testDataGenerator.py
zwischenstand late push

@@ -7,7 +7,8 @@ DATA_NAMES = ["origin", "destination", "reason", "HMV", "RED" "mean of Transport
 IGNORE_LINE_VALUES = ['ORIGEN DESTINO MOTIVO MODO HMV RED DURACION DISTKM EST EDAD SEXO FEV ',
                       " "]  # a list of values that should not be considered as a data containing line
 
-
+					  
+					  
 class DataEntry:
     """
     A class to store a single data entry. This is further used to mask the desired columns

### 2018-06-24 k-Means Miriam/Clustering.tex
miriams part

\section{Clustering with RapidMiner}

RapidMiner has different Modules for Clustering already implemented. We apply the k-Means algorithm with mixed measure (euclidean).

\includegraphics[width=0.9\textwidth]{ClusteringRapid.PNG}

The Process does the following steps:
\begin{description}
	\item[Retrieve] This Block gives the data in the process. For the comparison at the end we also retrieve the original data.
	\item[Nominal to Numerical] This changes the nominal data to numerical data, so that we can apply PCA in the next step
	\item[PCA] Here we apply the PCA reduction to the data with a variance threshold $0.95$
	\item[Clustering] Here the number of Clusters has to be fixed. We decided, that 10 runs should be enough. We can choose between differen measure types and tried mixed euclidean and squared euclidean distance
	\item[Generate ID] Is just needed to join the datasets 
	\item[Select Attribute] Here we just pick the Strata data for comparing with the clustering
	\item[Join] For comparing the Clustering and the Strata we join the two filtered data sets at the ID
\end{description}

\subsection{Original Data}

Applying this process to the original data you can see without alignment of strata and cluster, that the distribution of cluster and strata is not close to the same.

\ref{fig:OrgDist}. 
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidStrata.PNG}
  \caption{Strata}
  \label{fig:OrgSt}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidCluster.PNG}
  \caption{Cluster}
  \label{fig:OrgCl}
\end{subfigure}
\caption{Distribution of original data}
\label{fig:OrgDist}
\end{figure}

### 2018-06-24 k-Means Miriam/PCA.tex
miriams part

--- 2018-06-24 k-Means Miriam/Clustering.tex
@@
-\section{Clustering with RapidMiner}
+\section{PCA}
 
-RapidMiner has different Modules for Clustering already implemented. We apply the k-Means algorithm with mixed measure (euclidean).
+\textbf{PCA} uses an orthogonal transformation (vectors multiplied are 0) to
+convert a dataset with maybe correlated datas into a dataset withou
+correlations.
 
-\includegraphics[width=0.9\textwidth]{ClusteringRapid.PNG}
+Rapidminer has a tool applying the PCA algorithm on given data. Output are the
+new attributes, which are a combination of the old ones.
+
+\subsection{Rapidminer Process}
+\includegraphics[width=0.9\textwidth]{PCAClustering}
 
@@
 
-\subsection{Original Data}
+This is the basis for all datasets.
 
-Applying this process to the original data you can see without alignment of strata and cluster, that the distribution of cluster and strata is not close to the same.
+\subsection{Applying on original data}
+Applying the PCA module on the original data without EST and a variance
+threshold of 0.95 we get two Attributes, so we can assume, that the data is strongly connected. As expected the computating time decreases a lot.
 
-\ref{fig:OrgDist}. 
+For a clustering of 6 clusters we get a distribution that is complete different from the original distribution \ref{fig:OrgDist}. 
 \begin{figure}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidStrata.PNG}
+  \includegraphics[width=.4\linewidth]{ClusterOrigRapidStrata.PNG}
   \caption{Strata}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidCluster.PNG}
+  \includegraphics[width=.4\linewidth]{ClusterOrigRapidCluster.PNG}
   \caption{Cluster}

### 2018-06-24 testdata/testDataGenerator.py
Testsetgenerator fix
fixed the issue of ID computation after randomized picking

@@ -1,15 +1,19 @@
 import itertools
 import re
 import random
+import argparse
 
 FILE_NAME = "BaseDatosMDE.txt"  # the dataset to infer the subsets from
 
-# FILE_NAME = "Daten - Uhrzeiten.txt"  # the dataset to infer the subsets from
+# the header names that should be used
 DATA_NAMES = ["origin", "destination", "reason", "MODO", "HMV", "RED", "duration",
               "distance", "strata", "age", "gender", "FEV", "ID"]  # the header names
 
+# the lines that should be neglected in the FILE_NAME-file
 IGNORE_LINE_VALUES = ['ORIGEN DESTINO MOTIVO MODO HMV RED DURACION DISTKM EST EDAD SEXO FEV ',
-                      " "]  # a list of values that should not be considered as a data containing line
+                      " "]
+
+random.seed(7811)
 
 
 class DataEntry:
@@ -76,10 +80,25 @@ class DataEntry:
         return False
 
     def computeID(self):
-        # self.data[12] = "Age:{};Gender:{};Strata:{}".format(self.data[9], self.data[10], self.data[8])
+        """
+        computes an ID for this element. The ID is a random number between 1 and 150000.
+        This is done, because the value should be just an identifier.
+        If the numerical value is used it throws of every clustering and therefore indicate an error
+
+        :return: the ID (int, 1<= value <= 150000)
+        """
         self.data[12] = random.randint(1, 150000)
 
     def compareTo(self, otherEntry):
+        """
+        compares a DataEntry to a given other DataEntry.
+        The comparison is made based on the strata, gender and age.
+        If all those 3 are equal we consider two DataEntries to be from the same person.
+        Further logical reasoning could be made, through pathfinding with time and origin-destination constraints
+
+        :param otherEntry: the other DataEntry one should compare to
+        :return: a boolean value whether this DataEntry is equal to the given DataEntry
+        """
         if self.data[8] == otherEntry.data[8] and self.data[9] == otherEntry.data[9] and self.data[10] == \
                 otherEntry.data[10]:
             return True
@@ -116,11 +135,13 @@ class DataCollection:
         writes the information of all DataEntries and the length of the DataCollection
         :return: -undef-
         """
-        self.computeIDs()
-        print("Data Collection: length={}".format(len(self.data_entries)))
+        # self.computeIDs()
+        if args.debug: print("INFORMATION :: Data Collection: length={}".format(len(self.data_entries)))
         for entry in self.data_entries:
             entry.writeDataEntry()
 
[…]

### 2018-06-24 dokumentation/gower_distance.tex
structure and submission template test

\documentclass[11pt]{article}
\usepackage{amsmath} % Math
\usepackage{amssymb} % Math symbols
\usepackage[english]{babel} % Language
\usepackage{fancyhdr} % Header
\usepackage[a4paper, total={15cm, 20cm}]{geometry} % Dimensions of the paper and the text area
\usepackage[utf8]{inputenc} % encoding in UTF, needed for umlauts if German
\usepackage{mathtools} % Text above arrows
\usepackage{msc} % Drawing MSCs
\usepackage{multicol} % Multiple columns
\usepackage[explicit]{titlesec} % Automatic section titles
\usepackage{tikz} % Diagrams
\usetikzlibrary{arrows.meta, automata, shapes, matrix}

% Other packages that might be useful in the future
\usepackage{mathptmx,amssymb,amsmath}
%\usepackage{lingmacros}
%\usepackage{tree-dvips}
%\usepackage{ulem}
%\usepackage{amsthm}
%\usepackage{amsbsy}
%\usepackage{textcomp,gensymb}
%\usepackage{graphicx}
%\usepackage{mathtools}

% Custom variant of msc environment:
% - No "msc" keyword, longer partial messages
% - Increased vertical distance between messages
% - Less distance to the frame left and right
% - Less distance between header and processes
% - Less distance between footer and frame
% - Passing all given options down to the msc environment
\newenvironment{cmsc}[1][]{\msc[msc keyword={}, self message width=1.1cm, level height=0.6cm, environment distance=1.2cm, head top distance=0.75cm, foot distance=0.5cm, #1]}{\endmsc}

% No indentation at new paragraphs
\setlength{\parindent}{0pt}

% Distance between columns
\setlength{\columnsep}{1cm}
% Vertical line between columns
\setlength{\columnseprule}{0.5pt}
\def\columnseprulecolor{\color{gray}}

% Settings
%\newcommand{\sheetNr}{1}

%% Header
\fancyhf{}
\pagestyle{fancy}
\lhead{Clustering of Mobility Data}
%\rhead{Timo Bergerbusch: 344408}
\setlength{\headheight}{28pt}

%% Automatic section titles
%\titleformat{\section}{\normalfont\Large\bfseries}{}{0em}{Exercise #1}
%\titleformat{\subsection}{\normalfont\large\bfseries}{}{0em}{#1)}

\begin{document}

\section*{Gower Distance}

When clustering data with algorithms such as \textit{k-means} or \textit{k-medoids}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm. These distance measures come with constraints though, since they are only defined for numerical variables. In the scope of this work, we are facing data points with mixed variable types, therefore our choice of an appropriate distance measure fell on the Gower distance measure to calculate an overall \textit{similarity} between mixed data points. The gower similarity distance distinguishes between three types of variables:\\

\textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match [CITE]. In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}. Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values. Hence, we consider all variables, that are not explicitly numerical, as categorical variables.

\textbf{Categorical} variables form a set of unordered values and are comparible to ENUMs in programming languages.

\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.

\subsection*{Calculating the distances}

Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
\begin{equation*}
  S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
\end{equation*} 
where $s_{xy,k}$ denotes a score for the similarity of variables of index $k$ between the data points $x$ and $y$, that's definition is dependent on the type of the variable, as defined below. In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, where the value 1 means, when variables are comparable and 0 when not. This could happen, if some values are not defined. Within this work, the dataset is complete, therefore, the sum of $\delta_{xy,k}$ for all $0 <= k <= v$ equals to $v$, the number of considered variables in each data point. Thus, the similarity coefficient can be interpreted as the average value of all similarity scores.\\

With respect to the variable type, $s_{xy,k}$ is defined by
\begin{itemize}
	\item \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation. As pointed above, 0 values are not considered a match and even further, not considered to be comparable. Hence, the values result as in the table
	\begin{center}\begin{tabular}{l | c c c r}
		i & 1 & 1 & 0 & 0 \\
		j & 1 & 0 & 1 & 0 \\
		\hline
		$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
		$\delta_{xy,k}$ & 1 & 1 & 1 & 0
	\end{tabular}\end{center}

    \item \textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$ and 0, if they differ.

    \item \textbf{Numerical:} for numerical variables, the similarity score is calculated by
    \begin{equation}
      s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
    \end{equation}
    where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept. This can be a global range of acceptable values for variable $k$ or chosen on the basis of the dataset.
\end{itemize}
TODO: sources, citations
\end{document}

### 2018-06-24 dokumentation/submission.tex
structure and submission template test

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
	%
	\title{Clustering Analysis of Mobility Data}
	%
	%\titlerunning{Abbreviated paper title}
	% If the paper title is too long for the running head, you can set
	% an abbreviated paper title here
	%
	\author{Miriam NACHNAME\and
		Martin Breuer\and
		Moritz NACHNAME\and
		Timo Bergerbusch\and
		Walter NACHNAME}
	%
	\authorrunning{F. Author et al.} %TODO
	% First names are abbreviated in the running head.
	% If there are more than two authors, 'et al.' is used.
	%
	\institute{RWTH Aachen, Templergraben 55, 52062 Aachen, Germany}
	%
	\maketitle              % typeset the header of the contribution
	%
	\begin{abstract}
		The abstract should briefly summarize the contents of the paper in
		150--250 words.

		\keywords{Clustering \and Rapidminer \and Cluster \and Data Mining} %TODO: maybe more
	\end{abstract}
	%
	%
	%
	\section{Introduction}
	\section{Preprocessing}

\end{document}

### 2018-06-25 testdata/testDataGenerator.py
fixed reuse of IDs

@@ -13,6 +13,8 @@ DATA_NAMES = ["origin", "destination", "reason", "MODO", "HMV", "RED", "duration
 IGNORE_LINE_VALUES = ['ORIGEN DESTINO MOTIVO MODO HMV RED DURACION DISTKM EST EDAD SEXO FEV ',
                       " "]
 
+USED_IDS = []
+
 random.seed(7811)
 
 
@@ -87,7 +89,15 @@ class DataEntry:
 
         :return: the ID (int, 1<= value <= 150000)
         """
-        self.data[12] = random.randint(1, 150000)
+
+        guard = True
+        while guard:
+            id = random.randint(1, 500000)
+            if id not in USED_IDS:
+                guard = False
+
+        USED_IDS.append(id)
+        self.data[12] = id
 
     def compareTo(self, otherEntry):
         """
@@ -189,6 +199,7 @@ class DataCollection:
         # TODO: what if random Entries?
         # Then two action of the same person, which were consequtive in the original
         # are not neccassarily consequtive in the shrink-generate Set and therefore not be considered from the same person.
+        USED_IDS = []
         if len(self.data_entries) == 0:
             if args.debug: print("WARNING :: Can not compute the IDs because the list is empty")
         for i in range(0, len(self.data_entries)):
@@ -333,6 +344,7 @@ def parse_lines_of_content(linesOfContent):
             data_entries.append(data_entry)
     return data_entries
 
+
 # TODO: Merge both test set generators
 def create_equal_testset(mask, data_collection):
     numbers = [100, 200, 500, 1000, 2000]
@@ -420,7 +432,7 @@ mask = [
     True,  # 5 - average time
     True,  # 6 - duration
     True,  # 7 - distance
-    False,  # 8 - strata
+    True,  # 8 - strata
     True,  # 9 - age
     True,  # 10 - gender
     True,  # 11 - FEV

### 2018-06-25 Miriam-RapidMiner/DecisionTree.tex
miriams aktueller Stand

\section{Decision Tree}

Constructing a decision tree is a good way to see first correlations.

To get a decision tree from a dataset different ways are possible. We are concentrating on the simple information gain and applying again RapidMiner on the dataset.

\subsection*{RapidMiner and decisiontrees}
\includegraphics[width = 0.9\textwidth]{DecisionTreeRapidModel.PNG}

RapidMiner does the following steps:
\begin{description}
	\item[Retrieve] Includes the dataset
	\item[Generate Attribute] Changes the Numerical Attribute strata to a nominal one
	\item[Set Role] Gives strata the label role, so that the decision tree ends with this
	\item[Decision Tree] Creates a decision tree
\end{description}

Furthermore we choose information gain as splitting criterium (minimal gain 0.1) and a confidence of 0.25.

\subsection*{Full dataset}

Applying this process on the full dataset we get the not really helpful tree:

\includegraphics[width=0.9\textwidth]{DecisionTreeRapidTreeFull.PNG}

### 2018-06-25 Miriam-RapidMiner/PCA.tex
miriams aktueller Stand

--- 2018-06-24 k-Means Miriam/PCA.tex
@@
 \textbf{PCA} uses an orthogonal transformation (vectors multiplied are 0) to
-convert a dataset with maybe correlated datas into a dataset withou
+convert a dataset with maybe correlated datas into a dataset without
 correlations.
 
-Rapidminer has a tool applying the PCA algorithm on given data. Output are the
+The first component has to satisfy the following:
+
+\begin{align*}
+w_1 &= \text{argmax}_{||w|| = 1} \{\sum_i (t_1)_i^2\}\\
+	&= \text{argmax}_{||w|| = 1} \{\sum_i (x_iw)^2\}\\
+\end{align*}
+Which is achieved by the corresponding eigenvector.
+
+The next components need to fullfill:
+\begin{align*}
+\hat{X}_k &= X - \sum_{s=1}^{k-1} Xw_sw_s^T\\
+w_k &= \text{argmax}_{||w|| = 1} \{||\hat{X}_kw||^2\}\\
+	&= \text{argmax}_{||w|| = 1} \{\frac{w^T\hat{X}_k^T\hat{X}_kw}{w^Tw}\}\\
+\end{align*}
+
+Based on the "weight matrix" a mapping is possible from the original space to a new space in which the dataset is uncorrelated. Also the dimension can be reduced on this manner.
+
+Rapidminer has a tool applying the PCA algorithm on a given dataset. The Output are 
 new attributes, which are a combination of the old ones.
@@
 We can choose between differen measure types and tried mixed euclidean and squared euclidean distance
-	\item[Generate ID] Is just needed to join the datasets 
 	\item[Select Attribute] Here we just pick the Strata data for comparing with the clustering
@@
 
-This is the basis for all datasets.
+This is the basis for all datasets, where we set the role of id to id by adding the dataset to RapidMiner.
 
-\subsection{Applying on original data}
+\subsection{Searching for 6 Clusters}
+
+The first idea is to check, if on this way a clustering can be found, comparable to the strata groups. So we use the process for different datasets.
+
+\subsubsection{Applying on original data}
 Applying the PCA module on the original data without EST and a variance
@@
 
-For a clustering of 6 clusters we get a distribution that is complete different from the original distribution \ref{fig:OrgDist}. 
-\begin{figure}
+For a clustering of 6 clusters we get a distributions that are complete different.\ref{fig:OrgDist}
+\begin{figure}[h]
 \centering
@@
 \end{figure}
+
+Also the distribution in the strata groups of the clustering shows, that there is no real connection of strata and clustering.
+\includegraphics[width=0.9\textwidth]{ClusterPCAOrigRapidDistribution.PNG}

### 2018-06-26 testdata/testDataGenerator.py
small adjustments
just touch it. Dont break it

@@ -61,6 +61,7 @@ class DataEntry:
         :return: a list of the data masked by the values
         """
         assert len(mask) == len(self.data)
+        assert not self.isBroken()
         s = ""
         for i in range(0, len(mask)):
             if mask[i]:
@@ -409,6 +410,8 @@ if args.debug: print('Choosen flags: {}'.format(args))
 # ---------------- END: arguments ----------------
 
 # ---------------- START: preprocessing ----------------
+
+if args.debug: print("INFORMATION ::  Reading data file ({})".format(FILE_NAME))
 # read the whole file named FILE_NAME
 linesOfFile = read_file()
 # clean the file from empty lines and the header
@@ -418,7 +421,12 @@ dataEntries = parse_lines_of_content(linesOfContent)
 # create a new DataCollection
 data_collection = DataCollection(dataEntries)
 
+# TODO: create a set based an precomputed IDs
+# compute the IDs
+if args.debug: print("INFORMATION ::  Compute IDs")
 data_collection.computeIDs()
+
+if args.debug: print("INFORMATION ::  Preprocessing finished")
 
 # ---------------- END: preprocessing ----------------
 

### 2018-06-26 testdata/testDataGenerator.py
#N/A fix

@@ -78,7 +78,7 @@ class DataEntry:
         :return: a boolean value indicating if the entry is broken/invalid
         """
         for i in range(0, len(self.data)):
-            if self.data[i] == "-1":
+            if self.data[i] == "-1" or self.data[i] == "#N/A":
                 return True
         return False
 
@@ -232,7 +232,8 @@ class DataCollection:
     def shrink(self, length):
         # the absolute numbers in the original dataset
         # used for checking the ex. of that many elements
-        absolute_numbers = [6963, 52266, 49404, 8772, 5536, 2038]
+        absolute_numbers = [6963, 52265, 49404, 8772, 5536, 2038]
+        # absolute_numbers = [6963, 52266, 49404, 8772, 5536, 2038]
         # the distribution in the original dataset
         relative_numbers = []
         for i in range(0, 6):
@@ -259,6 +260,7 @@ class DataCollection:
             while count < abs_numbers[i]:
                 if index >= len(self.data_entries):
                     print("ERROR       :: index problem")
+                    print("ERROR       :: Strata {} has overall less elements ({}) than its supposed to have ({})".format(i+1,count,abs_numbers[i]))
                     exit()
                 if self.data_entries[index].data[8] == "{}".format(i + 1):
                     new_data_entries.append(self.data_entries[index])
@@ -340,7 +342,7 @@ def parse_lines_of_content(linesOfContent):
         data_entry = parse_DataEntry(element)
 
         if data_entry.isBroken():
-            print("ERROR :: BROKEN ENTITY! for element {}".format(element))
+            if args.debug: print("WARNING     :: BROKEN ENTITY! for element {}".format(element))
         else:
             data_entries.append(data_entry)
     return data_entries

### 2018-06-26 testdata/testDataGenerator.py
new sets and improved performance

@@ -13,9 +13,15 @@ DATA_NAMES = ["origin", "destination", "reason", "MODO", "HMV", "RED", "duration
 IGNORE_LINE_VALUES = ['ORIGEN DESTINO MOTIVO MODO HMV RED DURACION DISTKM EST EDAD SEXO FEV ',
                       " "]
 
-USED_IDS = []
-
+# a seed to generate pseudo-random numbers
+# (mainly for debug reasons)
 random.seed(7811)
+
+# the last ID used to enumerate all IDs
+LAST_ID = 1
+
+
+# USED_IDS = []
 
 
 class DataEntry:
@@ -91,14 +97,15 @@ class DataEntry:
         :return: the ID (int, 1<= value <= 150000)
         """
 
-        guard = True
-        while guard:
-            id = random.randint(1, 500000)
-            if id not in USED_IDS:
-                guard = False
-
-        USED_IDS.append(id)
-        self.data[12] = id
+        # guard = True
+        # while guard:
+        #     id = random.randint(1, 500000)
+        #     if id not in USED_IDS:
+        #         guard = False
+        # USED_IDS.append(id)
+        global LAST_ID
+        self.data[12] = LAST_ID
+        LAST_ID += 1  # random.randint(1, 9)
 
     def compareTo(self, otherEntry):
         """
@@ -196,19 +203,20 @@ class DataCollection:
                 file.write("\n")
 
     def computeIDs(self):
-        # !!!IMPORTANT!!!
-        # TODO: what if random Entries?
-        # Then two action of the same person, which were consequtive in the original
-        # are not neccassarily consequtive in the shrink-generate Set and therefore not be considered from the same person.
-        USED_IDS = []
+        # USED_IDS = []
         if len(self.data_entries) == 0:
             if args.debug: print("WARNING :: Can not compute the IDs because the list is empty")
         for i in range(0, len(self.data_entries)):
+
+            # compare if two adjacent DataEntrys are equal
             if i > 0 and self.data_entries[i].compareTo(self.data_entries[i - 1]):
+                # if so give them the same ID
                 self.data_entries[i].data[12] = self.data_entries[i - 1].data[12]
             else:
+                # otherwise it gets a new ID
                 self.data_entries[i].computeID()
 
+    @DeprecationWarning
     def findRandomEntry(self, strata):
         """
         searches for a randomly placed DataEntry in the set, which is in the given strata
@@ -230,6 +238,13 @@ class DataCollection:
[…]

### 2018-06-26 testdata/testDataGenerator.py
compute Persons collecting movements instead of only IDs

@@ -18,7 +18,9 @@ IGNORE_LINE_VALUES = ['ORIGEN DESTINO MOTIVO MODO HMV RED DURACION DISTKM EST ED
 random.seed(7811)
 
 # the last ID used to enumerate all IDs
-LAST_ID = 1
+LAST_ID = 0
+
+PERSONS = []
 
 
 # USED_IDS = []
@@ -71,7 +73,10 @@ class DataEntry:
         s = ""
         for i in range(0, len(mask)):
             if mask[i]:
-                s += "{}".format(self.data[i])
+                if i != 12:
+                    s += "{}".format(self.data[i])
+                else:
+                    s += "{}".format(self.data[12].getID())
                 s += " "
         return s
 
@@ -104,8 +109,11 @@ class DataEntry:
         #         guard = False
         # USED_IDS.append(id)
         global LAST_ID
-        self.data[12] = LAST_ID
-        LAST_ID += 1  # random.randint(1, 9)
+        LAST_ID += 1
+        last = Person(LAST_ID, self.data[8], self.data[9], self.data[10])
+        # last.append(self)
+        PERSONS.append(last)
+        self.data[12] = last
 
     def compareTo(self, otherEntry):
         """
@@ -121,6 +129,9 @@ class DataEntry:
                 otherEntry.data[10]:
             return True
         return False
+
+    def getPerson(self):
+        return self.data[12]
 
 
 class DataCollection:
@@ -148,7 +159,7 @@ class DataCollection:
             ret.append(entry.getData(mask))
         return ret
 
-    def writeDataCollection(self):
+    def writeDataCollection(self, mask):
         """
         writes the information of all DataEntries and the length of the DataCollection
         :return: -undef-
@@ -156,7 +167,8 @@ class DataCollection:
         # self.computeIDs()
         if args.debug: print("INFORMATION :: Data Collection: length={}".format(len(self.data_entries)))
         for entry in self.data_entries:
-            entry.writeDataEntry()
+            # entry.writeDataEntry()
+            print("{}".format(entry.getData(mask)))
 
     # TODO currently first appearing
     # SHOULD: random or decidable
@@ -208,13 +220,22 @@ class DataCollection:
             if args.debug: print("WARNING :: Can not compute the IDs because the list is empty")
         for i in range(0, len(self.data_entries)):
 
+            curr = self.data_entries[i]
+            prev = self.data_entries[i - 1]
+
             # compare if two adjacent DataEntrys are equal
[…]

### 2018-06-26 testdata/testDataGenerator.py
eigentliche Funk wieder herstellen

@@ -519,6 +519,9 @@ mask = [
 # data_collection.writeDataCollection(mask)
 
 
+# for person in PERSONS:
+#     print(person.getData())
+# exit()
 
 # create all testsets if all-flag
 # (NOTE: has to be checked against True, since it could be a size number and therefore is an int, but

### 2018-06-26 testdata/testDataGenerator.py
Vector Person Data Generation added

@@ -2,6 +2,7 @@ import itertools
 import re
 import random
 import argparse
+import numpy as np
 
 FILE_NAME = "BaseDatosMDE.txt"  # the dataset to infer the subsets from
 
@@ -335,6 +336,35 @@ class Person:
                                                                                        self.parameters[2],
                                                                                        len(self.movements))
 
+    def computeVector(self):
+        lst = [0] * (2 * 413 + 4 + 7 + 7)
+        sum_of_duration = 0
+        sum_of_way = 0
+        for entry in self.movements:
+            sum_of_duration += int(float(entry.data[6]))
+            sum_of_way += int(float(entry.data[7]))
+
+            lst[int(float(entry.data[0]))] += 1  # add one to origin count
+            lst[413 + int(float(entry.data[1]))] += 1  # add one to destination count
+
+            if entry.data[5] == "AM":
+                lst[2 * 413] += 1
+            elif entry.data[5] == "MD":
+                lst[2 * 413 + 1] += 1
+            elif entry.data[5] == "PM":
+                lst[2 * 413 + 2] += 1
+            elif entry.data[5] == "MN":
+                lst[2 * 413 + 3] += 1
+            lst[2 * 413 + 4 + int(float(entry.data[2])) - 1] += 1  # reason
+            lst[2 * 413 + 4 + 7 + int(float(entry.data[3])) - 1] += 1
+
+        lst.append(sum_of_duration)  # summe der dauer
+        lst.append(sum_of_way)  # summe der strecke
+        lst.append(int(float(self.parameters[2])))  # geschlecht
+        lst.append(int(float(self.parameters[1])))  # alter
+
+        return lst
+
 
 def get_used_headers(mask):
     """
@@ -438,6 +468,17 @@ def create_all_testsets(mask, data_collection):
     if args.debug: print("INFORMATION :: finished creating all Testsets")
 
 
+def create_Person_Vector_File():
+    if args.debug: print("INFORMATION ::  Vector Data of Persons")
+    file_Name = "person_vector_data.txt"
+    with open(file_Name, "w+") as file:
+
+        file.write("\n")
+        for person in PERSONS:
+            file.write(" ".join(str(x) for x in person.computeVector()))
+            file.write("\n")
+
+
 # ---------------- START: arguments ----------------
 parser = argparse.ArgumentParser(
     description='Create Testsets for of the mobility data using equal or normal distribution.')
@@ -518,10 +559,12 @@ mask = [
 
[…]

### 2018-06-27 testdata/testDataGenerator.py
vector flag

@@ -337,6 +337,17 @@ class Person:
                                                                                        len(self.movements))
 
     def computeVector(self):
+        """
+        The vetor has the form (with # as number):
+         #origin1 ... #origin413 #destination1 ... #destination413 AM MD PM MN #reason1 ... #reason7 #MoT1...#MoT7 SD SS G A
+        where MoT= mean of transportation
+        SD = sum duration
+        SS = sum distance
+        G = gender
+        A = age
+        Concluded this has the size 848.
+        :return: the vector
+        """
         lst = [0] * (2 * 413 + 4 + 7 + 7)
         sum_of_duration = 0
         sum_of_way = 0
@@ -356,7 +367,7 @@ class Person:
             elif entry.data[5] == "MN":
                 lst[2 * 413 + 3] += 1
             lst[2 * 413 + 4 + int(float(entry.data[2])) - 1] += 1  # reason
-            lst[2 * 413 + 4 + 7 + int(float(entry.data[3])) - 1] += 1
+            lst[2 * 413 + 4 + 7 + int(float(entry.data[3])) - 1] += 1  # mean of trans
 
         lst.append(sum_of_duration)  # summe der dauer
         lst.append(sum_of_way)  # summe der strecke
@@ -464,12 +475,19 @@ def create_all_testsets(mask, data_collection):
     if args.debug: print("INFORMATION :: save complete Testset of length: 124979")
     data_collection.printToFile("generatedTestSet-full-with-ID.txt", mask)
 
+    create_Person_Vector_File()
+
     # 'Finished'-message
     if args.debug: print("INFORMATION :: finished creating all Testsets")
 
 
 def create_Person_Vector_File():
-    if args.debug: print("INFORMATION ::  Vector Data of Persons")
+    """
+    creates the file 'person_vector_data.txt' with a vector for each person.
+
+    :return: -none-
+    """
+    if args.debug: print("INFORMATION :: Vector Data of Persons")
     file_Name = "person_vector_data.txt"
     with open(file_Name, "w+") as file:
 
@@ -485,6 +503,7 @@ parser = argparse.ArgumentParser(
 # set defaults
 parser.set_defaults(size=False)
 parser.set_defaults(debug=False)
+parser.set_defaults(vector=False)
 parser.set_defaults(distType="equal")
 
 # create  exclusive group
@@ -494,11 +513,14 @@ size_parser.add_argument('-all', dest='size', action='store_true',
                          help="create all test sets considered to be possibly relevant. This creates for the "
[…]

### 2018-06-27 Miriam-RapidMiner/Clustering.tex
miriams aktueller Stand
combined data

--- 2018-06-24 k-Means Miriam/Clustering.tex
@@
 
+\subsubsection{Searching for 6 Cluster}
 Applying this process to the original data you can see without alignment of strata and cluster, that the distribution of cluster and strata is not close to the same.
@@
 \end{figure}
+
+\subsubsection{Searching for 3 Cluster}
+
+The next idea is, that there are maybe just 3 obvious cluster to see and contain two 2 strata.
+
+\subsubsection{Applying on original data}
+
+For a clustering of 3 clusters we get a distributions that are complete different.\ref{fig:OrgDist}
+\begin{figure}[h]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{ClusterOrigRapidStrata2Cluster.PNG}
+  \caption{Strata}
+  \label{fig:OrgSt}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{ClusterOrigRapidCluster2Cluster.PNG}
+  \caption{Cluster}
+  \label{fig:OrgCl}
+\end{subfigure}
+\caption{Distribution of original data}
+\label{fig:OrgDist}
+\end{figure}
+
+Also the distribution in the strata groups of the clustering shows, that there is no real connection of strata and the clustering.
+\includegraphics[width=0.9\textwidth]{ClusterOrigRapidDistribution2Cluster.PNG}
+
+\subsection{Combined Data}
+
+Because of the not really convincing result of the clustering like above we had the idea to combine for every ID the different pathes in one big vector.
+This garants us, that one person just can be in one cluster aswell.
+The idea is to sum up all pathes in one big vector.
+\subsubsection{Code}
+%hier darfst du dich gerne verewigen Timo. Gerne auch vorher und nachher noch mehr schreiben. Titel darf auch geändert werden
+\subsubsection{Searching for 6 Clusters}
+Applying the process with maximal 100 steps on the data gives us the following results.
+
+Like above we check first the total distribution of strata and clusters with a pie. You already see, that it is not the 
+\begin{figure}[h]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{vectorclusteringcluster.PNG}
+  \caption{Strata}
+  \label{fig:OrgSt}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{vectorclusteringstrata.PNG}
+  \caption{Cluster}
+  \label{fig:OrgCl}
+\end{subfigure}
+\caption{Distribution of original data}
+\label{fig:OrgDist}
+\end{figure}
+
+Also the distribution in the strata groups of the clustering shows, that there is not really a connection of strata and the clustering to see.
+\includegraphics[width=0.9\textwidth]{vectorClustering.PNG}
+
+So we change the maximal stepsize to 1000 and let the algorithm run again.
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{vectorclusteringcluster1000.PNG}
+  \caption{Strata}
+  \label{fig:OrgSt}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{vectorclusteringstrata.PNG}
+  \caption{Cluster}
+  \label{fig:OrgCl}
+\end{subfigure}
+\caption{Distribution of original data}
+\label{fig:OrgDist}
+\end{figure}
+
+So we already get the idea, that there are not 6 Clusters, but just 3. Checking the distribution still gives us no strong correlation between strata and cluster.
+
+\includegraphics[width=0.9\textwidth]{vectorClustering1000}
+
+\subsubsection{Searching for 3 Clusters}
+
+Because of the result in the last part, we checked the behavior of the clustering by just searching for 3 Cluster. The first try is with 100 steps again.
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{vectorclusteringcluster3Cluster.PNG}
+  \caption{Strata}
+  \label{fig:OrgSt}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{vectorclusteringstrata.PNG}
+  \caption{Cluster}
+  \label{fig:OrgCl}
+\end{subfigure}
+\caption{Distribution of original data}
+\label{fig:OrgDist}
+\end{figure}
+
+This comes closer by the orginial distribution combining two stratas in one cluster.
+
+But having a look at the inbetween distribution does not show us a real correlation between cluster and strata.
+
+\includegraphics[width=0.9\textwidth]{vectorClustering3Cluster.PNG}
+
+We wanted to check what happens when giving the process 1000 steps maximal.

### 2018-06-27 Miriam-RapidMiner/PCA.tex
combined data

@@
 \subsubsection{Applying on original data}
-Applying the PCA module on the original data without EST and a variance
+Applying the PCA module on the original data without strata and a variance
 threshold of 0.95 we get two Attributes, so we can assume, that the data is strongly connected. As expected the computating time decreases a lot.
@@
 
-Also the distribution in the strata groups of the clustering shows, that there is no real connection of strata and clustering.
+Also the distribution in the strata groups of the clustering shows, that there is no real connection of strata and the clustering.
 \includegraphics[width=0.9\textwidth]{ClusterPCAOrigRapidDistribution.PNG}
+
+\subsection{Searching for 3 Cluster}
+
+The next idea is, that there are maybe just 3 obvious cluster to see and contain two 2 strata.
+
+\subsubsection{Applying on original data}
+Applying the PCA module on the original data without strata and a variance
+threshold of 0.95 we get two Attributes, so we can assume, that the data is strongly connected. As expected the computating time decreases a lot.
+
+For a clustering of 6 clusters we get a distributions that are complete different.\ref{fig:OrgDist}
+\begin{figure}[h]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidStrata2Cluster.PNG}
+  \caption{Strata}
+  \label{fig:OrgSt}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidCluster2Cluster.PNG}
+  \caption{Cluster}
+  \label{fig:OrgCl}
+\end{subfigure}
+\caption{Distribution of original data}
+\label{fig:OrgDist}
+\end{figure}
+
+Also the distribution in the strata groups of the clustering shows, that there is no real connection of strata and the clustering.
+\includegraphics[width=0.9\textwidth]{ClusterPCAOrigRapidDistribution2Cluster.PNG}

### 2018-06-27 testdata/testDataGenerator.py
name refactoring, little doc and full person_vector_data

@@ -2,7 +2,6 @@ import itertools
 import re
 import random
 import argparse
-import numpy as np
 
 FILE_NAME = "BaseDatosMDE.txt"  # the dataset to infer the subsets from
 
@@ -21,10 +20,8 @@ random.seed(7811)
 # the last ID used to enumerate all IDs
 LAST_ID = 0
 
+# the deduced different persons
 PERSONS = []
-
-
-# USED_IDS = []
 
 
 class DataEntry:
@@ -54,34 +51,34 @@ class DataEntry:
         """
         self.data = data
 
-    def writeDataEntry(self):
+    def write_data_entry(self):
         """
         simply writes the data to this DataEntry
         :return: -undef-
         """
         print("data = {}".format(self.data))
 
-    def getData(self, mask):
+    def get_data(self, entry_mask):
         """
         masks the data using a list of boolean values to indicate whether or not a column should be used.
 
         A True value indicates, that the column should be used
-        :param mask: a list of boolean values
+        :param entry_mask: a list of boolean values
         :return: a list of the data masked by the values
         """
-        assert len(mask) == len(self.data)
-        assert not self.isBroken()
+        assert len(entry_mask) == len(self.data)
+        assert not self.is_broken()
         s = ""
-        for i in range(0, len(mask)):
-            if mask[i]:
+        for i in range(0, len(entry_mask)):
+            if entry_mask[i]:
                 if i != 12:
                     s += "{}".format(self.data[i])
                 else:
-                    s += "{}".format(self.data[12].getID())
+                    s += "{}".format(self.data[12].get_id())
                 s += " "
         return s
 
-    def isBroken(self):
+    def is_broken(self):
         """
         tests whether or not a given DataEntry is valid.
         A DataEntry is considered broken if any of it's values is 0.
@@ -94,7 +91,7 @@ class DataEntry:
                 return True
         return False
 
-    def computeID(self):
+    def create_person(self):
         """
         computes an ID for this element. The ID is a random number between 1 and 150000.
         This is done, because the value should be just an identifier.
@@ -116,22 +113,22 @@ class DataEntry:
         PERSONS.append(last)
         self.data[12] = last
 
-    def compareTo(self, otherEntry):
+    def compare_to(self, other_entry):
         """
         compares a DataEntry to a given other DataEntry.
         The comparison is made based on the strata, gender and age.
         If all those 3 are equal we consider two DataEntries to be from the same person.
         Further logical reasoning could be made, through pathfinding with time and origin-destination constraints
 
[…]

### 2018-06-27 Miriam-RapidMiner/Clustering.tex
Added vector code in Tex

@@
 %hier darfst du dich gerne verewigen Timo. Gerne auch vorher und nachher noch mehr schreiben. Titel darf auch geändert werden
+% boa das ist voll liep un soo weischt :D
+
+Instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
+This class stores the ID, the parameters defining a person %TODO: ref zu code basics
+, and all movements from that person.\\
+Then we are able to compute the following vector, with 848 entries, for further usage:
+$$\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,
+ \underbrace{\mathit{AM}, \mathit{MD}, \mathit{PM}, \mathit{MN}}_{4}, 
+ \underbrace{\#r_1, \dots, \#r_7}_{7}, 
+ \underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, 
+ \underbrace{\mathit{SD}, \mathit{SS}, \mathit{G}, \mathit{A}}_{4}$$
+with the following abbreviations ($1 \le i \le 413$, $0 \le j \le 7$):
+\begin{multicols}{2}
+\begin{itemize}
+	\item[$o_i$:]  the $i$-th origin data point
+	\item[$d_i$:]  the $i$-th destination data point
+	\item[$\mathit{AM}$:] movements at time stamp AM
+	\item[$\mathit{MD}$:] movements at time stamp MD
+	\item[$\mathit{PM}$:] movements at time stamp PM
+	\item[$\mathit{MN}$:] movements at time stamp MN
+	\item[$r_j$:] the $j$-th reason
+	\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
+	\item[$\mathit{SD}$:] sum of all durations
+	\item[$\mathit{SS}$:] sum of all distances
+	\item[$\mathit{G}$:] the gender
+	\item[$\mathit{A}$:] the age
+\end{itemize}
+\end{multicols}
+
 \subsubsection{Searching for 6 Clusters}

### 2018-06-27 Miriam-RapidMiner/Clustering.tex
small addition

@@
 , and all movements from that person.\\
-Then we are able to compute the following vector, with 848 entries, for further usage:
+Then we are able to compute the following vector, with 848 entries, for further usage, that combines all movements of the person:
 $$\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,

### 2018-06-29 testdata/testDataGenerator.py
header line for vector data

@@ -522,7 +522,20 @@ def create_person_vector_file():
         print("INFORMATION :: Vector Data of Persons")
     file_name = "person_vector_data.txt"
     with open(file_name, "w+") as file:
-
+    
+        # generate header 
+        header_string = ""
+        for i in range(1, 413):
+            header_string += "o{} ".format(i)
+        for i in range(1, 413):
+            header_string += "d{} ".format(i)
+        header_string += "AM MD PM MN "
+        header_string += "r1 r2 r3 r4 r5 r6 r7"
+        header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7"
+        header_string += "SD SS G A"
+        
+        file.write(header_string)
+        
         file.write("\n")
         if args.debug:
             print_progressbar(0, len(PERSONS) - 1, prefix='INFORMATION ::', suffix='Complete', length=50)

### 2018-06-29 testdata/testDataGenerator.py
space typo

@@ -530,8 +530,8 @@ def create_person_vector_file():
         for i in range(1, 413):
             header_string += "d{} ".format(i)
         header_string += "AM MD PM MN "
-        header_string += "r1 r2 r3 r4 r5 r6 r7"
-        header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7"
+        header_string += "r1 r2 r3 r4 r5 r6 r7 "
+        header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7 "
         header_string += "SD SS G A"
         
         file.write(header_string)

### 2018-06-29 testdata/testDataGenerator.py
tab to space

@@ -438,7 +438,7 @@ def get_used_headers(header_mask):
     """
     s = ""
     for i in itertools.compress(DATA_NAMES, header_mask):
-        s += i + "\t"
+        s += i + " "
     return s
 
 

### 2018-06-29 testdata/testDataGenerator.py
equal distributed person vector

@@ -367,7 +367,7 @@ class Person:
         """
         self.movements.append(e)
 
-    def print(self):
+    def print_line(self):
         """
         prints a String of information about the person using 'get_data'
         :return:
@@ -377,6 +377,9 @@ class Person:
 
     def get_id(self):
         return self.pid
+
+    def get_strata(self):
+        return int(float(self.parameters[0]))
 
     def get_data(self):
         """
@@ -426,6 +429,11 @@ class Person:
         lst.append(int(float(self.parameters[2])))  # geschlecht
         lst.append(int(float(self.parameters[1])))  # alter
 
+        return lst
+
+    def compute_vector_with_strata(self):
+        lst = self.compute_vector()
+        lst.append("s_{}".format(self.parameters[0]))
         return lst
 
 
@@ -512,6 +520,21 @@ def parse_lines_of_content():
     return data_entries
 
 
+def compute_header_string(has_strata=False):
+    header_string = ""
+    for i in range(1, 413):
+        header_string += "o{} ".format(i)
+    for i in range(1, 413):
+        header_string += "d{} ".format(i)
+    header_string += "AM MD PM MN "
+    header_string += "r1 r2 r3 r4 r5 r6 r7 "
+    header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7 "
+    header_string += "SDest SDist G A"
+    if has_strata:
+        header_string += "strata"
+    return header_string
+
+
 def create_person_vector_file():
     """
     creates the file 'person_vector_data.txt' with a vector for each person.
@@ -522,20 +545,12 @@ def create_person_vector_file():
         print("INFORMATION :: Vector Data of Persons")
     file_name = "person_vector_data.txt"
     with open(file_name, "w+") as file:
-    
+
         # generate header 
-        header_string = ""
-        for i in range(1, 413):
-            header_string += "o{} ".format(i)
-        for i in range(1, 413):
-            header_string += "d{} ".format(i)
-        header_string += "AM MD PM MN "
-        header_string += "r1 r2 r3 r4 r5 r6 r7 "
-        header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7 "
-        header_string += "SD SS G A"
-        
+        header_string = compute_header_string()
+
         file.write(header_string)
-        
+
         file.write("\n")
         if args.debug:
[…]

### 2018-06-29 testdata/testDataGenerator.py
space fixed

@@ -433,7 +433,7 @@ class Person:
 
     def compute_vector_with_strata(self):
         lst = self.compute_vector()
-        lst.append("s_{}".format(self.parameters[0]))
+        lst.append("s_{}".format(self.get_strata()))
         return lst
 
 
@@ -520,7 +520,7 @@ def parse_lines_of_content():
     return data_entries
 
 
-def compute_header_string(has_strata=False):
+def compute_header_string():
     header_string = ""
     for i in range(1, 413):
         header_string += "o{} ".format(i)
@@ -530,8 +530,7 @@ def compute_header_string(has_strata=False):
     header_string += "r1 r2 r3 r4 r5 r6 r7 "
     header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7 "
     header_string += "SDest SDist G A"
-    if has_strata:
-        header_string += "strata"
+    header_string += " strata"
     return header_string
 
 
@@ -572,7 +571,7 @@ def create_equal_person_vector_file(length):
     file_name = "person_vector_data_with_strata_{}.txt".format(length)
     with open(file_name, "w+") as file:
         # generate header
-        header_string = compute_header_string(True)
+        header_string = compute_header_string()
         file.write(header_string)
         file.write("\n")
 

### 2018-06-29 testdata/testDataGenerator.py
off by 1. again

@@ -522,9 +522,9 @@ def parse_lines_of_content():
 
 def compute_header_string():
     header_string = ""
-    for i in range(1, 413):
+    for i in range(1, 413 + 1):
         header_string += "o{} ".format(i)
-    for i in range(1, 413):
+    for i in range(1, 413 + 1):
         header_string += "d{} ".format(i)
     header_string += "AM MD PM MN "
     header_string += "r1 r2 r3 r4 r5 r6 r7 "

### 2018-06-29 testdata/testDataGenerator.py
strataGrouped

@@ -434,6 +434,7 @@ class Person:
     def compute_vector_with_strata(self):
         lst = self.compute_vector()
         lst.append("s_{}".format(self.get_strata()))
+        lst.append("s_{}".format(int((self.get_strata() + 1) / 2)))
         return lst
 
 
@@ -589,6 +590,13 @@ def create_equal_person_vector_file(length):
                 new_person_list.append(PERSONS[item])
 
             item += 1
+
+            if item == len(PERSONS) - 1:
+                print("WARNING     :: There are not enough elements to have {} persons of each strata".format(length))
+                for i in range(0, 6):
+                    print("WARNING     :: There are currently {} persons for strata {}".format(buckets[i], i + 1))
+                print("ERROR       :: failed current Task")
+                exit()
 
         # for person in new_person_list:
         for person in new_person_list:

### 2018-06-29 testdata/testDataGenerator.py
repush

@@ -530,8 +530,10 @@ def compute_header_string():
     header_string += "AM MD PM MN "
     header_string += "r1 r2 r3 r4 r5 r6 r7 "
     header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7 "
-    header_string += "SDest SDist G A"
-    header_string += " strata"
+    header_string += "SDest SDist G A "
+    header_string += "strata "
+    header_string += "strataGrouped"
+
     return header_string
 
 

### 2018-06-29 testdata/testDataGenerator.py
script push

@@ -610,6 +610,49 @@ def create_equal_person_vector_file(length):
                 file.write("\n")
 
 
+def create_equal_person_vector_file_aggregated(length, suffix=""):
+    if args.debug:
+        print("INFORMATION :: Vector Data of Persons")
+
+    file_name = "person_vector_data_with_strata_{}{}.txt".format(length, suffix)
+    with open(file_name, "w+") as file:
+        # generate header
+        header_string = compute_header_string()
+        file.write(header_string)
+        file.write("\n")
+
+        buckets = [0, 0, 0]
+        new_person_list = []
+        lst = list(range(len(PERSONS) - 1))
+        random.shuffle(lst)
+
+        item = 0
+        # for item in lst:
+        while sum(buckets) < 3 * length:
+            strata = int((PERSONS[item].get_strata() + 1) / 2)
+            if buckets[strata - 1] < length:
+                buckets[strata - 1] += 1
+                new_person_list.append(PERSONS[item])
+
+            item += 1
+
+            if item == len(PERSONS) - 1:
+                print("WARNING     :: There are not enough elements to have {} persons of each strata".format(length))
+                for i in range(0, 6):
+                    print("WARNING     :: There are currently {} persons for strata {}".format(buckets[i], i + 1))
+                print("ERROR       :: failed current Task")
+                exit()
+
+        # for person in new_person_list:
+        for person in new_person_list:
+            # if args.debug:
+            #     print_progressbar(i, len(PERSONS) - 1, prefix='INFORMATION ::', suffix='Complete', length=50)
+            # file.write(" ".join(str(x) for x in person.compute_vector()))
+            file.write(" ".join(str(x) for x in person.compute_vector_with_strata()))
+            if new_person_list.index(person) != len(new_person_list) - 1:
+                file.write("\n")
+
+
 # Print iterations progress
 def print_progressbar(iteration, total, prefix='INFORMATION :: ', suffix='', decimals=1, length=100, fill='█'):
     """
@@ -639,6 +682,7 @@ parser = argparse.ArgumentParser(
 parser.set_defaults(size=False)
 parser.set_defaults(debug=False)
 parser.set_defaults(vector=False)
+parser.set_defaults(agg=False)
 parser.set_defaults(distType="equal")
 
 # create  exclusive group
@@ -656,6 +700,8 @@ size_parser.add_argument('-size', metavar='size', type=int,
 
 parser.add_argument('-vector', dest='vector', action="store_true",
[…]

### 2018-06-29 dokumentation/submission.tex
space fixed
Neural Net avg, und submission

@@
 \usepackage{graphicx}
+%\usepackage{ngerman}
+\usepackage[utf8x]{inputenc}
+\usepackage{fancyvrb}
+\usepackage{courier}
+\usepackage{helvet}
+\usepackage{tikz}
+\usepackage{xcolor}
+\usepackage{pdfpages}
+\usetikzlibrary{calc}
+\usepackage[strict]{changepage}
+\usepackage{xspace}
+\usepackage{hyperref}
+\usepackage{cleveref}
+\usetikzlibrary{shapes.geometric}
+\usetikzlibrary{arrows}
+\usetikzlibrary{positioning}
+\usetikzlibrary{arrows.meta, automata, shapes, matrix,positioning}
+\usepackage{amssymb}
+\usepackage{pifont}% http://ctan.org/pkg/pifont
+\usepackage{subcaption} 
+\usepackage{float}
+\usepackage{fixfoot}
+\usepackage{graphicx}
+\usepackage{wrapfig}
+\usepackage{multicol}
+\usepackage{amsmath}
+\usepackage{listings}
+
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 % Used for displaying a sample figure. If possible, figure files should
@@
 	%
-	\author{Miriam NACHNAME\and
+	\author{Miriam Wagner\and
 		Martin Breuer\and
-		Moritz NACHNAME\and
+		Moritz Werthebach\and
 		Timo Bergerbusch\and
-		Walter NACHNAME}
+		Walter Schikowski}
 	%
@@
 	%
-	\begin{abstract}
+	\begin{abstract} %TODO
 		The abstract should briefly summarize the contents of the paper in
@@
 	\section{Preprocessing}
+	\begin{itemize}
+		\item max equal of 2038 for a strata 
+		\item max equal of 595 per persona in strata
+		\item vector calc.
+	\end{itemize}
+	\subsection{Vector}
+	Instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
+This class stores the ID, the parameters defining a person %TODO: ref zu code basics
+	, and all movements from that person.\\
+	Then we are able to compute the following vector, with 848 entries, for further usage, that combines all movements of the person:
+	\begin{align*}
+	\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,
+	\underbrace{\mathit{AM}, \mathit{MD}, \mathit{PM}, \mathit{MN}}_{4}, 
+	\underbrace{\#r_1, \dots, \#r_7}_{7}, \\
+	\underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, \underbrace{\mathit{SDest}, \mathit{SDist}, \mathit{G}, \mathit{A} ,\mathit{strata}, \mathit{strataGrouped}}_{6}
+	\end{align*}
+	with the following abbreviations ($1 \le i \le 413$, $1 \le j \le 7$):
+	\begin{multicols}{2}
+		\begin{itemize}
+			\setlength{\itemindent}{.4cm}
+			\item[$o_i$:]  the $i$-th origin data point
+			\item[$d_i$:]  the $i$-th destination data point
+			\item[$\mathit{AM}$:] movements at time stamp AM
+			\item[$\mathit{MD}$:] movements at time stamp MD
+			\item[$\mathit{PM}$:] movements at time stamp PM
+			\item[$\mathit{MN}$:] movements at time stamp MN
+			\item[$r_j$:] the $j$-th reason
+			\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
+			\item[$\mathit{SDest}$:] sum of all durations
+			\item[$\mathit{SDist}$:] sum of all distances
+			\item[$\mathit{G}$:] the gender
+			\item[$\mathit{A}$:] the age
+			\item[$strata$:] the strata (used for comparison)
+			\item[$strataGrouped$:] the aggregated stratas
+		\end{itemize}
+	\end{multicols}
+	\section{Predicting}
+
+	\subsection{Classification}
+	\subsection{Neural Net}
+	Also we have a max. of 595 persons for strata 6 to have an equally distributed test set.\\	
+	All using the vector data\\
+
+	We consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
+	As an example of other strata aggregation we combine the stratas 1--2, 3--4 and 5--6 together and call them $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$.
+For that we take an equal distribution of all original stratas and map them correspondingly to the new stratas.
+	\setlength\tabcolsep{.2cm}
+	\begin{figure}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|c|c|}
+		\hline
+		                         &   \#    &        & \multicolumn{3}{c|}{Set size} \\
+		          Name           & Neurons &   AG   &  100  &  200  &      595      \\ \hline
+		    $\mathcal{N}_5$      &    5    & \xmark & 60.03 & 59.92 &     60.18     \\
+		 $\mathcal{N}_5^\star$   &    5    & \cmark & 87.6  & 89.7  &     71.05     \\
+		   $\mathcal{N}_{10}$    &   10    & \xmark & 75.83 & 73.54 &     69.56     \\
+		$\mathcal{N}_{10}^\star$ &   10    & \cmark & 92.93 & 93.48 &     74.58     \\		
+		%		$\mathcal{N}_{10}^\star$ &   10    & \cmark & 88.33 {\small $\pm$7.49} & 90.67{\small $\pm$2.81} & 92.14 {\small $\pm$ 2.59} \\
+		   $\mathcal{N}_{20}$    &   20    & \xmark & 75.45 & 71.14 &     61.87     \\
+		$\mathcal{N}_{20}^\star$ &   20    & \cmark & 92.87 & 94.4  &     78.32     \\ \hline
+	\end{tabular}
+	\caption{The accuracy values of the neural nets. (See excel-spreadsheet)}
+	\label{tab: nn-accuracy}
+	% mittels deep-learning-vector-average
+	\end{figure}
+
+If we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be to small to have sufficient training.\\
+
+The greater the population, the more the borders between the stratas blur.
 

### 2018-06-29 testdata/testDataGenerator.py
strata as categorial

@@ -71,7 +71,10 @@ class DataEntry:
         s = ""
         for i in range(0, len(entry_mask)):
             if entry_mask[i]:
-                if i != 12:
+
+                if i == 8:
+                    s += "s_{}".format(self.data[i])
+                elif i != 12:
                     s += "{}".format(self.data[i])
                 else:
                     s += "{}".format(self.data[12].get_id())

### 2018-06-30 dokumentation/submission.tex
submission testdatagenerator

@@
 	\section{Preprocessing}
-	\begin{itemize}
-		\item max equal of 2038 for a strata 
-		\item max equal of 595 per persona in strata
-		\item vector calc.
-	\end{itemize}
-	\subsection{Vector}
-	Instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
+	In order to classify the given data into smaller test sets or mask different aspects, we have to perform analysis.\\
+	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
+	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments the script is able to create all test sets considered by our clustering and neural net approaches.\\
+	%TODO: should this be within the paper or more in a ReadMe-file delivered with the script itself?
+%	The possible flags and an explanation can be taken from \Cref{fig: testdatagenerator flags}.	
+%	\begin{figure}[H]
+%		\centering
+%		\includegraphics[scale=0.5]{src/pic/testDataGenerator-h.PNG}
+%		\caption{The possible arguments to the \texttt{testDataGenerator}}
+%		\label{fig: testdatagenerator flags}
+%	\end{figure}
+	We observe the following distribution over the whole dataset:\\
+	{\hspace*{2cm}\setlength\tabcolsep{.2cm}\begin{tabular}{c|ccccccc}
+		strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
+		 abs   & 6963 & 52265 & 49404 & 8772 & 5536 & 2038 &  124978  \\
+		  \%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
+	\end{tabular}}\\
+	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
+
+	In addition to the original paper we compute the value \texttt{ID}, which is used to combine movements considered to be from the same person.
+	We consider two movements to coincide on the underlying person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
+	{\hspace*{2cm}\setlength\tabcolsep{.2cm}\begin{tabular}{c|ccccccc}
+		strata &  1   &   2   &   3   &  4   &  5   &  6  & $\Sigma$ \\ \hline
+		 abs   & 3153 & 23367 & 21418 & 3497 & 2083 & 595 &  54113   \\
+		  \%   & 5.83 & 43.18 & 39.58 & 6.46 & 3.85 & 1.1 &   100
+	\end{tabular}}\\
+	So we also have through strata 6 an upper bound of 595 for equally distributed person vector data (see \Cref{subsec: person vector data}).
+
+	\subsection{Vector}\label{subsec: person vector data}
+	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
 This class stores the ID, the parameters defining a person %TODO: ref zu code basics

### 2018-06-30 testdata/testDataGenerator.py

@@ -275,7 +275,7 @@ class DataCollection:
         # the distribution in the original dataset
         relative_numbers = []
         for i in range(0, 6):
-            relative_numbers.append(absolute_numbers[i] / 124979)
+            relative_numbers.append(absolute_numbers[i] / 124978)
         abs_numbers = []
         new_data_entries = []
         for i in range(0, 6):
@@ -701,9 +701,13 @@ size_parser.add_argument('-size', metavar='size', type=int,
                               'possible)')
 
 parser.add_argument('-vector', dest='vector', action="store_true",
-                    help='create a vector of person data including the strata')
+                    help='create a vector of person data including the strata.'
+                         'This can be combined with the "-size size" and/or "-agg" flag ')
 
-parser.add_argument('-agg', dest='agg', action='store_true', help='TODO')
+parser.add_argument('-agg', dest='agg', action='store_true',
+                    help='creates the person vector data for the aggregation of strata 1-2, 3-4 and 5-6. '
+                         'Also the mapping is done before picking a possible test set and therefore may contain for '
+                         'the field of 1-2, only entries from strata 2')
 
 # debug output flag
 parser.add_argument('-debug', dest='debug', action='store_true', help='write (debug) information of current processes')
@@ -714,7 +718,7 @@ parser.add_argument('-debug', dest='debug', action='store_true', help='write (de
 # only if rapidminer can set flags
 
 # change the distribution
-parser.add_argument('-distType', metavar='distType', help='the type of distribution: (0) equal; (1) normal')
+parser.add_argument('-distType', metavar='distType', help='the type of distribution: "equal" (default); "normal"')
 # parse the arguments
 args = parser.parse_args()
 

### 2018-06-30 dokumentation/submission.tex
NN started

@@
 	\section{Introduction}
-	\section{Preprocessing}
+	\section{Preprocessing}\label{sec: proprocessing}
 	In order to classify the given data into smaller test sets or mask different aspects, we have to perform analysis.\\
@@
 	\subsection{Neural Net}
-	Also we have a max. of 595 persons for strata 6 to have an equally distributed test set.\\	
-	All using the vector data\\
+	For all the neural net computations done we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
+	We do this, because results on the normal datasets had an unacceptable performance. An example is given in \Cref{fig: NN without vector}.	
+	\begin{figure}
+		\centering
+		\includegraphics[scale = 0.4]{src/pic/NN_without_vector.png}
+		\caption{An example of a neural net trained without person vector data.}
+		\label{fig: NN without vector}
+	\end{figure}
 
-	We consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
+	In the following we consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
 	As an example of other strata aggregation we combine the stratas 1--2, 3--4 and 5--6 together and call them $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$.
-For that we take an equal distribution of all original stratas and map them correspondingly to the new stratas.
+This builds a superset of the original stratas and since the stratas themselves are logically connected this task should be easier to fulfill.\\
+	The sets are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
 	\setlength\tabcolsep{.2cm}
@@
 
-If we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be to small to have sufficient training.\\
+	For each we performed 5 independent runs and take an average over those accuracy values in order to have a sophisticated statement.\\
 
-The greater the population, the more the borders between the stratas blur.
+	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be to small to have sufficient training.\\
+
+	%TODO: what can we oberve?
+%	The greater the population, the more the borders between the stratas blur.
 

### 2018-06-30 dokumentation/clustering_algorithms.tex
Began paper section on clustering algorithms

\documentclass[11pt]{article}
\usepackage{amsmath} % Math
\usepackage{amssymb} % Math symbols
\usepackage[english]{babel} % Language
\usepackage{fancyhdr} % Header
\usepackage[a4paper, total={15cm, 20cm}]{geometry} % Dimensions of the paper and the text area
\usepackage[utf8]{inputenc} % encoding in UTF, needed for umlauts if German
\usepackage{mathtools} % Text above arrows
\usepackage{msc} % Drawing MSCs
\usepackage{multicol} % Multiple columns
\usepackage[explicit]{titlesec} % Automatic section titles
\usepackage{tikz} % Diagrams
\usetikzlibrary{arrows.meta, automata, shapes, matrix}

% Other packages that might be useful in the future
\usepackage{mathptmx,amssymb,amsmath}
%\usepackage{lingmacros}
%\usepackage{tree-dvips}
%\usepackage{ulem}
%\usepackage{amsthm}
%\usepackage{amsbsy}
%\usepackage{textcomp,gensymb}
%\usepackage{graphicx}
%\usepackage{mathtools}

% Custom variant of msc environment:
% - No "msc" keyword, longer partial messages
% - Increased vertical distance between messages
% - Less distance to the frame left and right
% - Less distance between header and processes
% - Less distance between footer and frame
% - Passing all given options down to the msc environment
\newenvironment{cmsc}[1][]{\msc[msc keyword={}, self message width=1.1cm, level height=0.6cm, environment distance=1.2cm, head top distance=0.75cm, foot distance=0.5cm, #1]}{\endmsc}

% No indentation at new paragraphs
\setlength{\parindent}{0pt}

% Distance between columns
\setlength{\columnsep}{1cm}
% Vertical line between columns
\setlength{\columnseprule}{0.5pt}
\def\columnseprulecolor{\color{gray}}

% Settings
%\newcommand{\sheetNr}{1}

%% Header
\fancyhf{}
\pagestyle{fancy}
\lhead{Clustering of Mobility Data}
%\rhead{Timo Bergerbusch: 344408}
\setlength{\headheight}{28pt}

%% Automatic section titles
%\titleformat{\section}{\normalfont\Large\bfseries}{}{0em}{Exercise #1}
%\titleformat{\subsection}{\normalfont\large\bfseries}{}{0em}{#1)}

\begin{document}

\section*{Clustering the data}

In the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} is a process of grouping data objects from a dataset into multiple groups or clusters. The essential criterion for the quality of the clustering is a certain \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.

In the scope of this work, we decided to use well-known partitioning methods such as k-means or k-medoids. In general, given $n$ data objects, partitioning methods distribute the data object into $k$ clusters with $k<n$, using a distance measure to evaluate the repsective similarity. Partitioning methods form exclusive clusters by ensuring that each cluster contains at least one object. Note that the number $k$ of clusters has to be chosen manually prior to the partitioning process.

\subsection{k-Means}

As pointed above, k-Means is one of the most used clustering algorithms. k-means is a \textit{centroid based technique}, which means that each cluster is represented by a data point that also is the center of the cluster. A distance measure is then used to assigns every remaining data object from the data set to the best fitting cluster according to it's similarity to the center of this cluster and it's dissimilarity to the centers of any other cluster.

The data objects within a dataset are considered to reside in a euclidean space. Thus, the euclidean distance is used to calculate a score for the similarity of two data points. When using k-Means, the quality of a cluster $C_i$ can b1e evaluated by computing the sum of squared erros between all data points and the centroid in $C_i$. This method is known as \textit{within-cluster variation} and defined as follows:

[TODO: LEAST SQUARES]

Given $k$, the first step of k-Means is to select $k$ random objects in the data set as initial representatives for the cluster centers. After that, each remaining data object is assigned to the best fitting cluster in terms of similarity, based on a similarity score that is determined by the euclidean distance between the data point and every cluster center. The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation. In order to achieve this, for each cluster centroid is reassigned as the mean of all objects within that cluster. By considering the updated centroids, every data point is redistributed to the now best fitting cluster. This iterative process will continue, until no better clustering can be found, which means that the latest clustering equals the previous distribution.

[TODO: ALGORITHMUS]

\subsection{k-Medoids}

By the choic1e of using euclidean distance as a similarity meaasure, k-Means is sensitive to outliers in a wây, that they can heavily distort the mean of a cluster and, as a result, worsen the within-cluster variation.

k-medoids is a modification of k-means,

\end{document}

### 2018-07-01 testdata/testDataGenerator.py
praesi

@@ -437,7 +437,8 @@ class Person:
     def compute_vector_with_strata(self):
         lst = self.compute_vector()
         lst.append("s_{}".format(self.get_strata()))
-        lst.append("s_{}".format(int((self.get_strata() + 1) / 2)))
+        if args.agg:
+            lst.append("s_{}".format(int((self.get_strata() + 1) / 2)))
         return lst
 
 
@@ -536,7 +537,8 @@ def compute_header_string():
     header_string += "MoT1 MoT2 MoT3 MoT4 MoT5 MoT6 MoT7 "
     header_string += "SDest SDist G A "
     header_string += "strata "
-    header_string += "strataGrouped"
+    if args.agg:
+        header_string += "strataGrouped"
 
     return header_string
 

### 2018-07-08 Programmierkram/python/gower_implementation/gower_custom.py
commented final gower code. The old testfiles were moved to "gower_discarded"

@@ -1,26 +1,0 @@
-# Origin | Destination | Reason | Mean of Transportation | Average Time | Time Interval | Duration | Distance | Age | Gender
-
-
-categorical = [True, True, True, True, False, True, False, False, False, True]
-
-def distance (X, Y, categorical):
-    # Check array sizes
-    
-    if (len(X) != len(Y)):
-           print("Array sizes do not Match!")
-           return;
-    
-    # Algorithm
-    
-    s = [0 for i in range(len(X))]
-    
-    ranges = [0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
-       
-    for i in range(len(X)):
-        if (categorical[i] == True):
-            s[i] = 1 if (X[i] == Y[i]) else 0
-        else:
-            s[i] = 1 - abs(X[i] - Y[i]) / ranges[i]
-            
-    return sum(s)/len(X) 
-    

### 2018-07-08 Programmierkram/python/gower_implementation/gower_final_code.py

@@ -1,0 +1,49 @@
+"""
+Custom implementation of the Gower distance in the scope of the practicle course on the topic "clustering of mobility
+data" at HumTec, RWTH, SS18.
+
+parameters:
+
+(list) X, Y:        Data objects with variables of arbitrary types (Numerical, Categorical, [Binary]).
+(list) categorical: Value at a position i is 'True', if the variable type of X[i], Y[i] are categorical, else 'False'.
+(list) ranges:      Value at a position i is the value range of the variable type of X[i], Y[i]
+"""
+
+def distance (X, Y, categorical, ranges):
+    # Check array sizes
+    
+    if (len(X) != len(Y)):
+           print("Array sizes do not Match!")
+           return;
+    
+    # Algorithm
+    
+    s = [0 for i in range(len(X))]
+       
+    for i in range(len(X)):
+        if (categorical[i] == True):
+            s[i] = 1 if (X[i] == Y[i]) else 0
+        else:
+            s[i] = 1 - abs(X[i] - Y[i]) / ranges[i]
+            
+    return sum(s)/len(X) 
+
+"""
+EXAMPLE
+
+The data points have the form X, Y = (TravelDistance | Age | Gender),
+whereas 'Gender' is a categorical variable, given in the 'categorical' list.
+The ranges to each variable type are saved in the 'ranges' list.
+"""
+
+"""
+X = [87, 40, "M"]
+Y = [98, 46, "M"]
+
+categorical = [False, False, True]
+ranges = [250, 90, 0]
+
+print(distance(X, Y, categorical, ranges))
+
+# Output: 0.963111111111111
+"""

### 2018-07-09 dokumentation/submission.tex
kleine Fehler
Neural net Teil erweitert

@@
 \usepackage{hyperref}
-\usepackage{cleveref}
 \usetikzlibrary{shapes.geometric}
@@
 \usepackage{amsmath}
+\usepackage{cleveref}
 \usepackage{listings}
@@
 	\subsection{Classification}
+%	\input{MiriamClustering}
 	\subsection{Neural Net}
-	For all the neural net computations done we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
-	We do this, because results on the normal datasets had an unacceptable performance. An example is given in \Cref{fig: NN without vector}.	
-	\begin{figure}
+	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
+
+	We do this, because results on the normal datasets had an unacceptable performance, since only single movements and not complete paths of individuals are considered.
+An example training and performance measure is given in \Cref{fig: NN without vector} where unprocessed data is used.
+The performance is measured using 10-fold cross validation, i.
+e.
+the data is split into 10 subsets where in each iteration exactly one data set is used as test set and the other 9 as training set.
+The average value of the accuracy values lead
+	 to the total accuracy value of the neural net.
+	\begin{figure}[H]
 		\centering
@@
 This builds a superset of the original stratas and since the stratas themselves are logically connected this task should be easier to fulfill.\\
-	The sets are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
+
+	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata which are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
+For every neural net and every set size we performed 5 independent runs and calculated the average over those accuracy values in order to have a sophisticated, comparable statements.
 	\setlength\tabcolsep{.2cm}
@@
 
-	For each we performed 5 independent runs and take an average over those accuracy values in order to have a sophisticated statement.\\
-
 	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be to small to have sufficient training.\\

### 2018-07-10 dokumentation/gower_distance.tex
gower documentation overhaul

@@
 \usepackage[explicit]{titlesec} % Automatic section titles
-\usepackage{tikz} % Diagrams
-\usetikzlibrary{arrows.meta, automata, shapes, matrix}
+
+\usepackage{biblatex}
 
@@
 
+TODO: citations
+
 When clustering data with algorithms such as \textit{k-means} or \textit{k-medoids}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm.
 These distance measures come with constraints though, since they are only defined for numerical variables.
-In the scope of this work, we are facing data points with mixed variable types, therefore our choice of an appropriate distance measure fell on the Gower distance measure to calculate an overall \textit{similarity} between mixed data points.
-The gower similarity distance distinguishes between three types of variables:\\
+In the scope of this work, we are facing data points with mixed variable types, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points.
+The Gower distance measure distinguishes between three types of variables:\\
 
-\textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match [CITE].
+\textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match \cite{gower1971general}.
 In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}.
@@
 
-\textbf{Categorical} variables form a set of unordered values and are comparible to ENUMs in programming languages.
+\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.
 
@@
 Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
-\begin{equation*}
+\begin{equation}\label{eq1}
   S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
-\end{equation*} 
-where $s_{xy,k}$ denotes a score for the similarity of variables of index $k$ between the data points $x$ and $y$, that's definition is dependent on the type of the variable, as defined below.
-In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, where the value 1 means, when variables are comparable and 0 when not.
-This could happen, if some values are not defined.
-Within this work, the dataset is complete, therefore, the sum of $\delta_{xy,k}$ for all $0 <= k <= v$ equals to $v$, the number of considered variables in each data point.
-Thus, the similarity coefficient can be interpreted as the average value of all similarity scores.\\
+\end{equation} 
+where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at position $k$ in the data points $x$ and $y$.
+Note, that the definition of the score dependends on the type of the variable, as defined below.
+In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, when the two variables are comparable and to 0 when not.
+For example, variables are not comparable, if values are undefined in the data points or the variable types do not match at index $k$, among other reasons.
+Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$, the number of considered variables in each data point, for all $x, y$ in the data set.
+Thus, the similarity coefficient in (\ref{eq1}) can be interpreted as the average value of all similarity scores.\\
 
-With respect to the variable type, $s_{xy,k}$ is defined by
+With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:
 \begin{itemize}
@@
     \item \textbf{Numerical:} for numerical variables, the similarity score is calculated by
-    \begin{equation}
+    \begin{equation*}
       s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
-    \end{equation}
+    \end{equation*}
     where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
@@
 \end{itemize}
-TODO: sources, citations
+
+\begin{thebibliography}{9}
+
+	\bibitem{gower1971general}
+	Gower, John C,
+	\emph{A general coefficient of similarity and some of its properties},
+	Biometrics, JSTOR pp. 857-871,
+	1971.
+
+\end{thebibliography}
+
 \end{document}

### 2018-07-10 dokumentation/submission.tex
small fixes/additions

@@
 	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
+	Furthermore we have to make sure that  two different data points, which belong to the very same person, are assigned to the same cluster.
+To do so we compute the value \texttt{ID} which identifies each person and can be used to combine movements that are considered to be from the same person, i.
+e.
+two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
+This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.
 
-	In addition to the original paper we compute the value \texttt{ID}, which is used to combine movements considered to be from the same person.
-	We consider two movements to coincide on the underlying person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
 	{\hspace*{2cm}\setlength\tabcolsep{.2cm}\begin{tabular}{c|ccccccc}
@@
 	\end{tabular}}\\
-	So we also have through strata 6 an upper bound of 595 for equally distributed person vector data (see \Cref{subsec: person vector data}).
+	In Section \Cref{subsec: person vector data} we introduce vectors representing single persons.
+Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.
 
@@
 
-	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be to small to have sufficient training.\\
+	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be too small to have sufficient training.\\
 

### 2018-07-10 dokumentation/submission.tex
Abstract hinzugefügt

@@
 	\begin{abstract} %TODO
-		The abstract should briefly summarize the contents of the paper in
-		150--250 words.
+		We analyze movements in the urban environment of the columbian city Medellín.
+Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
+Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
+Due to bad results we introduce our preprocessing steps and switch to supervised learning.
+Decision trees and neural networks did neither match our performance expectations which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
 
@@
 	\subsection{Classification}
-%	\input{MiriamClustering}
+	\input{MiriamClustering}
 	\subsection{Neural Net}

### 2018-07-10 dokumentation/MiriamClustering.tex
relative Pfade gefixt

--- 2018-06-27 Miriam-RapidMiner/Clustering.tex
@@
+%Miriam Clustering
+
 \section{Clustering with RapidMiner}
@@
 
-\includegraphics[width=0.9\textwidth]{ClusteringRapid.PNG}
+\includegraphics[width=0.9\textwidth]{../Miriam-RapidMiner/ClusteringRapid.PNG}
 
@@
 	\item[Nominal to Numerical] This changes the nominal data to numerical data, so that we can apply PCA in the next step
-	\item[PCA] Here we apply the PCA reduction to the data with a variance threshold $0.95$
+	\item[PCA] Here we apply the PCA reduction to the data with a variance threshold $0.95$ \colorbox{red}{where?}
 	\item[Clustering] Here the number of Clusters has to be fixed.
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidStrata.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/ClusterPCAOrigRapidStrata.PNG}
   \caption{Strata}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterPCAOrigRapidCluster.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/ClusterPCAOrigRapidCluster.PNG}
   \caption{Cluster}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterOrigRapidStrata2Cluster.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/ClusterOrigRapidStrata2Cluster.PNG}
   \caption{Strata}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterOrigRapidCluster2Cluster.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/ClusterOrigRapidCluster2Cluster.PNG}
   \caption{Cluster}
@@
 Also the distribution in the strata groups of the clustering shows, that there is no real connection of strata and the clustering.
-\includegraphics[width=0.9\textwidth]{ClusterOrigRapidDistribution2Cluster.PNG}
+\includegraphics[width=0.9\textwidth]{../Miriam-RapidMiner/ClusterOrigRapidDistribution2Cluster.PNG}
 
@@
   \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringcluster.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/vectorclusteringcluster.PNG}
   \caption{Strata}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringstrata.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/vectorclusteringstrata.PNG}
   \caption{Cluster}
@@
 Also the distribution in the strata groups of the clustering shows, that there is not really a connection of strata and the clustering to see.
-\includegraphics[width=0.9\textwidth]{vectorClustering.PNG}
+\includegraphics[width=0.9\textwidth]{../Miriam-RapidMiner/vectorClustering.PNG}
 
@@
   \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringcluster1000.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/vectorclusteringcluster1000.PNG}
   \caption{Strata}
@@
 
-\includegraphics[width=0.9\textwidth]{vectorClustering1000}
+\includegraphics[width=0.9\textwidth]{../Miriam-RapidMiner/vectorClustering1000}
 
@@
   \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringcluster3Cluster.PNG}
+  \includegraphics[width=.4\linewidth]{../Miriam-RapidMiner/vectorclusteringcluster3Cluster.PNG}
   \caption{Strata}
@@
 
-\includegraphics[width=0.9\textwidth]{vectorClustering3Cluster.PNG}
+\includegraphics[width=0.9\textwidth]{../Miriam-RapidMiner/vectorClustering3Cluster.PNG}
 

### 2018-07-10 dokumentation/submission.tex
Introduction and Observation

--- 2018-07-09 dokumentation/submission.tex
@@
 \begin{document}
+
 	%
-	\title{Clustering Analysis of Mobility Data}
+%	\title{Clustering Analysis of Mobility Data}
+	\title{Are you moving predictably?}
 	%
@@
 	%
+	%TODO: Order by Lastname?
 	\author{Miriam Wagner\and
@@
 
-		\keywords{Clustering \and Rapidminer \and Cluster \and Data Mining} %TODO: maybe more
+		\keywords{Data Mining \and Clustering \and Rapidminer \and Cluster \and Neural Nets} %TODO: maybe more
 	\end{abstract}
@@
 	\section{Introduction}
+%	TODO (1 page)
+	% could at more in depth 
+	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the data collected?
+In particular is it possible to determine the wealth of a person, only given movements of a single day?
+	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
+There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín.
+All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
+	The data entries consist of data about the movement, like endpoints or length, and also some meta parameters like gender, age or the so-called strata of the person.
+The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.\\
+	Our goal is to ascertain if there is a correlation between the movements and the strata.
+
 	\section{Preprocessing}\label{sec: proprocessing}
-	In order to classify the given data into smaller test sets or mask different aspects, we have to perform analysis.\\
+	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
 	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
-	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments the script is able to create all test sets considered by our clustering and neural net approaches.\\
-	%TODO: should this be within the paper or more in a ReadMe-file delivered with the script itself?
-%	The possible flags and an explanation can be taken from \Cref{fig: testdatagenerator flags}.	
-%	\begin{figure}[H]
-%		\centering
-%		\includegraphics[scale=0.5]{src/pic/testDataGenerator-h.PNG}
-%		\caption{The possible arguments to the \texttt{testDataGenerator}}
-%		\label{fig: testdatagenerator flags}
-%	\end{figure}
+	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets considered by our clustering and neural net approaches.\\
 	We observe the following distribution over the whole dataset:\\
-	{\hspace*{2cm}\setlength\tabcolsep{.2cm}\begin{tabular}{c|ccccccc}
+	\begin{figure}[H]
+		\centering
+		\setlength\tabcolsep{.2cm}
+		\begin{tabular}{c|ccccccc}
 			strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
@@
 			\%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
-	\end{tabular}}\\
+		\end{tabular}
+		\label{table: distribution normal}
+	\end{figure}
 	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
@@
 	We consider two movements to coincide on the underlying person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
-	{\hspace*{2cm}\setlength\tabcolsep{.2cm}\begin{tabular}{c|ccccccc}
+	\begin{figure}[H]
+		\centering
+		\setlength\tabcolsep{.2cm}
+		\begin{tabular}{c|ccccccc}
 			strata &  1   &   2   &   3   &  4   &  5   &  6  & $\Sigma$ \\ \hline
@@
 			  \%   & 5.83 & 43.18 & 39.58 & 6.46 & 3.85 & 1.1 &   100
-	\end{tabular}}\\
+		\end{tabular}
+	\end{figure}
 	So we also have through strata 6 an upper bound of 595 for equally distributed person vector data (see \Cref{subsec: person vector data}).
 
-	\subsection{Vector}\label{subsec: person vector data}
+	\subsection{Stratified Person Data}\label{subsec: person vector data}
 	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
@@
 	, and all movements from that person.\\
-	Then we are able to compute the following vector, with 848 entries, for further usage, that combines all movements of the person:
+	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:
 	\begin{align*}
@@
 	\underbrace{\#r_1, \dots, \#r_7}_{7}, \\
-	\underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, \underbrace{\mathit{SDest}, \mathit{SDist}, \mathit{G}, \mathit{A} ,\mathit{strata}, \mathit{strataGrouped}}_{6}
+	\underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, \underbrace{\mathit{S_{Dest}}, \mathit{S_{Dist}}, \mathit{G}, \mathit{A} ,\mathit{strata}, \mathit{strataGrouped}}_{6}
 	\end{align*}
@@
 			\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
-			\item[$\mathit{SDest}$:] sum of all durations
-			\item[$\mathit{SDist}$:] sum of all distances
+			\item[$\mathit{S_{Dest}}$:] sum of all durations
+			\item[$\mathit{S_{Dist}}$:] sum of all distances
 			\item[$\mathit{G}$:] the gender
@@
 
+	\subsection{Distance Measures}
+	%TODO: distance measures
+	TODO (1 page)
+
 	\subsection{Classification}
+	%TODO: clustering
+	TODO (5 pages)	
 %	\input{MiriamClustering}
@@
 
+	%TODO: recompute
 	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be to small to have sufficient training.\\
 
-	%TODO: what can we oberve?
+	\section{Observations}
+%	TODO (1 page)
+%	TODO: restructure. Have like 3 parts
 %	The greater the population, the more the borders between the stratas blur. 
+	As an overall result we get, that we can not determine the strata based on the information we have.
+	Using various datasets (c.f.
+\Cref{sec: proprocessing}), we witnessed that having equally distributed datasets lead to an overall higher accuracy.
+This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f.
+\Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
+	During clustering we observed, that using a different distance measure formula would not lead to a huge difference.
+Also, we detected that reducing the numbers of clusters leads to better results, which reduces the significance of result that could be stated.\\
+	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
+	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
+Therefore we see that the more data we have the higher the variance within the stratas themselves is.
+We cannot draw some kind of lines to distinguish between different entries.
+Datapoints, which were outliers in the smaller sets, are now not considered to be outliers, because of large numbers having the same characteristics.
+So the lines, we were able to draw for smaller datasets, are blurring and create some kind of transition phase.\\ %TODO: letzter satz doof
 
+	\section{Discussion}
+	%TODO: discussion
+	TODO (2 page)
+
+\begin{thebibliography}{8}
+	\bibitem{rich_do_not_rise_early}
+	Lotero, Laura, et al. "Rich do not rise early: spatio-temporal patterns in the mobility networks of different socio-economic classes." Royal Society open science 3.10 (2016): 150654.
+
+	\bibitem{strata_def}
+	Hudson, Rex A. Colombia: A country study. Government Printing Office, 2010.
+
+%	\bibitem{ref_lncs1}
+%	Author, F., Author, S.: Title of a proceedings paper. In: Editor,
+%	F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
+%	Springer, Heidelberg (2016). \doi{10.10007/1234567890}
+%	
+%	\bibitem{ref_book1}
+%	Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
+%	Location (1999)
+%	
+%	\bibitem{ref_proc1}
+%	Author, A.-B.: Contribution title. In: 9th International Proceedings
+%	on Proceedings, pp. 1--2. Publisher, Location (2010)
+%	
+%	\bibitem{ref_url1}
+%	LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
+%	Oct 2017
+\end{thebibliography}
 \end{document}

### 2018-07-10 dokumentation/submission.tex
Hatte was vergessen :dash:

@@
 	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
-
-	In addition to the original paper we compute the value \texttt{ID}, which is used to combine movements considered to be from the same person.
-	We consider two movements to coincide on the underlying person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
+	Furthermore we have to make sure that  two different data points, which belong to the very same person, are assigned to the same cluster.
+To do so we compute the value \texttt{ID} which identifies each person and can be used to combine movements that are considered to be from the same person, i.
+e.
+two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
+This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.
 	\begin{figure}[H]
@@
 	\end{figure}
-	So we also have through strata 6 an upper bound of 595 for equally distributed person vector data (see \Cref{subsec: person vector data}).
+	In Section \Cref{subsec: person vector data} we introduce vectors representing single persons.
+Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.
 
@@
 	%TODO: recompute
-	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be to small to have sufficient training.\\
+	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be too small to have sufficient training.\\
 
@@
 
-		%	\bibitem{ref_lncs1}
-		%	Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-		%	F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-		%	Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-		%	
-		%	\bibitem{ref_book1}
-		%	Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-		%	Location (1999)
-		%	
-		%	\bibitem{ref_proc1}
-		%	Author, A.-B.: Contribution title. In: 9th International Proceedings
-		%	on Proceedings, pp. 1--2. Publisher, Location (2010)
-		%	
-		%	\bibitem{ref_url1}
-		%	LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-		%	Oct 2017
 	\end{thebibliography}

### 2018-07-11 dokumentation/submission.tex

@@
 	%
-	\section{Introduction}
+	\section{Introduction} \label{sec: introduction}
 	%	TODO (1 page)
@@
 
-	\subsection{Distance Measures}
+	\subsection{Distance Measures} \label{subsec: distance measures}
 	%TODO: distance measures
@@
 
-	\subsection{Classification}
+	\subsection{Classification} \label{subsec: classification}
 	%TODO: clustering
@@
 	%	\input{MiriamClustering}
-	\subsection{Neural Net}
+	\subsection{Neural Net} \label{subsec: neural net}
 	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
@@
 
-	\section{Observations}
+	\section{Observations} \label{sec: observation}
 	%	TODO (1 page)
@@
 
-	\section{Discussion}
+	\section{Discussion} \label{sec: discussion}
 	%TODO: discussion
-	TODO (2 page)
+%	TODO (2 page)
+%	\begin{itemize}
+%		\item repr. working day
+%		\item over/under std's
+%		\item indiv. of lifestyle
+%		\item inv. behaviour in same strata
+%	\end{itemize}
+	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we though about various aspects having an impact on the movement.
+We state 4 main influencing aspects, explaining the variance throughout the data.
+
+	\subsection{Representativity of the day}
+	The day the data was collected on (c.f.
+\Cref{sec: introduction}) is not mentioned.
+Therefore we can not infer that it is representative.
+The people asked to enter their movement could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
+\\
+	Also the day itself is not mentioned, so we don't have information if it was even a working day or a weekend. This influences the behavior drastically.
+
+	\subsection{People living over/under standards}
+	%TODO
+
+	\subsection{Individuality of lifestyle}
+	Obviously, people are individual in their way of life.
+So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply don't like driving.
+On the other end of the spectrum, some people, who have little money still own and drive a car daily in order to go to work, since they can't afford an other apartment.\\
+	One can think of multitude scenarios, of people behaving different, cause by their individuality.
+
+	\subsection{Inverse behavior in a strata}
+	As an example, take a look at the strata 6, which denotes the richest people considered. Their we have inter alia two groups: 
+	\begin{enumerate}
+		\setlength{\itemindent}{1cm}
+		\item[1. Group]
+		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
+		\item[2. Group]
+		People, who are rich just by birth, which don't have the necessity to work or move at all. They could possibly stay home and don't have to leave at all.
+	\end{enumerate} 
+	Those two groups are completely opposite, but still belong to the same strata.
+Now also considering strata one, we can find the exact same movement behavior in there as well.
+Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work.
+Also there are people, who don't move at all, because of the same reason.
+
+	So we can see that based on the information we have it is very unlikely to have a precise correlation between the given data and the strata.
+However given more information about the previously mentioned aspects and more entries in general there could be some kind correlation to be found.\\
+	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
+Also one could use different network visualization tools in order to infer their own patterns.
+An example would be to take the origin (destination) sectors as nodes, an edge if there is a movement, directed or undirected, and a different thickness or color gradient based on the number of this edge being taken.\\
+
+	Overall we can conclude that with the data we got we are not able to find a correlation.
 

### 2018-07-11 dokumentation/MiriamClustering.tex
miriam ist dumm

--- 2018-06-27 Miriam-RapidMiner/Clustering.tex
@@
-\section{Clustering with RapidMiner}
+%Miriam Clustering
+The first question was: is it possible without knowing the social classes to reproduce them based on the movement data.
+Therefore we wanted to look for clusters and compare those with the strata.
+Also we had a look if the possibly found clusters have special properties.
+\subsubsection{Clustering with RapidMiner}
 
-RapidMiner has different Modules for Clustering already implemented. We apply the k-Means algorithm with mixed measure (euclidean).
+RapidMiner has different Modules for Clustering already implemented. We decided to concentrate on the k-means clustering algorithm.
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.9\textwidth]{ClusteringRapid.PNG}
+\caption{Process of k-means clustering}
+\label{fig: kclust}
+\end{figure}
 
-\includegraphics[width=0.9\textwidth]{ClusteringRapid.PNG}
-
-The Process does the following steps:
+The Process, figure \ref{fig: kclust} contains the following steps:
 \begin{description}
-	\item[Retrieve] This Block gives the data in the process. For the comparison at the end we also retrieve the original data.
-	\item[Nominal to Numerical] This changes the nominal data to numerical data, so that we can apply PCA in the next step
-	\item[PCA] Here we apply the PCA reduction to the data with a variance threshold $0.95$
-	\item[Clustering] Here the number of Clusters has to be fixed.
-We decided, that 10 runs should be enough.
-We can choose between differen measure types and tried mixed euclidean and squared euclidean distance
-	\item[Select Attribute] Here we just pick the Strata data for comparing with the clustering
-	\item[Join] For comparing the Clustering and the Strata we join the two filtered data sets at the ID
+	\item[Retrieve] gives the data into the process. 
+  \item[Generate ID] creates an ID such that we can make the comparsion step at the end through joining the sets
+  \item[Multiply] creates two identical data sets
+  \item[Select Attrbiutes] thoughs away the strata before the clustering step, everything behalve cluster and id after the clustering and just keeps id and strata for the join step
+	\item[Clustering] runs the k-means clustering algorithm. The number of Clusters has to be fixed.
+	\item[Join] For comparing the clustering result and the strata we join the two filtered data sets by the id
 \end{description}
 
-\subsection{Original Data}
+In the clustering block we can chose between different distance measures and maximal step numbers.
+We decided to concentrate on almost everywhere basic configurations and chose the squared euclidean distance in the mixed version.
 
-\subsubsection{Searching for 6 Cluster}
-Applying this process to the original data you can see without alignment of strata and cluster, that the distribution of cluster and strata is not close to the same.
+In the first step we tried to cluster the \textbf{Original Data} in \textbf{6 Cluster}. Therefore we retrieved the original data set in RapidMiner and chose k as 6.
 
 \ref{fig:OrgDist}. 
-\begin{figure}
+\begin{figure}[!htbp]
 \centering
@@
 
-\subsubsection{Searching for 3 Cluster}
+In figure \ref{fig:OrgDist} is the result to see of the first try.
+Figure \ref{fig:OrgSt} shows the strata distribution as pie chart and \ref{fig:OrgCl} the resulted cluster distribution.
+It can already been seen, that the distributions are not similar.
+In the next step we tried it with more steps, but the result was not looking better.
 
-The next idea is, that there are maybe just 3 obvious cluster to see and contain two 2 strata.
-
-\subsubsection{Applying on original data}
-
-For a clustering of 3 clusters we get a distributions that are complete different.\ref{fig:OrgDist}
-\begin{figure}[h]
+We asked ourselfs, if 6 cluster is not too fine, so we searched for \textbf{3 clusters} in the next step. The idea is to combine two stratas in 1, such that we just have 3 stratas left.
+\begin{figure}[!htbp]
 \centering

### 2018-07-11 dokumentation/DecisionTree.tex

\subsection{Decision Tree} \label{subsec: decisiontree}
Decision trees are a good manner to figure out, which parts of the data set have the most influence on the decision. We used again \textbf{RapidMiner} for building trees based on different data sets.

\begin{figure}[!htbp]
\centering
\includegraphics[width = 0.9\textwidth]{DecisionTreeRapidModel.PNG}
\caption{Process for decision trees in RapidMiner}
\label{fig: RapDec}
\end{figure}

RapidMiner does the following steps, to see in figure \ref{fig: RapDec}:
\begin{description}
	\item[Retrieve] includes the dataset
	\item[Generate Attribute] changes the Numerical Attribute strata to four binomial ones
	\item[Set Role] gives strata the label role, so that the decision tree has those as leafs
	\item[Decision Tree] creates the decision tree
\end{description}

Furthermore we choose information gain as splitting criterium (minimal gain 0.1) and a confidence of 0.25. Other configuration does not show different results.

In the first step we applied the process on the whole data set and the resulting tree was just the leaf "strata 2". So we tried it with different other data sets and the best result we got was for \textbf{stratified person data} equally distributed 3 stratas and just 200 in every aggerated strata group.

\begin{figure}[!htbp]
\centering
\begin{subfigure}{0.7\textwidth}
\includegraphics
{}

### 2018-07-11 dokumentation/MiriamClustering.tex

@@
 Also we had a look if the possibly found clusters have special properties.
+
 \subsubsection{Clustering with RapidMiner}
@@
 
-The Process, figure \ref{fig: kclust} contains the following steps:
+The Process, figure \ref{fig: kclust}, contains the following steps:
 \begin{description}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterOrigRapidStrata.PNG}
+  \includegraphics[width=.8\linewidth]{ClusterOrigRapidStrata.PNG}
   \caption{Strata}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterOrigRapidCluster.PNG}
+  \includegraphics[width=.8\linewidth]{ClusterOrigRapidCluster.PNG}
   \caption{Cluster}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterOrigRapidStrata2Cluster.PNG}
+  \includegraphics[width=.8\linewidth]{ClusterOrigRapidStrata2Cluster.PNG}
   \caption{Strata}
@@
   \centering
-  \includegraphics[width=.4\linewidth]{ClusterOrigRapidCluster2Cluster.PNG}
+  \includegraphics[width=.8\linewidth]{ClusterOrigRapidCluster2Cluster.PNG}
   \caption{Cluster}
@@
 \end{subfigure}
-\caption{Distribution of original data}
-\label{fig:OrgDist}
+\caption{Distribution of original data for just 3 clusters}
+\label{fig:OrgDist3Cl}
 \end{figure}
 
-Also the distribution in the strata groups of the clustering shows, that there is no real connection of strata and the clustering.
+\begin{figure}[!htbp]
+\centering
 \includegraphics[width=0.9\textwidth]{ClusterOrigRapidDistribution2Cluster.PNG}
-
-\subsection{Combined Data}
-
-Because of the not really convincing result of the clustering like above we had the idea to combine for every ID the different pathes in one big vector.
-This garants us, that one person just can be in one cluster aswell.
-The idea is to sum up all pathes in one big vector.
-\subsubsection{Code}
-%hier darfst du dich gerne verewigen Timo. Gerne auch vorher und nachher noch mehr schreiben. Titel darf auch geändert werden
-% boa das ist voll liep un soo weischt :D
-
-Instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person %TODO: ref zu code basics
-, and all movements from that person.\\
-Then we are able to compute the following vector, with 848 entries, for further usage, that combines all movements of the person:
-$$\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,
- \underbrace{\mathit{AM}, \mathit{MD}, \mathit{PM}, \mathit{MN}}_{4}, 
- \underbrace{\#r_1, \dots, \#r_7}_{7}, 
- \underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, 
- \underbrace{\mathit{SD}, \mathit{SS}, \mathit{G}, \mathit{A}}_{4}$$
-with the following abbreviations ($1 \le i \le 413$, $0 \le j \le 7$):
-\begin{multicols}{2}
-\begin{itemize}
-	\item[$o_i$:]  the $i$-th origin data point
-	\item[$d_i$:]  the $i$-th destination data point
-	\item[$\mathit{AM}$:] movements at time stamp AM
-	\item[$\mathit{MD}$:] movements at time stamp MD
-	\item[$\mathit{PM}$:] movements at time stamp PM
-	\item[$\mathit{MN}$:] movements at time stamp MN
-	\item[$r_j$:] the $j$-th reason
-	\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
-	\item[$\mathit{SD}$:] sum of all durations
-	\item[$\mathit{SS}$:] sum of all distances
-	\item[$\mathit{G}$:] the gender
-	\item[$\mathit{A}$:] the age
-\end{itemize}
-\end{multicols}
-
-\subsubsection{Searching for 6 Clusters}
-Applying the process with maximal 100 steps on the data gives us the following results.
-
-Like above we check first the total distribution of strata and clusters with a pie. You already see, that it is not the 
-\begin{figure}[h]
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringcluster.PNG}
-  \caption{Strata}
-  \label{fig:OrgSt}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringstrata.PNG}
-  \caption{Cluster}
-  \label{fig:OrgCl}
-\end{subfigure}
-\caption{Distribution of original data}
-\label{fig:OrgDist}
+\caption{Distribution of the clusters in between the grouped strata}
+\label{fig:Groupdist}
 \end{figure}
 
-Also the distribution in the strata groups of the clustering shows, that there is not really a connection of strata and the clustering to see.
-\includegraphics[width=0.9\textwidth]{vectorClustering.PNG}
+After having a look at the pie charts, \ref{fig:OrgDist3Cl} and so the distribution in between the variable it seems to be a better result, so we had a look at the cluster distribution in the 3 grouped stratas, \ref{fig:Groupdist}.
+This figure shows clearly, that there is no real correlation between strata
 
-So we change the maximal stepsize to 1000 and let the algorithm run again.
+\begin{figure}[!htbp]
+\centering
+\begin{subfigure}{0.9\textwidth}
+\includegraphics[width=\linewidth]{ClusterOrigRapidDistribution2038eq.PNG}
+\caption{For 6 clusters}
+\label{fig:2038_6}
+\end{subfigure}
+\begin{subfigure}{0.9\textwidth}
+\includegraphics[width=\linewidth]{ClusterOrigRapidDistribution2038eq2.PNG}
+\caption{For 3 clusters}
+\label{fig:2038_3}
+\end{subfigure}
+\caption{Distribution of the clusters in between the strata}
+\label{fig:2038_Clust}
+\end{figure}
+
+The result for different datasizes and equal distribution of the stratas does not change the result.
+The biggest equal distributed dataset has 2038 data rows for every strata and 4076 in between the grouped strata.
+In figure \ref{fig:2038_Clust} the two clustering results can be seen.
+Again we could not really see a significant correlation.
+
+\textbf{Stratified Person Data}
+
+After those not really convincing results we applied the process on the stratified person data, because those represented the movement of one person.
 
@@
   \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringcluster1000.PNG}
+  \includegraphics[width=.9\linewidth]{vectorclusteringcluster.PNG}
   \caption{Strata}
-  \label{fig:OrgSt}
+  \label{fig:VecSt}
 \end{subfigure}%
@@
   \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringstrata.PNG}
+  \includegraphics[width=.9\linewidth]{vectorclusteringstrata.PNG}
   \caption{Cluster}
-  \label{fig:OrgCl}
+  \label{fig:VecCl}
 \end{subfigure}
-\caption{Distribution of original data}
-\label{fig:OrgDist}
+\caption{Distribution of stratified person data}
+\label{fig:VecDist}
 \end{figure}
 
-So we already get the idea, that there are not 6 Clusters, but just 3. Checking the distribution still gives us no strong correlation between strata and cluster.
-
-\includegraphics[width=0.9\textwidth]{vectorClustering1000}
-
-\subsubsection{Searching for 3 Clusters}
-
-Because of the result in the last part, we checked the behavior of the clustering by just searching for 3 Cluster. The first try is with 100 steps again.
-
-\begin{figure}[h]
+\begin{figure}[!htbp]
 \centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringcluster3Cluster.PNG}
-  \caption{Strata}
-  \label{fig:OrgSt}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{vectorclusteringstrata.PNG}
-  \caption{Cluster}
-  \label{fig:OrgCl}
-\end{subfigure}
-\caption{Distribution of original data}
-\label{fig:OrgDist}
+\includegraphics[width=0.3\textwidth]{vectorclusteringcluster1000.PNG}
+\includegraphics[width=0.69\textwidth]{vectorClustering1000.PNG}
+\caption{1000 steps clustering in 6 clusters}
+\label{fig:1000vect}
 \end{figure}
 
-This comes closer by the orginial distribution combining two stratas in one cluster.
+The results for the whole data set without equalization are shown in figure \ref{fig:2038_Clust}.
+We changed the number of steps to 1000 for comparsion and the result, \ref{fig:1000vect}, let us assume, that 3 clusters better would fit.
 
-But having a look at the inbetween distribution does not show us a real correlation between cluster and strata.
+So we applied the process on the \textbf{stratified person data} and searched for \textbf{3 clusters}. We directly run the algorithm 1000 steps.
 
-\includegraphics[width=0.9\textwidth]{vectorClustering3Cluster.PNG}
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.69\textwidth]{vectorClustering31000.PNG}
+\caption{1000 steps clustering in 3 clusters}
+\label{fig:1000vect3}
+\end{figure}
 
-We wanted to check what happens when giving the process 1000 steps maximal.
+Figure \ref{fig:1000vect3} shows clearly, that again no correlation can be found. Furthermore we applied this for the different datasets we generated, but the result was always similar.

### 2018-07-11 dokumentation/DecisionTree.tex
Mein Teil fertig ausser PCA, da das nicht mehr passt.

@@
 	\item[Retrieve] includes the dataset
-	\item[Generate Attribute] changes the Numerical Attribute strata to four binomial ones
+	\item[Select Attributes] makes it possible to or have a look at grouped strata or normal strata
 	\item[Set Role] gives strata the label role, so that the decision tree has those as leafs
+	\item[Multiply] clones the data set 
 	\item[Decision Tree] creates the decision tree
+	\item[Apply Model] is used creates the labeled data set for the \textbf{Performance} step
+	\item[Performance] gives the performance result of the created model
 \end{description}
@@
 In the first step we applied the process on the whole data set and the resulting tree was just the leaf "strata 2".
-So we tried it with different other data sets and the best result we got was for \textbf{stratified person data} equally distributed 3 stratas and just 200 in every aggerated strata group.
+So we tried it with different other data sets and the best result we got was for \textbf{stratified person data} equally distribute and just 200 in every strata group.
 
@@
 \centering
-\begin{subfigure}{0.7\textwidth}
-\includegraphics
-{}
+\begin{subfigure}{0.9\textwidth}
+\includegraphics[width = \linewidth]{Dec200eqPrec.PNG}
+\caption{200 in every strata}
+\label{fig:decvec200}
+\end{subfigure}
+\begin{subfigure}{0.9\textwidth}
+\includegraphics[width= \linewidth]{decvec585pre.PNG}
+\caption{585 in every strata}
+\label{fig:DecVec585}
+\end{subfigure}
+\caption{Performance for stratified person data}
+\label{fig:DecVec}
+\end{figure}
+
+In figure \ref{fig:DecVec} the best and worst outcome can be seen for 6 clusters.
+For all othe configurations the outcome was similar, so just small data sets had a good result and big data sets had no really good result for our question, such that we still had no explicit result with which we could work.

### 2018-07-11 dokumentation/submission.tex
miriam ist dumm

@@
 	%
-	\authorrunning{F. Author et al.} %TODO
+	\authorrunning{T. Bergerbusch et al.} %TODO
 	% First names are abbreviated in the running head.
@@
 	Our goal is to ascertain if there is a correlation between the movements and the strata.
+
+	\section{Theoretical backgrounds}
+
+	\subsection{Clustering the data}
+
+	In the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} is a process of grouping data objects from a dataset into multiple groups or clusters.
+The essential criterion for the quality of the clustering is a certain \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.
+
+	In the scope of this work, we decided to use well-known partitioning methods such as k-means or k-medoids.
+In general, given $n$ data objects, partitioning methods distribute the data object into $k$ clusters with $k<n$, using a distance measure to evaluate the repsective similarity.
+Partitioning methods form exclusive clusters by ensuring that each cluster contains at least one object.
+Note that the number $k$ of clusters has to be chosen manually prior to the partitioning process.
+
+	\subsubsection{k-Means}
+
+	is a \textit{centroid based technique}, which means that each cluster is represented by a data point that also is the center of the cluster.
+A distance measure is then used to assigns every remaining data object from the data set to the best fitting cluster according to it's similarity to the center of this cluster and it's dissimilarity to the centers of any other cluster.
+
+	The data objects within a dataset are considered to reside in a euclidean space.
+Thus, the euclidean distance is used to calculate a score for the similarity of two data points.
+When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared erros between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
+This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows:
+	\begin{equation}
+		E = \sum_{i=1}^{k} \sum_{p \in C_i} dist(p, c_i)^2,
+	\end{equation}
+	Given $k$, the first step of k-Means is to select $k$ random objects in the data set as initial representatives for the cluster centers.
+After that, each remaining data object is assigned to the best fitting cluster in terms of similarity, based on a similarity score that is determined by the euclidean distance between the data point and every cluster center.
+The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation.
+In order to achieve this, each cluster centroid is reassigned as the mean of all objects within that cluster.
+By considering the updated centroids, every data point is redistributed to the now best fitting cluster.
+This iterative process will continue, until no better clustering can be found, which means that the latest clustering equals the previous distribution.
+
+	\subsection{Gower Distance}\label{gower}
+
+	When clustering data with algorithms such as \textit{k-means}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm.
+These distance measures come with constraints though, since they are only defined for numerical variables.
+In the scope of this work, we are facing data points with mixed variable types, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points.
+The Gower distance measure distinguishes between three types of variables:\\
+	\\ \textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match \cite{gower1971general}.
+In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}.
+Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values.
+Hence, we consider all variables, that are not explicitly numerical, as categorical variables.\\
+	\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.\\	
+	\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.
+
+	\subsubsection{Calculating the distances.}
+
+	Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
+	\begin{equation}\label{eq1}
+	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
+	\end{equation} 
+	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at position $k$ in the data points $x$ and $y$.
+Note, that the definition of the score dependends on the type of the variable, as defined below.
+In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, when the two variables are comparable and to 0 when not.
+For example, variables are not comparable, if values are undefined in the data points or the variable types do not match at index $k$, among other reasons.
+Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$, the number of considered variables in each data point, for all $x, y$ in the data set.
+Thus, the similarity coefficient in (\ref{eq1}) can be interpreted as the average value of all similarity scores.
+	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
+	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
+As pointed above, 0 values are not considered a match and even further, not considered to be comparable.
+Hence, the values result as in the table
+		\begin{center}
+			\begin{tabular}{l | c c c r}
+				i & 1 & 1 & 0 & 0 \\
+				j & 1 & 0 & 1 & 0 \\
+				\hline
+				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
+				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
+			\end{tabular}
+		\end{center}	
+	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$ and 0, if they differ.\\
+	\\ \textbf{Numerical:} for numerical variables, the similarity score is calculated by
+		\begin{equation*}
+		s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
+		\end{equation*}
+		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
+This can be a global range of acceptable values for variable $k$ or chosen on the basis of the dataset.
 
@@
 	%TODO: distance measures
-	TODO (1 page)
+	%TODO (1 page)
 
@@
 	%TODO: clustering
-	TODO (5 pages)	
-	%	\input{MiriamClustering}
+	%TODO (5 pages)	
+	{}\input{MiriamClustering}
+
+	\input{DecisionTree}
+	%\subsection{Neural Net}
+	%TODO (5 pages)	
 	\subsection{Neural Net} \label{subsec: neural net}
@@
 
+		\bibitem{gower1971general}
+		Gower, John C,
+		\emph{A general coefficient of similarity and some of its properties},
+		Biometrics, pp. 857-871,
+		1971.
+
+		\bibitem{data_mining}
+		Han, Jiawei and Pei, Jian and Kamber, Micheline,
+		\emph{Data mining: concepts and techniques},
+		Elsevier, pp. 444-454,
+		2011.
 	\end{thebibliography}

### 2018-07-11 dokumentation/submission.tex
fertig
added small introduction to neural nets

@@
 	%TODO (5 pages)	
-	{}\input{MiriamClustering}
+	\input{MiriamClustering}
 
@@
 	\subsection{Neural Net} \label{subsec: neural net}
-	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
+	In this section we want to further improve the accuracy of predicting the strata of one person by using neural networks.
+Like in the previous section, when we used decision trees, we need ground truth, in this case the strata, to be able to train the neural networks.
+For all the neural net computations we considered person vector data sets of different sizes (c.f.
+\Cref{subsec: person vector data}).\\
 
@@
 	\subsection{People living over/under standards}
-	%TODO
+	There are a lot of people spending more or less money than they actually have.
+So for example there are people not earning a lot of money, but still having a car.
+Or people earning a lot, but spend it just on holidays or never for bad times.
 

### 2018-07-12 dokumentation/MiriamClustering.tex
Mein Teil fertig ausser PCA, da das nicht mehr passt.
fertig
Eingebauter anmerkungen
In Introduction und abstract

@@
 %Miriam Clustering
+\setlength{\parindent}{0em}
 The first question was: is it possible without knowing the social classes to reproduce them based on the movement data.
@@
 	\item[Retrieve] gives the data into the process. 
-  \item[Generate ID] creates an ID such that we can make the comparsion step at the end through joining the sets
+  \item[Generate ID] creates an ID such that we can make the comparison step at the end through joining the sets
   \item[Multiply] creates two identical data sets
@@
 \ref{fig:OrgDist}. 
-\begin{figure}[!htbp]
+\begin{figure}[H]
 \centering
@@
 
-We asked ourselfs, if 6 cluster is not too fine, so we searched for \textbf{3 clusters} in the next step. The idea is to combine two stratas in 1, such that we just have 3 stratas left.
-\begin{figure}[!htbp]
+We asked ourselves, if 6 cluster is not too fine, so we searched for \textbf{3 clusters} in the next step. The idea is to combine two stratas in 1, such that we just have 3 stratas left.
+\begin{figure}[H]
 \centering
@@
 Figure \ref{fig:1000vect3} shows clearly, that again no correlation can be found. Furthermore we applied this for the different datasets we generated, but the result was always similar.
+
+%PCA??

### 2018-07-12 dokumentation/submission.tex
Eingebauter anmerkungen
In Introduction und abstract

@@
 	%
-	\authorrunning{T. Bergerbusch et al.} %TODO
+	\authorrunning{T. Bergerbusch et al.}
 	% First names are abbreviated in the running head.
@@
 	\begin{abstract} %TODO
-		We analyze movements in the urban environment of the columbian city Medellín.
+		We analyze movements in the urban environment of the Columbian city Medellín.
 Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
-Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
+		% Marías sentence
+		The objective of the paper is twofold: try to find groups of movement patterns and see whether they correspond to socio-economic stratus, and then, since we have the economic stratus of each person in the data set, we will apply supervised learning to classify patterns into stratus.
+%		Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
 Due to bad results we introduce our preprocessing steps and switch to supervised learning.
@@
 	\section{Introduction} \label{sec: introduction}
-	%	TODO (1 page)
-	% could at more in depth 
+	% could add more in depth 
 	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the data collected?
@@
 	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
-There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín.
+There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín, which is the second largest Colombian town with an estimated population of 2.5 million as of 2017 \cite{population_number} .
 All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
 	The data entries consist of data about the movement, like endpoints or length, and also some meta parameters like gender, age or the so-called strata of the person.
-The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.\\
-	Our goal is to ascertain if there is a correlation between the movements and the strata.
+The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.
+Those costs are defined by Colombians laws, which classifies households to regulate the access to public utility servies, having as a result six socio-economic strata \cite{rich_do_not_rise_early}\\
+	Our goal is to ascertain if there is a correlation between the movements and the strata, in order to be able to predict the strata based on the movements.
+This is done by clustering using the k-means algorithm, considering different distance measures, with optionally including the principal component analysis (PCA).
+In order to figure out the most important aspects of the data, a decision tree and neural net gets also trained and tested.
 
@@
 	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person %TODO: ref zu code basics
-	, and all movements from that person.\\
+This class stores the ID, the parameters defining a person (c.f.
+\Cref{sec: proprocessing}), and all movements from that person.\\
 	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:
@@
 		2011.
+
+		\bibitem{population_number}
+		Estimates and projections of the total national, departmental and municipal population by area 1985-2020 (XLS). NADS. Retrieved 1 September 2014.
 	\end{thebibliography}

### 2018-07-12 dokumentation/DecisionTree.tex
fertig
OVERHAUL

--- 2018-07-11 dokumentation/DecisionTree.tex
@@
-\subsection{Decision Tree} \label{subsec: decisiontree}
-Decision trees are a good manner to figure out, which parts of the data set have the most influence on the decision. We used again \textbf{RapidMiner} for building trees based on different data sets.
+\setlength{\parindent}{0em}
 
-\begin{figure}[!htbp]
+\subsection{Supervised learning}
+
+\textbf{Decision Trees} \label{subsec: decisiontree}
+%Decision trees are a good manner to figure out, which parts of the data set have the most influence on the decision.
+Therefore labeled data is needed and we have given the strata.
+We used again \textbf{RapidMiner} for building trees based on different data sets.
+%
+%\begin{figure}[!htbp]
+%\centering
+%\includegraphics[width = 0.9\textwidth]{DecisionTreeRapidModel.PNG}
+%\caption{Process for decision trees in RapidMiner}
+%\label{fig: RapDec}
+%\end{figure}
+%
+%RapidMiner does the following steps, to see in figure \ref{fig: RapDec}:
+%\begin{description}
+%	\item[Retrieve] includes the dataset
+%	\item[Select Attributes] makes it possible to or have a look at grouped strata or normal strata
+%	\item[Set Role] gives strata the label role, so that the decision tree has those as leafs
+%	\item[Multiply] clones the data set 
+%	\item[Decision Tree] creates the decision tree
+%	\item[Apply Model] is used creates the labeled data set for the \textbf{Performance} step
+%	\item[Performance] gives the performance result of the created model
+%\end{description}
+%
+%Furthermore we choose information gain as splitting criterium (minimal gain 0.1) and a confidence of 0.25. Other configuration does not show different results.
+
+In the first step we apply the process on the \textit{original data set} and the resulting tree is just the leaf `strata 2'.
+So we try it with different other data sets and the best result we get is for \textit{stratified person data} equally distributed and just 200 data entries in every strata.
+
+\begin{figure}[H]
 \centering
-\includegraphics[width = 0.9\textwidth]{DecisionTreeRapidModel.PNG}
-\caption{Process for decision trees in RapidMiner}
-\label{fig: RapDec}
+\begin{subfigure}{0.9\textwidth}
+\includegraphics[width = \linewidth]{Dec200eqPrec.PNG}
+\caption{200 in every strata}
+\label{fig:decvec200}
+\end{subfigure}
+\begin{subfigure}{0.9\textwidth}
+\includegraphics[width= \linewidth]{decvec585pre.PNG}
+\caption{585 in every strata}
+\label{fig:DecVec585}
+\end{subfigure}
+\caption{Performance for stratified person data}
+\label{fig:DecVec}
 \end{figure}
 
-RapidMiner does the following steps, to see in figure \ref{fig: RapDec}:
-\begin{description}
-	\item[Retrieve] includes the dataset
-	\item[Generate Attribute] changes the Numerical Attribute strata to four binomial ones
-	\item[Set Role] gives strata the label role, so that the decision tree has those as leafs
-	\item[Decision Tree] creates the decision tree
-\end{description}
-
-Furthermore we choose information gain as splitting criterium (minimal gain 0.1) and a confidence of 0.25. Other configuration does not show different results.
-
-In the first step we applied the process on the whole data set and the resulting tree was just the leaf "strata 2".
-So we tried it with different other data sets and the best result we got was for \textbf{stratified person data} equally distributed 3 stratas and just 200 in every aggerated strata group.
-
-\begin{figure}[!htbp]
-\centering
-\begin{subfigure}{0.7\textwidth}
-\includegraphics
-{}
+In \Cref{fig:DecVec} the best and worst outcome can be seen for 6 clusters.
+For all other configurations the outcome is similar, so just small data sets have an acceptable result and larger datasets have no sufficient results for our question.\\

### 2018-07-12 dokumentation/MiriamClustering.tex
OVERHAUL

--- 2018-07-11 dokumentation/MiriamClustering.tex
@@
 %Miriam Clustering
-The first question was: is it possible without knowing the social classes to reproduce them based on the movement data.
-Therefore we wanted to look for clusters and compare those with the strata.
-Also we had a look if the possibly found clusters have special properties.
+\setlength{\parindent}{0em}
+We look for clusters and compare those with the strata. Also we try to observe, if the possibly found clusters have special properties. % Miriam ist doof #truth
 
-\subsubsection{Clustering with RapidMiner}
+%Timo hat fineliner
+%oder doch eilaina
+In the first step we try to cluster the \textit{Original Data} in \textit{6 Cluster}. Therefore we retrieve the original data set in RapidMiner and choose k as 6.
 
-RapidMiner has different Modules for Clustering already implemented. We decided to concentrate on the k-means clustering algorithm.
 \begin{figure}[!htbp]
+\vspace*{-1.5em}
 \centering
-\includegraphics[width=0.9\textwidth]{ClusteringRapid.PNG}
-\caption{Process of k-means clustering}
-\label{fig: kclust}
-\end{figure}
-
-The Process, figure \ref{fig: kclust}, contains the following steps:
-\begin{description}
-	\item[Retrieve] gives the data into the process. 
-  \item[Generate ID] creates an ID such that we can make the comparsion step at the end through joining the sets
-  \item[Multiply] creates two identical data sets
-  \item[Select Attrbiutes] thoughs away the strata before the clustering step, everything behalve cluster and id after the clustering and just keeps id and strata for the join step
-	\item[Clustering] runs the k-means clustering algorithm. The number of Clusters has to be fixed.
-	\item[Join] For comparing the clustering result and the strata we join the two filtered data sets by the id
-\end{description}
-
-In the clustering block we can chose between different distance measures and maximal step numbers.
-We decided to concentrate on almost everywhere basic configurations and chose the squared euclidean distance in the mixed version.
-
-In the first step we tried to cluster the \textbf{Original Data} in \textbf{6 Cluster}. Therefore we retrieved the original data set in RapidMiner and chose k as 6.
-
-\ref{fig:OrgDist}. 
-\begin{figure}[!htbp]
-\centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.35\textwidth}
   \centering
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.35\textwidth}
   \centering
@@
 \end{figure}
+\vspace*{-1.5em}
 
-In figure \ref{fig:OrgDist} is the result to see of the first try.
-Figure \ref{fig:OrgSt} shows the strata distribution as pie chart and \ref{fig:OrgCl} the resulted cluster distribution.
+In \Cref{fig:OrgDist} is the result to see of the first try.
+\Cref{fig:OrgSt} shows the strata distribution as pie chart and \Cref{fig:OrgCl} the resulted cluster distribution.
 It can already been seen, that the distributions are not similar.
-In the next step we tried it with more steps, but the result was not looking better.
+For more steps the outcome is analogous.
 
-We asked ourselfs, if 6 cluster is not too fine, so we searched for \textbf{3 clusters} in the next step. The idea is to combine two stratas in 1, such that we just have 3 stratas left.
-\begin{figure}[!htbp]
+The next idea is: are 6 cluster too fine? Based on the grouped strata we look at 3 clusters.
+
+\begin{figure}[H]
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.4\textwidth}
   \centering
@@
   \caption{Strata}
-  \label{fig:OrgSt}
+  \label{fig:OrgSt3}
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.35\textwidth}
   \centering
@@
   \caption{Cluster}
-  \label{fig:OrgCl}
+  \label{fig:OrgCl3}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{ClusterOrigRapidDistribution2Cluster.PNG}
+\includegraphics[width=0.5\textwidth]{ClusterOrigRapidDistribution2Cluster.PNG}
 \caption{Distribution of the clusters in between the grouped strata}
@@
 \end{figure}
+%#ArschFuerNachbarn
+%#BloederKommentar
+%#0:15
 
-After having a look at the pie charts, \ref{fig:OrgDist3Cl} and so the distribution in between the variable it seems to be a better result, so we had a look at the cluster distribution in the 3 grouped stratas, \ref{fig:Groupdist}.
-This figure shows clearly, that there is no real correlation between strata
+\Cref{fig:OrgDist3Cl} looks promising, so we had a look at the cluster distribution in the 3 grouped stratas, visualized \Cref{fig:Groupdist}.
+It can be seen, that all clusters appear in all grouped strata.
 
 \begin{figure}[!htbp]
+\vspace{-3em}
 \centering
-\begin{subfigure}{0.9\textwidth}
+\begin{subfigure}{0.45\textwidth}
 \includegraphics[width=\linewidth]{ClusterOrigRapidDistribution2038eq.PNG}
@@
 \end{subfigure}
-\begin{subfigure}{0.9\textwidth}
+\begin{subfigure}{0.45\textwidth}
 \includegraphics[width=\linewidth]{ClusterOrigRapidDistribution2038eq2.PNG}
@@
 \end{subfigure}
-\caption{Distribution of the clusters in between the strata}
+\caption{Distribution of the clusters in between the strata dataset size 2038}
 \label{fig:2038_Clust}
+\vspace{-3em}
 \end{figure}
-
-The result for different datasizes and equal distribution of the stratas does not change the result.
-The biggest equal distributed dataset has 2038 data rows for every strata and 4076 in between the grouped strata.
-In figure \ref{fig:2038_Clust} the two clustering results can be seen.
-Again we could not really see a significant correlation.
-
-\textbf{Stratified Person Data}
-
-After those not really convincing results we applied the process on the stratified person data, because those represented the movement of one person.
-
-\begin{figure}[h]
+\vspace{-3em}
+\begin{figure}[H]
+	\vspace{-3em}
 	\centering
-\begin{subfigure}{.5\textwidth}
+	\begin{subfigure}{.43\textwidth}
 		\centering
@@
 	\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+	\begin{subfigure}{.4\textwidth}
 		\centering
@@
 
-\begin{figure}[!htbp]
+\begin{figure}[H]
 	\centering
-\includegraphics[width=0.3\textwidth]{vectorclusteringcluster1000.PNG}
-\includegraphics[width=0.69\textwidth]{vectorClustering1000.PNG}
-\caption{1000 steps clustering in 6 clusters}
+	\includegraphics[width=0.4\textwidth]{vectorclusteringcluster1000.PNG}
+	\includegraphics[width=0.49\textwidth]{vectorClustering1000.PNG}
+	\caption{1000 steps clustering in 6 clusters on maximal stratified person dataset}
 	\label{fig:1000vect}
@@
 
-The results for the whole data set without equalization are shown in figure \ref{fig:2038_Clust}.
-We changed the number of steps to 1000 for comparsion and the result, \ref{fig:1000vect}, let us assume, that 3 clusters better would fit.
+Different dataset sizes and equal distribution of the stratas do not effect the result.
+In  \Cref{fig:2038_Clust} two clustering results can be seen.
+Again we could not observe a significant correlation.
 
-So we applied the process on the \textbf{stratified person data} and searched for \textbf{3 clusters}. We directly run the algorithm 1000 steps.
+After those not convincing results we applied the process to the \textit{stratified person data}.
+
+The results for the whole data set without equalization are shown in \Cref{fig:2038_Clust}.
+We change the number of steps to 1000 for comparison and the result, illustrated in \Cref{fig:1000vect}, lets us assume, that 3 clusters would fit better.
 
@@
 \includegraphics[width=0.69\textwidth]{vectorClustering31000.PNG}
-\caption{1000 steps clustering in 3 clusters}
+\caption{1000 steps clustering in 3 clusters on maximal stratified person dataset}
 \label{fig:1000vect3}
@@
 
-Figure \ref{fig:1000vect3} shows clearly, that again no correlation can be found. Furthermore we applied this for the different datasets we generated, but the result was always similar.
+\Cref{fig:1000vect3} shows clearly, that again no correlation can be found. Applying the procedure to the different datasets we generated creates similar results.
+
+%PCA?

### 2018-07-12 dokumentation/conclusion.tex
OVERHAUL

--- 2018-07-11 dokumentation/submission.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-%\usepackage{ngerman}
-\usepackage[utf8x]{inputenc}
-\usepackage{fancyvrb}
-\usepackage{courier}
-\usepackage{helvet}
-\usepackage{tikz}
-\usepackage{xcolor}
-\usepackage{pdfpages}
-\usetikzlibrary{calc}
-\usepackage[strict]{changepage}
-\usepackage{xspace}
-\usepackage{hyperref}
-\usetikzlibrary{shapes.geometric}
-\usetikzlibrary{arrows}
-\usetikzlibrary{positioning}
-\usetikzlibrary{arrows.meta, automata, shapes, matrix,positioning}
-\usepackage{amssymb}
-\usepackage{pifont}% http://ctan.org/pkg/pifont
-\usepackage{subcaption} 
-\usepackage{float}
-\usepackage{fixfoot}
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage{multicol}
-\usepackage{amsmath}
-\usepackage{cleveref}
-\usepackage{listings}
-
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\begin{document}
-
-	%
-	%	\title{Clustering Analysis of Mobility Data}
-	\title{Are you moving predictably?}
-	%
-	%\titlerunning{Abbreviated paper title}
-	% If the paper title is too long for the running head, you can set
-	% an abbreviated paper title here
-	%
-	%TODO: Order by Lastname?
-	\author{Miriam Wagner\and
-		Martin Breuer\and
-		Moritz Werthebach\and
-		Timo Bergerbusch\and
-		Walter Schikowski}
-	%
-	\authorrunning{T. Bergerbusch et al.} %TODO
-	% First names are abbreviated in the running head.
-	% If there are more than two authors, 'et al.' is used.
-	%
-	\institute{RWTH Aachen, Templergraben 55, 52062 Aachen, Germany}
-	%
-	\maketitle              % typeset the header of the contribution
-	%
-	\begin{abstract} %TODO
-		We analyze movements in the urban environment of the columbian city Medellín.
-Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
-Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
-Due to bad results we introduce our preprocessing steps and switch to supervised learning.
-Decision trees and neural networks did neither match our performance expectations which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
-
-		\keywords{Data Mining \and Clustering \and Rapidminer \and Cluster \and Neural Nets} %TODO: maybe more
-	\end{abstract}
-	%
-	%
-	%
-	\section{Introduction} \label{sec: introduction}
-	%	TODO (1 page)
-	% could at more in depth 
-	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the data collected?
-In particular is it possible to determine the wealth of a person, only given movements of a single day?
-	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
-There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín.
-All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
-	The data entries consist of data about the movement, like endpoints or length, and also some meta parameters like gender, age or the so-called strata of the person.
-The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.\\
-	Our goal is to ascertain if there is a correlation between the movements and the strata.
-
-	\section{Theoretical backgrounds}
-
-	\subsection{Clustering the data}
-
-	In the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} is a process of grouping data objects from a dataset into multiple groups or clusters.
-The essential criterion for the quality of the clustering is a certain \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.
-
-	In the scope of this work, we decided to use well-known partitioning methods such as k-means or k-medoids.
-In general, given $n$ data objects, partitioning methods distribute the data object into $k$ clusters with $k<n$, using a distance measure to evaluate the repsective similarity.
-Partitioning methods form exclusive clusters by ensuring that each cluster contains at least one object.
-Note that the number $k$ of clusters has to be chosen manually prior to the partitioning process.
-
-	\subsubsection{k-Means}
-
-	is a \textit{centroid based technique}, which means that each cluster is represented by a data point that also is the center of the cluster.
-A distance measure is then used to assigns every remaining data object from the data set to the best fitting cluster according to it's similarity to the center of this cluster and it's dissimilarity to the centers of any other cluster.
-
-	The data objects within a dataset are considered to reside in a euclidean space.
-Thus, the euclidean distance is used to calculate a score for the similarity of two data points.
-When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared erros between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
-This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows:
-	\begin{equation}
-		E = \sum_{i=1}^{k} \sum_{p \in C_i} dist(p, c_i)^2,
-	\end{equation}
-	Given $k$, the first step of k-Means is to select $k$ random objects in the data set as initial representatives for the cluster centers.
-After that, each remaining data object is assigned to the best fitting cluster in terms of similarity, based on a similarity score that is determined by the euclidean distance between the data point and every cluster center.
-The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation.
-In order to achieve this, each cluster centroid is reassigned as the mean of all objects within that cluster.
-By considering the updated centroids, every data point is redistributed to the now best fitting cluster.
-This iterative process will continue, until no better clustering can be found, which means that the latest clustering equals the previous distribution.
-
-	\subsection{Gower Distance}\label{gower}
-
-	When clustering data with algorithms such as \textit{k-means}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm.
-These distance measures come with constraints though, since they are only defined for numerical variables.
-In the scope of this work, we are facing data points with mixed variable types, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points.
-The Gower distance measure distinguishes between three types of variables:\\
-	\\ \textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match \cite{gower1971general}.
-In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}.
-Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values.
-Hence, we consider all variables, that are not explicitly numerical, as categorical variables.\\
-	\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.\\	
-	\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.
-
-	\subsubsection{Calculating the distances.}
-
-	Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
-	\begin{equation}\label{eq1}
-	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
-	\end{equation} 
-	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at position $k$ in the data points $x$ and $y$.
-Note, that the definition of the score dependends on the type of the variable, as defined below.
-In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, when the two variables are comparable and to 0 when not.
-For example, variables are not comparable, if values are undefined in the data points or the variable types do not match at index $k$, among other reasons.
-Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$, the number of considered variables in each data point, for all $x, y$ in the data set.
-Thus, the similarity coefficient in (\ref{eq1}) can be interpreted as the average value of all similarity scores.
-	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
-	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
-As pointed above, 0 values are not considered a match and even further, not considered to be comparable.
-Hence, the values result as in the table
-		\begin{center}
-			\begin{tabular}{l | c c c r}
-				i & 1 & 1 & 0 & 0 \\
-				j & 1 & 0 & 1 & 0 \\
-				\hline
-				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
-				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
-			\end{tabular}
-		\end{center}	
-	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$ and 0, if they differ.\\
-	\\ \textbf{Numerical:} for numerical variables, the similarity score is calculated by
-		\begin{equation*}
-		s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
-		\end{equation*}
-		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
-This can be a global range of acceptable values for variable $k$ or chosen on the basis of the dataset.
-
-	\section{Preprocessing}\label{sec: proprocessing}
-	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
-	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
-	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets considered by our clustering and neural net approaches.\\
-	We observe the following distribution over the whole dataset:\\
-	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
-			abs   & 6963 & 52265 & 49404 & 8772 & 5536 & 2038 &  124978  \\
-			\%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
-		\end{tabular}
-		\label{table: distribution normal}
-	\end{figure}
-	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
-	Furthermore we have to make sure that  two different data points, which belong to the very same person, are assigned to the same cluster.
-To do so we compute the value \texttt{ID} which identifies each person and can be used to combine movements that are considered to be from the same person, i.
-e.
-two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
-This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.
-	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6  & $\Sigma$ \\ \hline
-			abs   & 3153 & 23367 & 21418 & 3497 & 2083 & 595 &  54113   \\
-			\%   & 5.83 & 43.18 & 39.58 & 6.46 & 3.85 & 1.1 &   100
-		\end{tabular}
-	\end{figure}
-	In Section \Cref{subsec: person vector data} we introduce vectors representing single persons.
-Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.
-
-	\subsection{Stratified Person Data}\label{subsec: person vector data}
-	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person %TODO: ref zu code basics
-	, and all movements from that person.\\
-	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:
-	\begin{align*}
-	\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,
-	\underbrace{\mathit{AM}, \mathit{MD}, \mathit{PM}, \mathit{MN}}_{4}, 
-	\underbrace{\#r_1, \dots, \#r_7}_{7}, \\
-	\underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, \underbrace{\mathit{S_{Dest}}, \mathit{S_{Dist}}, \mathit{G}, \mathit{A} ,\mathit{strata}, \mathit{strataGrouped}}_{6}
-	\end{align*}
-	with the following abbreviations ($1 \le i \le 413$, $1 \le j \le 7$):
-	\begin{multicols}{2}
-		\begin{itemize}
-			\setlength{\itemindent}{.4cm}
-			\item[$o_i$:]  the $i$-th origin data point
-			\item[$d_i$:]  the $i$-th destination data point
-			\item[$\mathit{AM}$:] movements at time stamp AM
-			\item[$\mathit{MD}$:] movements at time stamp MD
-			\item[$\mathit{PM}$:] movements at time stamp PM
-			\item[$\mathit{MN}$:] movements at time stamp MN
-			\item[$r_j$:] the $j$-th reason
-			\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
-			\item[$\mathit{S_{Dest}}$:] sum of all durations
-			\item[$\mathit{S_{Dist}}$:] sum of all distances
-			\item[$\mathit{G}$:] the gender
-			\item[$\mathit{A}$:] the age
-			\item[$strata$:] the strata (used for comparison)
-			\item[$strataGrouped$:] the aggregated stratas
-		\end{itemize}
-	\end{multicols}
-	\section{Predicting}
-
-	\subsection{Distance Measures} \label{subsec: distance measures}
-	%TODO: distance measures
-	%TODO (1 page)
-
-	\subsection{Classification} \label{subsec: classification}
-	%TODO: clustering
-	%TODO (5 pages)	
-	{}\input{MiriamClustering}
-
-	\input{DecisionTree}
-	%\subsection{Neural Net}
-	%TODO (5 pages)	
-	\subsection{Neural Net} \label{subsec: neural net}
-	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
-
-	We do this, because results on the normal datasets had an unacceptable performance, since only single movements and not complete paths of individuals are considered.
-An example training and performance measure is given in \Cref{fig: NN without vector} where unprocessed data is used.
-The performance is measured using 10-fold cross validation, i.
-e.
-the data is split into 10 subsets where in each iteration exactly one data set is used as test set and the other 9 as training set.
-The average value of the accuracy values lead
-	to the total accuracy value of the neural net.
-	\begin{figure}[H]
-		\centering
-		\includegraphics[scale = 0.4]{src/pic/NN_without_vector.png}
-		\caption{An example of a neural net trained without person vector data.}
-		\label{fig: NN without vector}
-	\end{figure}
-
-	In the following we consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
-	As an example of other strata aggregation we combine the stratas 1--2, 3--4 and 5--6 together and call them $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$.
-This builds a superset of the original stratas and since the stratas themselves are logically connected this task should be easier to fulfill.\\
-
-	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata which are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
-For every neural net and every set size we performed 5 independent runs and calculated the average over those accuracy values in order to have a sophisticated, comparable statements.
-	\setlength\tabcolsep{.2cm}
-	\begin{figure}[H]
-		\centering
-		\begin{tabular}{|c|c|c|c|c|c|}
-			\hline
-			&   \#    &        & \multicolumn{3}{c|}{Set size} \\
-			Name           & Neurons &   AG   &  100  &  200  &      595      \\ \hline
-			$\mathcal{N}_5$      &    5    & \xmark & 60.03 & 59.92 &     60.18     \\
-			$\mathcal{N}_5^\star$   &    5    & \cmark & 87.6  & 89.7  &     71.05     \\
-			$\mathcal{N}_{10}$    &   10    & \xmark & 75.83 & 73.54 &     69.56     \\
-			$\mathcal{N}_{10}^\star$ &   10    & \cmark & 92.93 & 93.48 &     74.58     \\		
-			%		$\mathcal{N}_{10}^\star$ &   10    & \cmark & 88.33 {\small $\pm$7.49} & 90.67{\small $\pm$2.81} & 92.14 {\small $\pm$ 2.59} \\
-			$\mathcal{N}_{20}$    &   20    & \xmark & 75.45 & 71.14 &     61.87     \\
-			$\mathcal{N}_{20}^\star$ &   20    & \cmark & 92.87 & 94.4  &     78.32     \\ \hline
-		\end{tabular}
-		\caption{The accuracy values of the neural nets. (See excel-spreadsheet)}
-		\label{tab: nn-accuracy}
-		% mittels deep-learning-vector-average
-	\end{figure}
-
-	%TODO: recompute
-	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be too small to have sufficient training.\\
-
-	\section{Observations} \label{sec: observation}
-	%	TODO (1 page)
-	%	TODO: restructure. Have like 3 parts
-	%	The greater the population, the more the borders between the stratas blur. 
-	As an overall result we get, that we can not determine the strata based on the information we have.
-	Using various datasets (c.f.
-\Cref{sec: proprocessing}), we witnessed that having equally distributed datasets lead to an overall higher accuracy.
-This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f.
-\Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
-	During clustering we observed, that using a different distance measure formula would not lead to a huge difference.
-Also, we detected that reducing the numbers of clusters leads to better results, which reduces the significance of result that could be stated.\\
-	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
-	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
-Therefore we see that the more data we have the higher the variance within the stratas themselves is.
-We cannot draw some kind of lines to distinguish between different entries.
-Datapoints, which were outliers in the smaller sets, are now not considered to be outliers, because of large numbers having the same characteristics.
-So the lines, we were able to draw for smaller datasets, are blurring and create some kind of transition phase.\\ %TODO: letzter satz doof
-
 	\section{Discussion} \label{sec: discussion}
-	%TODO: discussion
-%	TODO (2 page)
 %	\begin{itemize}
@@
 \Cref{sec: introduction}) is not mentioned.
-Therefore we can not infer that it is representative.
-The people asked to enter their movement could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
+Therefore we cannot infer, that it is representative.
+The people, asked to enter their movement, could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
 \\
-	Also the day itself is not mentioned, so we don't have information if it was even a working day or a weekend. This influences the behavior drastically.
+	Also the day itself is not mentioned, so we do not have information, if it was even a working day or a weekend. This influences the behavior drastically.
 
 	\subsection{People living over/under standards}
-	%TODO
+	There are a lot of people spending more or less money than they actually have.
+So for example there are people not earning a lot of money, but still having a car.
+Or people earning a lot, but spend it just on holidays or save it for bad times.
+
+%	In Liebe
+%	deine Miriam
 
@@
 	Obviously, people are individual in their way of life.
-So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply don't like driving.
-On the other end of the spectrum, some people, who have little money still own and drive a car daily in order to go to work, since they can't afford an other apartment.\\
-	One can think of multitude scenarios, of people behaving different, cause by their individuality.
+So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply do not like driving.
+On the other end of the spectrum, some people, who have little money, still own and drive a car daily in order to go to work, since they cannot afford an other apartment.\\
+	One can think of multitude scenarios of people behaving different caused by their individuality.
 
@@
 		\item[2. Group]
-		People, who are rich just by birth, which don't have the necessity to work or move at all. They could possibly stay home and don't have to leave at all.
+		People, who are rich just by birth, which do not have the necessity to work or move at all. They could possibly stay home and do not have to leave at all.
 	\end{enumerate} 
@@
 Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work.
-Also there are people, who don't move at all, because of the same reason.
+Also there are people, who do not move at all, because of the same reason.
 
-	So we can see that based on the information we have it is very unlikely to have a precise correlation between the given data and the strata.
-However given more information about the previously mentioned aspects and more entries in general there could be some kind correlation to be found.\\
+	So we can see that based on the information we have, it is very unlikely to have a clear correlation between the given data and the strata.
+However, given more information about the previously mentioned aspects and more entries in general, there could be some kind of correlation to be found.\\
 	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
-Also one could use different network visualization tools in order to infer their own patterns.
-An example would be to take the origin (destination) sectors as nodes, an edge if there is a movement, directed or undirected, and a different thickness or color gradient based on the number of this edge being taken.\\
+Also one could use different network visualization tools in order to infer patterns.
+An example would be to take the origin (destination) sectors as nodes, an edge, if there is a movement between, directed or undirected, and a different thickness or color gradient, based on the number of this edge being taken.\\
 
-	Overall we can conclude that with the data we got we are not able to find a correlation.
-
-	\begin{thebibliography}{8}
-		\bibitem{rich_do_not_rise_early}
-		Lotero, Laura, et al. "Rich do not rise early: spatio-temporal patterns in the mobility networks of different socio-economic classes." Royal Society open science 3.10 (2016): 150654.
-
-		\bibitem{strata_def}
-		Hudson, Rex A. Colombia: A country study. Government Printing Office, 2010.
-
-		\bibitem{gower1971general}
-		Gower, John C,
-		\emph{A general coefficient of similarity and some of its properties},
-		Biometrics, pp. 857-871,
-		1971.
-
-		\bibitem{data_mining}
-		Han, Jiawei and Pei, Jian and Kamber, Micheline,
-		\emph{Data mining: concepts and techniques},
-		Elsevier, pp. 444-454,
-		2011.
-	\end{thebibliography}
-\end{document}
+	Overall we can conclude that with the data we got, we are not able to find a correlation.

### 2018-07-12 dokumentation/introduction.tex
OVERHAUL

\section{Introduction} \label{sec: introduction}
	% could add more in depth 
	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the collected data? In particular is it possible to determine the wealth of a person, only given movements of a single day?
	For this we considered the dataset stated in \cite{rich_do_not_rise_early}. There we have a set of 124979 rows of movement data of various persons from the Columbian town Medellín, which is the second largest Colombian town with an estimated population of 2.5 million as of 2017 \cite{population_number} . All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
	The data entries consist of data about the movement, like endpoints, length, and also some meta parameters like gender, age or the so-called strata of the person. The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs. Those costs are defined by Colombians laws, which classifies households to regulate the access to public utility services, having as a result six socio-economic strata \cite{rich_do_not_rise_early}.\\
	Our goal is to ascertain, if there is a correlation between the movements and the strata, in order to be able to predict the strata based on the movements. This is done in two steps:\\
	First, by clustering using the k-means algorithm, considering different distance measures, with optionally including the principal component analysis (PCA).\\
	And second, a decision tree and neural net were trained and tested.

### 2018-07-12 dokumentation/neuralnet.tex
OVERHAUL

--- 2018-07-11 dokumentation/submission.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-%\usepackage{ngerman}
-\usepackage[utf8x]{inputenc}
-\usepackage{fancyvrb}
-\usepackage{courier}
-\usepackage{helvet}
-\usepackage{tikz}
-\usepackage{xcolor}
-\usepackage{pdfpages}
-\usetikzlibrary{calc}
-\usepackage[strict]{changepage}
-\usepackage{xspace}
-\usepackage{hyperref}
-\usetikzlibrary{shapes.geometric}
-\usetikzlibrary{arrows}
-\usetikzlibrary{positioning}
-\usetikzlibrary{arrows.meta, automata, shapes, matrix,positioning}
-\usepackage{amssymb}
-\usepackage{pifont}% http://ctan.org/pkg/pifont
-\usepackage{subcaption} 
-\usepackage{float}
-\usepackage{fixfoot}
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage{multicol}
-\usepackage{amsmath}
-\usepackage{cleveref}
-\usepackage{listings}
-
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\begin{document}
-
-	%
-	%	\title{Clustering Analysis of Mobility Data}
-	\title{Are you moving predictably?}
-	%
-	%\titlerunning{Abbreviated paper title}
-	% If the paper title is too long for the running head, you can set
-	% an abbreviated paper title here
-	%
-	%TODO: Order by Lastname?
-	\author{Miriam Wagner\and
-		Martin Breuer\and
-		Moritz Werthebach\and
-		Timo Bergerbusch\and
-		Walter Schikowski}
-	%
-	\authorrunning{T. Bergerbusch et al.} %TODO
-	% First names are abbreviated in the running head.
-	% If there are more than two authors, 'et al.' is used.
-	%
-	\institute{RWTH Aachen, Templergraben 55, 52062 Aachen, Germany}
-	%
-	\maketitle              % typeset the header of the contribution
-	%
-	\begin{abstract} %TODO
-		We analyze movements in the urban environment of the columbian city Medellín.
-Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
-Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
-Due to bad results we introduce our preprocessing steps and switch to supervised learning.
-Decision trees and neural networks did neither match our performance expectations which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
-
-		\keywords{Data Mining \and Clustering \and Rapidminer \and Cluster \and Neural Nets} %TODO: maybe more
-	\end{abstract}
-	%
-	%
-	%
-	\section{Introduction} \label{sec: introduction}
-	%	TODO (1 page)
-	% could at more in depth 
-	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the data collected?
-In particular is it possible to determine the wealth of a person, only given movements of a single day?
-	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
-There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín.
-All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
-	The data entries consist of data about the movement, like endpoints or length, and also some meta parameters like gender, age or the so-called strata of the person.
-The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.\\
-	Our goal is to ascertain if there is a correlation between the movements and the strata.
-
-	\section{Theoretical backgrounds}
-
-	\subsection{Clustering the data}
-
-	In the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} is a process of grouping data objects from a dataset into multiple groups or clusters.
-The essential criterion for the quality of the clustering is a certain \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.
-
-	In the scope of this work, we decided to use well-known partitioning methods such as k-means or k-medoids.
-In general, given $n$ data objects, partitioning methods distribute the data object into $k$ clusters with $k<n$, using a distance measure to evaluate the repsective similarity.
-Partitioning methods form exclusive clusters by ensuring that each cluster contains at least one object.
-Note that the number $k$ of clusters has to be chosen manually prior to the partitioning process.
-
-	\subsubsection{k-Means}
-
-	is a \textit{centroid based technique}, which means that each cluster is represented by a data point that also is the center of the cluster.
-A distance measure is then used to assigns every remaining data object from the data set to the best fitting cluster according to it's similarity to the center of this cluster and it's dissimilarity to the centers of any other cluster.
-
-	The data objects within a dataset are considered to reside in a euclidean space.
-Thus, the euclidean distance is used to calculate a score for the similarity of two data points.
-When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared erros between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
-This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows:
-	\begin{equation}
-		E = \sum_{i=1}^{k} \sum_{p \in C_i} dist(p, c_i)^2,
-	\end{equation}
-	Given $k$, the first step of k-Means is to select $k$ random objects in the data set as initial representatives for the cluster centers.
-After that, each remaining data object is assigned to the best fitting cluster in terms of similarity, based on a similarity score that is determined by the euclidean distance between the data point and every cluster center.
-The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation.
-In order to achieve this, each cluster centroid is reassigned as the mean of all objects within that cluster.
-By considering the updated centroids, every data point is redistributed to the now best fitting cluster.
-This iterative process will continue, until no better clustering can be found, which means that the latest clustering equals the previous distribution.
-
-	\subsection{Gower Distance}\label{gower}
-
-	When clustering data with algorithms such as \textit{k-means}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm.
-These distance measures come with constraints though, since they are only defined for numerical variables.
-In the scope of this work, we are facing data points with mixed variable types, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points.
-The Gower distance measure distinguishes between three types of variables:\\
-	\\ \textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match \cite{gower1971general}.
-In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}.
-Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values.
-Hence, we consider all variables, that are not explicitly numerical, as categorical variables.\\
-	\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.\\	
-	\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.
-
-	\subsubsection{Calculating the distances.}
-
-	Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
-	\begin{equation}\label{eq1}
-	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
-	\end{equation} 
-	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at position $k$ in the data points $x$ and $y$.
-Note, that the definition of the score dependends on the type of the variable, as defined below.
-In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, when the two variables are comparable and to 0 when not.
-For example, variables are not comparable, if values are undefined in the data points or the variable types do not match at index $k$, among other reasons.
-Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$, the number of considered variables in each data point, for all $x, y$ in the data set.
-Thus, the similarity coefficient in (\ref{eq1}) can be interpreted as the average value of all similarity scores.
-	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
-	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
-As pointed above, 0 values are not considered a match and even further, not considered to be comparable.
-Hence, the values result as in the table
-		\begin{center}
-			\begin{tabular}{l | c c c r}
-				i & 1 & 1 & 0 & 0 \\
-				j & 1 & 0 & 1 & 0 \\
-				\hline
-				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
-				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
-			\end{tabular}
-		\end{center}	
-	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$ and 0, if they differ.\\
-	\\ \textbf{Numerical:} for numerical variables, the similarity score is calculated by
-		\begin{equation*}
-		s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
-		\end{equation*}
-		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
-This can be a global range of acceptable values for variable $k$ or chosen on the basis of the dataset.
-
-	\section{Preprocessing}\label{sec: proprocessing}
-	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
-	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
-	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets considered by our clustering and neural net approaches.\\
-	We observe the following distribution over the whole dataset:\\
+\textbf{Neural Net} \label{subsec: neural net}\\
+	In this section we want to further improve the accuracy of predicting the strata of one person by using neural networks.
+Like in the previous section, where we used decision trees, we need ground truth, in this case the strata, to be able to train the neural networks.
+For all the neural net computations we considered person vector data sets of different sizes (c.f.
+\Cref{sec: proprocessing}).\\
+	We do this, because results on the normal datasets had an unacceptable performance, since only single movements and not complete paths of individuals are considered.
+An example training and performance measure is given in \Cref{fig: NN without vector}, where unprocessed data is used.
+The performance is measured using 10-fold cross validation, i.e.
+the data is split into 10 subsets, where in each iteration exactly one data set is used as test set and the other 9 as training set.
+The average value of the accuracy values leads to the total accuracy of the neural net.
 	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
-			abs   & 6963 & 52265 & 49404 & 8772 & 5536 & 2038 &  124978  \\
-			\%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
-		\end{tabular}
-		\label{table: distribution normal}
-	\end{figure}
-	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
-	Furthermore we have to make sure that  two different data points, which belong to the very same person, are assigned to the same cluster.
-To do so we compute the value \texttt{ID} which identifies each person and can be used to combine movements that are considered to be from the same person, i.
-e.
-two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
-This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.
-	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6  & $\Sigma$ \\ \hline
-			abs   & 3153 & 23367 & 21418 & 3497 & 2083 & 595 &  54113   \\
-			\%   & 5.83 & 43.18 & 39.58 & 6.46 & 3.85 & 1.1 &   100
-		\end{tabular}
-	\end{figure}
-	In Section \Cref{subsec: person vector data} we introduce vectors representing single persons.
-Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.
-
-	\subsection{Stratified Person Data}\label{subsec: person vector data}
-	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person %TODO: ref zu code basics
-	, and all movements from that person.\\
-	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:
-	\begin{align*}
-	\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,
-	\underbrace{\mathit{AM}, \mathit{MD}, \mathit{PM}, \mathit{MN}}_{4}, 
-	\underbrace{\#r_1, \dots, \#r_7}_{7}, \\
-	\underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, \underbrace{\mathit{S_{Dest}}, \mathit{S_{Dist}}, \mathit{G}, \mathit{A} ,\mathit{strata}, \mathit{strataGrouped}}_{6}
-	\end{align*}
-	with the following abbreviations ($1 \le i \le 413$, $1 \le j \le 7$):
-	\begin{multicols}{2}
-		\begin{itemize}
-			\setlength{\itemindent}{.4cm}
-			\item[$o_i$:]  the $i$-th origin data point
-			\item[$d_i$:]  the $i$-th destination data point
-			\item[$\mathit{AM}$:] movements at time stamp AM
-			\item[$\mathit{MD}$:] movements at time stamp MD
-			\item[$\mathit{PM}$:] movements at time stamp PM
-			\item[$\mathit{MN}$:] movements at time stamp MN
-			\item[$r_j$:] the $j$-th reason
-			\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
-			\item[$\mathit{S_{Dest}}$:] sum of all durations
-			\item[$\mathit{S_{Dist}}$:] sum of all distances
-			\item[$\mathit{G}$:] the gender
-			\item[$\mathit{A}$:] the age
-			\item[$strata$:] the strata (used for comparison)
-			\item[$strataGrouped$:] the aggregated stratas
-		\end{itemize}
-	\end{multicols}
-	\section{Predicting}
-
-	\subsection{Distance Measures} \label{subsec: distance measures}
-	%TODO: distance measures
-	%TODO (1 page)
-
-	\subsection{Classification} \label{subsec: classification}
-	%TODO: clustering
-	%TODO (5 pages)	
-	{}\input{MiriamClustering}
-
-	\input{DecisionTree}
-	%\subsection{Neural Net}
-	%TODO (5 pages)	
-	\subsection{Neural Net} \label{subsec: neural net}
-	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
-
-	We do this, because results on the normal datasets had an unacceptable performance, since only single movements and not complete paths of individuals are considered.
-An example training and performance measure is given in \Cref{fig: NN without vector} where unprocessed data is used.
-The performance is measured using 10-fold cross validation, i.
-e.
-the data is split into 10 subsets where in each iteration exactly one data set is used as test set and the other 9 as training set.
-The average value of the accuracy values lead
-	to the total accuracy value of the neural net.
-	\begin{figure}[H]
+		\vspace*{-1em}
 		\centering
@@
 	\end{figure}
+	\vspace*{-1.5em}
 
@@
 	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata which are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
-For every neural net and every set size we performed 5 independent runs and calculated the average over those accuracy values in order to have a sophisticated, comparable statements.
+For every neural net and every set size we perform 5 independent runs and calculate the average over those accuracy values in order to have a sophisticated, comparable statement.
 	\setlength\tabcolsep{.2cm}
@@
 	%TODO: recompute
-	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be too small to have sufficient training.\\
-
-	\section{Observations} \label{sec: observation}
-	%	TODO (1 page)
-	%	TODO: restructure. Have like 3 parts
-	%	The greater the population, the more the borders between the stratas blur. 
-	As an overall result we get, that we can not determine the strata based on the information we have.
-	Using various datasets (c.f.
-\Cref{sec: proprocessing}), we witnessed that having equally distributed datasets lead to an overall higher accuracy.
-This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f.
-\Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
-	During clustering we observed, that using a different distance measure formula would not lead to a huge difference.
-Also, we detected that reducing the numbers of clusters leads to better results, which reduces the significance of result that could be stated.\\
-	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
-	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
-Therefore we see that the more data we have the higher the variance within the stratas themselves is.
-We cannot draw some kind of lines to distinguish between different entries.
-Datapoints, which were outliers in the smaller sets, are now not considered to be outliers, because of large numbers having the same characteristics.
-So the lines, we were able to draw for smaller datasets, are blurring and create some kind of transition phase.\\ %TODO: letzter satz doof
-
-	\section{Discussion} \label{sec: discussion}
-	%TODO: discussion
-%	TODO (2 page)
-%	\begin{itemize}
-%		\item repr. working day
-%		\item over/under std's
-%		\item indiv. of lifestyle
-%		\item inv. behaviour in same strata
-%	\end{itemize}
-	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we though about various aspects having an impact on the movement.
-We state 4 main influencing aspects, explaining the variance throughout the data.
-
-	\subsection{Representativity of the day}
-	The day the data was collected on (c.f.
-\Cref{sec: introduction}) is not mentioned.
-Therefore we can not infer that it is representative.
-The people asked to enter their movement could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
-\\
-	Also the day itself is not mentioned, so we don't have information if it was even a working day or a weekend. This influences the behavior drastically.
-
-	\subsection{People living over/under standards}
-	%TODO
-
-	\subsection{Individuality of lifestyle}
-	Obviously, people are individual in their way of life.
-So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply don't like driving.
-On the other end of the spectrum, some people, who have little money still own and drive a car daily in order to go to work, since they can't afford an other apartment.\\
-	One can think of multitude scenarios, of people behaving different, cause by their individuality.
-
-	\subsection{Inverse behavior in a strata}
-	As an example, take a look at the strata 6, which denotes the richest people considered. Their we have inter alia two groups: 
-	\begin{enumerate}
-		\setlength{\itemindent}{1cm}
-		\item[1. Group]
-		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
-		\item[2. Group]
-		People, who are rich just by birth, which don't have the necessity to work or move at all. They could possibly stay home and don't have to leave at all.
-	\end{enumerate} 
-	Those two groups are completely opposite, but still belong to the same strata.
-Now also considering strata one, we can find the exact same movement behavior in there as well.
-Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work.
-Also there are people, who don't move at all, because of the same reason.
-
-	So we can see that based on the information we have it is very unlikely to have a precise correlation between the given data and the strata.
-However given more information about the previously mentioned aspects and more entries in general there could be some kind correlation to be found.\\
-	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
-Also one could use different network visualization tools in order to infer their own patterns.
-An example would be to take the origin (destination) sectors as nodes, an edge if there is a movement, directed or undirected, and a different thickness or color gradient based on the number of this edge being taken.\\
-
-	Overall we can conclude that with the data we got we are not able to find a correlation.
-
-	\begin{thebibliography}{8}
-		\bibitem{rich_do_not_rise_early}
-		Lotero, Laura, et al. "Rich do not rise early: spatio-temporal patterns in the mobility networks of different socio-economic classes." Royal Society open science 3.10 (2016): 150654.
-
-		\bibitem{strata_def}
-		Hudson, Rex A. Colombia: A country study. Government Printing Office, 2010.
-
-		\bibitem{gower1971general}
-		Gower, John C,
-		\emph{A general coefficient of similarity and some of its properties},
-		Biometrics, pp. 857-871,
-		1971.
-
-		\bibitem{data_mining}
-		Han, Jiawei and Pei, Jian and Kamber, Micheline,
-		\emph{Data mining: concepts and techniques},
-		Elsevier, pp. 444-454,
-		2011.
-	\end{thebibliography}
-\end{document}
+	The size of larger nets in terms of neurons is counter-productive, since, if we take 50 neurons per layer, we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input dataset would be too small to perform sufficient training.\\

### 2018-07-12 dokumentation/observation.tex
OVERHAUL

--- 2018-07-11 dokumentation/submission.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-%\usepackage{ngerman}
-\usepackage[utf8x]{inputenc}
-\usepackage{fancyvrb}
-\usepackage{courier}
-\usepackage{helvet}
-\usepackage{tikz}
-\usepackage{xcolor}
-\usepackage{pdfpages}
-\usetikzlibrary{calc}
-\usepackage[strict]{changepage}
-\usepackage{xspace}
-\usepackage{hyperref}
-\usetikzlibrary{shapes.geometric}
-\usetikzlibrary{arrows}
-\usetikzlibrary{positioning}
-\usetikzlibrary{arrows.meta, automata, shapes, matrix,positioning}
-\usepackage{amssymb}
-\usepackage{pifont}% http://ctan.org/pkg/pifont
-\usepackage{subcaption} 
-\usepackage{float}
-\usepackage{fixfoot}
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage{multicol}
-\usepackage{amsmath}
-\usepackage{cleveref}
-\usepackage{listings}
-
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\begin{document}
-
-	%
-	%	\title{Clustering Analysis of Mobility Data}
-	\title{Are you moving predictably?}
-	%
-	%\titlerunning{Abbreviated paper title}
-	% If the paper title is too long for the running head, you can set
-	% an abbreviated paper title here
-	%
-	%TODO: Order by Lastname?
-	\author{Miriam Wagner\and
-		Martin Breuer\and
-		Moritz Werthebach\and
-		Timo Bergerbusch\and
-		Walter Schikowski}
-	%
-	\authorrunning{T. Bergerbusch et al.} %TODO
-	% First names are abbreviated in the running head.
-	% If there are more than two authors, 'et al.' is used.
-	%
-	\institute{RWTH Aachen, Templergraben 55, 52062 Aachen, Germany}
-	%
-	\maketitle              % typeset the header of the contribution
-	%
-	\begin{abstract} %TODO
-		We analyze movements in the urban environment of the columbian city Medellín.
-Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
-Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
-Due to bad results we introduce our preprocessing steps and switch to supervised learning.
-Decision trees and neural networks did neither match our performance expectations which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
-
-		\keywords{Data Mining \and Clustering \and Rapidminer \and Cluster \and Neural Nets} %TODO: maybe more
-	\end{abstract}
-	%
-	%
-	%
-	\section{Introduction} \label{sec: introduction}
-	%	TODO (1 page)
-	% could at more in depth 
-	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the data collected?
-In particular is it possible to determine the wealth of a person, only given movements of a single day?
-	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
-There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín.
-All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
-	The data entries consist of data about the movement, like endpoints or length, and also some meta parameters like gender, age or the so-called strata of the person.
-The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.\\
-	Our goal is to ascertain if there is a correlation between the movements and the strata.
-
-	\section{Theoretical backgrounds}
-
-	\subsection{Clustering the data}
-
-	In the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} is a process of grouping data objects from a dataset into multiple groups or clusters.
-The essential criterion for the quality of the clustering is a certain \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.
-
-	In the scope of this work, we decided to use well-known partitioning methods such as k-means or k-medoids.
-In general, given $n$ data objects, partitioning methods distribute the data object into $k$ clusters with $k<n$, using a distance measure to evaluate the repsective similarity.
-Partitioning methods form exclusive clusters by ensuring that each cluster contains at least one object.
-Note that the number $k$ of clusters has to be chosen manually prior to the partitioning process.
-
-	\subsubsection{k-Means}
-
-	is a \textit{centroid based technique}, which means that each cluster is represented by a data point that also is the center of the cluster.
-A distance measure is then used to assigns every remaining data object from the data set to the best fitting cluster according to it's similarity to the center of this cluster and it's dissimilarity to the centers of any other cluster.
-
-	The data objects within a dataset are considered to reside in a euclidean space.
-Thus, the euclidean distance is used to calculate a score for the similarity of two data points.
-When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared erros between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
-This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows:
-	\begin{equation}
-		E = \sum_{i=1}^{k} \sum_{p \in C_i} dist(p, c_i)^2,
-	\end{equation}
-	Given $k$, the first step of k-Means is to select $k$ random objects in the data set as initial representatives for the cluster centers.
-After that, each remaining data object is assigned to the best fitting cluster in terms of similarity, based on a similarity score that is determined by the euclidean distance between the data point and every cluster center.
-The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation.
-In order to achieve this, each cluster centroid is reassigned as the mean of all objects within that cluster.
-By considering the updated centroids, every data point is redistributed to the now best fitting cluster.
-This iterative process will continue, until no better clustering can be found, which means that the latest clustering equals the previous distribution.
-
-	\subsection{Gower Distance}\label{gower}
-
-	When clustering data with algorithms such as \textit{k-means}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm.
-These distance measures come with constraints though, since they are only defined for numerical variables.
-In the scope of this work, we are facing data points with mixed variable types, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points.
-The Gower distance measure distinguishes between three types of variables:\\
-	\\ \textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match \cite{gower1971general}.
-In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}.
-Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values.
-Hence, we consider all variables, that are not explicitly numerical, as categorical variables.\\
-	\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.\\	
-	\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.
-
-	\subsubsection{Calculating the distances.}
-
-	Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
-	\begin{equation}\label{eq1}
-	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
-	\end{equation} 
-	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at position $k$ in the data points $x$ and $y$.
-Note, that the definition of the score dependends on the type of the variable, as defined below.
-In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, when the two variables are comparable and to 0 when not.
-For example, variables are not comparable, if values are undefined in the data points or the variable types do not match at index $k$, among other reasons.
-Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$, the number of considered variables in each data point, for all $x, y$ in the data set.
-Thus, the similarity coefficient in (\ref{eq1}) can be interpreted as the average value of all similarity scores.
-	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
-	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
-As pointed above, 0 values are not considered a match and even further, not considered to be comparable.
-Hence, the values result as in the table
-		\begin{center}
-			\begin{tabular}{l | c c c r}
-				i & 1 & 1 & 0 & 0 \\
-				j & 1 & 0 & 1 & 0 \\
-				\hline
-				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
-				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
-			\end{tabular}
-		\end{center}	
-	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$ and 0, if they differ.\\
-	\\ \textbf{Numerical:} for numerical variables, the similarity score is calculated by
-		\begin{equation*}
-		s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
-		\end{equation*}
-		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
-This can be a global range of acceptable values for variable $k$ or chosen on the basis of the dataset.
-
-	\section{Preprocessing}\label{sec: proprocessing}
-	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
-	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
-	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets considered by our clustering and neural net approaches.\\
-	We observe the following distribution over the whole dataset:\\
-	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
-			abs   & 6963 & 52265 & 49404 & 8772 & 5536 & 2038 &  124978  \\
-			\%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
-		\end{tabular}
-		\label{table: distribution normal}
-	\end{figure}
-	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
-	Furthermore we have to make sure that  two different data points, which belong to the very same person, are assigned to the same cluster.
-To do so we compute the value \texttt{ID} which identifies each person and can be used to combine movements that are considered to be from the same person, i.
-e.
-two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
-This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.
-	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6  & $\Sigma$ \\ \hline
-			abs   & 3153 & 23367 & 21418 & 3497 & 2083 & 595 &  54113   \\
-			\%   & 5.83 & 43.18 & 39.58 & 6.46 & 3.85 & 1.1 &   100
-		\end{tabular}
-	\end{figure}
-	In Section \Cref{subsec: person vector data} we introduce vectors representing single persons.
-Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.
-
-	\subsection{Stratified Person Data}\label{subsec: person vector data}
-	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person %TODO: ref zu code basics
-	, and all movements from that person.\\
-	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:
-	\begin{align*}
-	\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,
-	\underbrace{\mathit{AM}, \mathit{MD}, \mathit{PM}, \mathit{MN}}_{4}, 
-	\underbrace{\#r_1, \dots, \#r_7}_{7}, \\
-	\underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, \underbrace{\mathit{S_{Dest}}, \mathit{S_{Dist}}, \mathit{G}, \mathit{A} ,\mathit{strata}, \mathit{strataGrouped}}_{6}
-	\end{align*}
-	with the following abbreviations ($1 \le i \le 413$, $1 \le j \le 7$):
-	\begin{multicols}{2}
-		\begin{itemize}
-			\setlength{\itemindent}{.4cm}
-			\item[$o_i$:]  the $i$-th origin data point
-			\item[$d_i$:]  the $i$-th destination data point
-			\item[$\mathit{AM}$:] movements at time stamp AM
-			\item[$\mathit{MD}$:] movements at time stamp MD
-			\item[$\mathit{PM}$:] movements at time stamp PM
-			\item[$\mathit{MN}$:] movements at time stamp MN
-			\item[$r_j$:] the $j$-th reason
-			\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
-			\item[$\mathit{S_{Dest}}$:] sum of all durations
-			\item[$\mathit{S_{Dist}}$:] sum of all distances
-			\item[$\mathit{G}$:] the gender
-			\item[$\mathit{A}$:] the age
-			\item[$strata$:] the strata (used for comparison)
-			\item[$strataGrouped$:] the aggregated stratas
-		\end{itemize}
-	\end{multicols}
-	\section{Predicting}
-
-	\subsection{Distance Measures} \label{subsec: distance measures}
-	%TODO: distance measures
-	%TODO (1 page)
-
-	\subsection{Classification} \label{subsec: classification}
-	%TODO: clustering
-	%TODO (5 pages)	
-	{}\input{MiriamClustering}
-
-	\input{DecisionTree}
-	%\subsection{Neural Net}
-	%TODO (5 pages)	
-	\subsection{Neural Net} \label{subsec: neural net}
-	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
-
-	We do this, because results on the normal datasets had an unacceptable performance, since only single movements and not complete paths of individuals are considered.
-An example training and performance measure is given in \Cref{fig: NN without vector} where unprocessed data is used.
-The performance is measured using 10-fold cross validation, i.
-e.
-the data is split into 10 subsets where in each iteration exactly one data set is used as test set and the other 9 as training set.
-The average value of the accuracy values lead
-	to the total accuracy value of the neural net.
-	\begin{figure}[H]
-		\centering
-		\includegraphics[scale = 0.4]{src/pic/NN_without_vector.png}
-		\caption{An example of a neural net trained without person vector data.}
-		\label{fig: NN without vector}
-	\end{figure}
-
-	In the following we consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
-	As an example of other strata aggregation we combine the stratas 1--2, 3--4 and 5--6 together and call them $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$.
-This builds a superset of the original stratas and since the stratas themselves are logically connected this task should be easier to fulfill.\\
-
-	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata which are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
-For every neural net and every set size we performed 5 independent runs and calculated the average over those accuracy values in order to have a sophisticated, comparable statements.
-	\setlength\tabcolsep{.2cm}
-	\begin{figure}[H]
-		\centering
-		\begin{tabular}{|c|c|c|c|c|c|}
-			\hline
-			&   \#    &        & \multicolumn{3}{c|}{Set size} \\
-			Name           & Neurons &   AG   &  100  &  200  &      595      \\ \hline
-			$\mathcal{N}_5$      &    5    & \xmark & 60.03 & 59.92 &     60.18     \\
-			$\mathcal{N}_5^\star$   &    5    & \cmark & 87.6  & 89.7  &     71.05     \\
-			$\mathcal{N}_{10}$    &   10    & \xmark & 75.83 & 73.54 &     69.56     \\
-			$\mathcal{N}_{10}^\star$ &   10    & \cmark & 92.93 & 93.48 &     74.58     \\		
-			%		$\mathcal{N}_{10}^\star$ &   10    & \cmark & 88.33 {\small $\pm$7.49} & 90.67{\small $\pm$2.81} & 92.14 {\small $\pm$ 2.59} \\
-			$\mathcal{N}_{20}$    &   20    & \xmark & 75.45 & 71.14 &     61.87     \\
-			$\mathcal{N}_{20}^\star$ &   20    & \cmark & 92.87 & 94.4  &     78.32     \\ \hline
-		\end{tabular}
-		\caption{The accuracy values of the neural nets. (See excel-spreadsheet)}
-		\label{tab: nn-accuracy}
-		% mittels deep-learning-vector-average
-	\end{figure}
-
-	%TODO: recompute
-	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be too small to have sufficient training.\\
-
-	\section{Observations} \label{sec: observation}
+\section{Conclusion} \label{sec: observation}
 	%	TODO (1 page)
@@
 	Using various datasets (c.f.
-\Cref{sec: proprocessing}), we witnessed that having equally distributed datasets lead to an overall higher accuracy.
+\Cref{sec: proprocessing}), we witnessed, that having equally distributed datasets leads to an overall higher accuracy.
 This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f.
@@
 	During clustering we observed, that using a different distance measure formula would not lead to a huge difference.
-Also, we detected that reducing the numbers of clusters leads to better results, which reduces the significance of result that could be stated.\\
+Also, we detected that reducing the numbers of clusters leads to better results, but decreases the significance of the statement, that could be made.\\
 	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
 	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
-Therefore we see that the more data we have the higher the variance within the stratas themselves is.
-We cannot draw some kind of lines to distinguish between different entries.
-Datapoints, which were outliers in the smaller sets, are now not considered to be outliers, because of large numbers having the same characteristics.
-So the lines, we were able to draw for smaller datasets, are blurring and create some kind of transition phase.\\ %TODO: letzter satz doof
-
-	\section{Discussion} \label{sec: discussion}
-	%TODO: discussion
-%	TODO (2 page)
-%	\begin{itemize}
-%		\item repr. working day
-%		\item over/under std's
-%		\item indiv. of lifestyle
-%		\item inv. behaviour in same strata
-%	\end{itemize}
-	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we though about various aspects having an impact on the movement.
-We state 4 main influencing aspects, explaining the variance throughout the data.
-
-	\subsection{Representativity of the day}
-	The day the data was collected on (c.f.
-\Cref{sec: introduction}) is not mentioned.
-Therefore we can not infer that it is representative.
-The people asked to enter their movement could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
-\\
-	Also the day itself is not mentioned, so we don't have information if it was even a working day or a weekend. This influences the behavior drastically.
-
-	\subsection{People living over/under standards}
-	%TODO
-
-	\subsection{Individuality of lifestyle}
-	Obviously, people are individual in their way of life.
-So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply don't like driving.
-On the other end of the spectrum, some people, who have little money still own and drive a car daily in order to go to work, since they can't afford an other apartment.\\
-	One can think of multitude scenarios, of people behaving different, cause by their individuality.
-
-	\subsection{Inverse behavior in a strata}
-	As an example, take a look at the strata 6, which denotes the richest people considered. Their we have inter alia two groups: 
-	\begin{enumerate}
-		\setlength{\itemindent}{1cm}
-		\item[1. Group]
-		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
-		\item[2. Group]
-		People, who are rich just by birth, which don't have the necessity to work or move at all. They could possibly stay home and don't have to leave at all.
-	\end{enumerate} 
-	Those two groups are completely opposite, but still belong to the same strata.
-Now also considering strata one, we can find the exact same movement behavior in there as well.
-Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work.
-Also there are people, who don't move at all, because of the same reason.
-
-	So we can see that based on the information we have it is very unlikely to have a precise correlation between the given data and the strata.
-However given more information about the previously mentioned aspects and more entries in general there could be some kind correlation to be found.\\
-	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
-Also one could use different network visualization tools in order to infer their own patterns.
-An example would be to take the origin (destination) sectors as nodes, an edge if there is a movement, directed or undirected, and a different thickness or color gradient based on the number of this edge being taken.\\
-
-	Overall we can conclude that with the data we got we are not able to find a correlation.
-
-	\begin{thebibliography}{8}
-		\bibitem{rich_do_not_rise_early}
-		Lotero, Laura, et al. "Rich do not rise early: spatio-temporal patterns in the mobility networks of different socio-economic classes." Royal Society open science 3.10 (2016): 150654.
-
-		\bibitem{strata_def}
-		Hudson, Rex A. Colombia: A country study. Government Printing Office, 2010.
-
-		\bibitem{gower1971general}
-		Gower, John C,
-		\emph{A general coefficient of similarity and some of its properties},
-		Biometrics, pp. 857-871,
-		1971.
-
-		\bibitem{data_mining}
-		Han, Jiawei and Pei, Jian and Kamber, Micheline,
-		\emph{Data mining: concepts and techniques},
-		Elsevier, pp. 444-454,
-		2011.
-	\end{thebibliography}
-\end{document}
+Therefore we see, the more data we have, the higher is the variance within the stratas themselves.
+We cannot draw some kind of lines to distinguish between different stratas.
+Data points, which were outliers in the smaller sets, are now not considered to be outliers, because many others have the same characteristics.
+So the lines, we were able to draw for smaller datasets, are blurring.\\ %TODO: letzter satz doof	-> hihi

### 2018-07-12 dokumentation/preprocessing.tex
OVERHAUL

--- 2018-07-11 dokumentation/submission.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-%\usepackage{ngerman}
-\usepackage[utf8x]{inputenc}
-\usepackage{fancyvrb}
-\usepackage{courier}
-\usepackage{helvet}
-\usepackage{tikz}
-\usepackage{xcolor}
-\usepackage{pdfpages}
-\usetikzlibrary{calc}
-\usepackage[strict]{changepage}
-\usepackage{xspace}
-\usepackage{hyperref}
-\usetikzlibrary{shapes.geometric}
-\usetikzlibrary{arrows}
-\usetikzlibrary{positioning}
-\usetikzlibrary{arrows.meta, automata, shapes, matrix,positioning}
-\usepackage{amssymb}
-\usepackage{pifont}% http://ctan.org/pkg/pifont
-\usepackage{subcaption} 
-\usepackage{float}
-\usepackage{fixfoot}
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage{multicol}
-\usepackage{amsmath}
-\usepackage{cleveref}
-\usepackage{listings}
-
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\begin{document}
-
-	%
-	%	\title{Clustering Analysis of Mobility Data}
-	\title{Are you moving predictably?}
-	%
-	%\titlerunning{Abbreviated paper title}
-	% If the paper title is too long for the running head, you can set
-	% an abbreviated paper title here
-	%
-	%TODO: Order by Lastname?
-	\author{Miriam Wagner\and
-		Martin Breuer\and
-		Moritz Werthebach\and
-		Timo Bergerbusch\and
-		Walter Schikowski}
-	%
-	\authorrunning{T. Bergerbusch et al.} %TODO
-	% First names are abbreviated in the running head.
-	% If there are more than two authors, 'et al.' is used.
-	%
-	\institute{RWTH Aachen, Templergraben 55, 52062 Aachen, Germany}
-	%
-	\maketitle              % typeset the header of the contribution
-	%
-	\begin{abstract} %TODO
-		We analyze movements in the urban environment of the columbian city Medellín.
-Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
-Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
-Due to bad results we introduce our preprocessing steps and switch to supervised learning.
-Decision trees and neural networks did neither match our performance expectations which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
-
-		\keywords{Data Mining \and Clustering \and Rapidminer \and Cluster \and Neural Nets} %TODO: maybe more
-	\end{abstract}
-	%
-	%
-	%
-	\section{Introduction} \label{sec: introduction}
-	%	TODO (1 page)
-	% could at more in depth 
-	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the data collected?
-In particular is it possible to determine the wealth of a person, only given movements of a single day?
-	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
-There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín.
-All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
-	The data entries consist of data about the movement, like endpoints or length, and also some meta parameters like gender, age or the so-called strata of the person.
-The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.\\
-	Our goal is to ascertain if there is a correlation between the movements and the strata.
-
-	\section{Theoretical backgrounds}
-
-	\subsection{Clustering the data}
-
-	In the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} is a process of grouping data objects from a dataset into multiple groups or clusters.
-The essential criterion for the quality of the clustering is a certain \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.
-
-	In the scope of this work, we decided to use well-known partitioning methods such as k-means or k-medoids.
-In general, given $n$ data objects, partitioning methods distribute the data object into $k$ clusters with $k<n$, using a distance measure to evaluate the repsective similarity.
-Partitioning methods form exclusive clusters by ensuring that each cluster contains at least one object.
-Note that the number $k$ of clusters has to be chosen manually prior to the partitioning process.
-
-	\subsubsection{k-Means}
-
-	is a \textit{centroid based technique}, which means that each cluster is represented by a data point that also is the center of the cluster.
-A distance measure is then used to assigns every remaining data object from the data set to the best fitting cluster according to it's similarity to the center of this cluster and it's dissimilarity to the centers of any other cluster.
-
-	The data objects within a dataset are considered to reside in a euclidean space.
-Thus, the euclidean distance is used to calculate a score for the similarity of two data points.
-When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared erros between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
-This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows:
-	\begin{equation}
-		E = \sum_{i=1}^{k} \sum_{p \in C_i} dist(p, c_i)^2,
-	\end{equation}
-	Given $k$, the first step of k-Means is to select $k$ random objects in the data set as initial representatives for the cluster centers.
-After that, each remaining data object is assigned to the best fitting cluster in terms of similarity, based on a similarity score that is determined by the euclidean distance between the data point and every cluster center.
-The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation.
-In order to achieve this, each cluster centroid is reassigned as the mean of all objects within that cluster.
-By considering the updated centroids, every data point is redistributed to the now best fitting cluster.
-This iterative process will continue, until no better clustering can be found, which means that the latest clustering equals the previous distribution.
-
-	\subsection{Gower Distance}\label{gower}
-
-	When clustering data with algorithms such as \textit{k-means}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm.
-These distance measures come with constraints though, since they are only defined for numerical variables.
-In the scope of this work, we are facing data points with mixed variable types, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points.
-The Gower distance measure distinguishes between three types of variables:\\
-	\\ \textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match \cite{gower1971general}.
-In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}.
-Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values.
-Hence, we consider all variables, that are not explicitly numerical, as categorical variables.\\
-	\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.\\	
-	\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.
-
-	\subsubsection{Calculating the distances.}
-
-	Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
-	\begin{equation}\label{eq1}
-	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
-	\end{equation} 
-	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at position $k$ in the data points $x$ and $y$.
-Note, that the definition of the score dependends on the type of the variable, as defined below.
-In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, when the two variables are comparable and to 0 when not.
-For example, variables are not comparable, if values are undefined in the data points or the variable types do not match at index $k$, among other reasons.
-Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$, the number of considered variables in each data point, for all $x, y$ in the data set.
-Thus, the similarity coefficient in (\ref{eq1}) can be interpreted as the average value of all similarity scores.
-	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
-	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
-As pointed above, 0 values are not considered a match and even further, not considered to be comparable.
-Hence, the values result as in the table
-		\begin{center}
-			\begin{tabular}{l | c c c r}
-				i & 1 & 1 & 0 & 0 \\
-				j & 1 & 0 & 1 & 0 \\
-				\hline
-				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
-				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
-			\end{tabular}
-		\end{center}	
-	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$ and 0, if they differ.\\
-	\\ \textbf{Numerical:} for numerical variables, the similarity score is calculated by
-		\begin{equation*}
-		s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
-		\end{equation*}
-		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
-This can be a global range of acceptable values for variable $k$ or chosen on the basis of the dataset.
-
-	\section{Preprocessing}\label{sec: proprocessing}
+\subsection{Preprocessing}\label{sec: proprocessing}
 	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
 	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
-	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets considered by our clustering and neural net approaches.\\
+	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets used by our clustering and neural net approaches.\\
 	We observe the following distribution over the whole dataset:\\
+
+	\vspace*{-3em}
 	\begin{figure}[H]
@@
 		\end{tabular}
+		\vspace*{-2.5em}
 		\label{table: distribution normal}
 	\end{figure}
-	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
+	There is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
 	Furthermore we have to make sure that two different data points, which belong to the very same person, are assigned to the same cluster.
-To do so we compute the value \texttt{ID} which identifies each person and can be used to combine movements that are considered to be from the same person, i.
-e.
+To do so, we compute the value \texttt{ID}, which identifies each person and can be used to combine movements that are considered to be from the same person.
+I.e.
 two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
-This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.
+This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.\\
+
+	\vspace*{-3em}
 	\begin{figure}[H]
@@
 		\end{tabular}
+		\vspace*{-2.5em}
 	\end{figure}
-	In Section \Cref{subsec: person vector data} we introduce vectors representing single persons.
-Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.
+	Following, we introduce vectors representing single persons.
+Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.\\
 
-	\subsection{Stratified Person Data}\label{subsec: person vector data}
+	\textbf{Stratified Person Data}\label{subsubsec: person vector data}\\
 	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person %TODO: ref zu code basics
-	, and all movements from that person.\\
+This class stores the ID, the parameters defining a person (c.f.
+\Cref{sec: proprocessing}), and all movements from that person.\\
 	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:
@@
 	with the following abbreviations ($1 \le i \le 413$, $1 \le j \le 7$):
-	\begin{multicols}{2}
+	\begin{figure}[H]
+		\centering
+		\hspace*{-.3cm}
+		\begin{subfigure}{0.50\textwidth}
 		\begin{itemize}
@@
 			\item[$r_j$:] the $j$-th reason
+		\end{itemize}
+	\end{subfigure}\hspace*{1cm}
+	\begin{subfigure}{0.48\textwidth}
+	\begin{itemize}
 			\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
@@
 		\end{itemize}
-	\end{multicols}
-	\section{Predicting}
-
-	\subsection{Distance Measures} \label{subsec: distance measures}
-	%TODO: distance measures
-	%TODO (1 page)
-
-	\subsection{Classification} \label{subsec: classification}
-	%TODO: clustering
-	%TODO (5 pages)	
-	{}\input{MiriamClustering}
-
-	\input{DecisionTree}
-	%\subsection{Neural Net}
-	%TODO (5 pages)	
-	\subsection{Neural Net} \label{subsec: neural net}
-	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
-
-	We do this, because results on the normal datasets had an unacceptable performance, since only single movements and not complete paths of individuals are considered.
-An example training and performance measure is given in \Cref{fig: NN without vector} where unprocessed data is used.
-The performance is measured using 10-fold cross validation, i.
-e.
-the data is split into 10 subsets where in each iteration exactly one data set is used as test set and the other 9 as training set.
-The average value of the accuracy values lead
-	to the total accuracy value of the neural net.
-	\begin{figure}[H]
-		\centering
-		\includegraphics[scale = 0.4]{src/pic/NN_without_vector.png}
-		\caption{An example of a neural net trained without person vector data.}
-		\label{fig: NN without vector}
+	\end{subfigure}
 	\end{figure}
-
-	In the following we consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
-	As an example of other strata aggregation we combine the stratas 1--2, 3--4 and 5--6 together and call them $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$.
-This builds a superset of the original stratas and since the stratas themselves are logically connected this task should be easier to fulfill.\\
-
-	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata which are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
-For every neural net and every set size we performed 5 independent runs and calculated the average over those accuracy values in order to have a sophisticated, comparable statements.
-	\setlength\tabcolsep{.2cm}
-	\begin{figure}[H]
-		\centering
-		\begin{tabular}{|c|c|c|c|c|c|}
-			\hline
-			&   \#    &        & \multicolumn{3}{c|}{Set size} \\
-			Name           & Neurons &   AG   &  100  &  200  &      595      \\ \hline
-			$\mathcal{N}_5$      &    5    & \xmark & 60.03 & 59.92 &     60.18     \\
-			$\mathcal{N}_5^\star$   &    5    & \cmark & 87.6  & 89.7  &     71.05     \\
-			$\mathcal{N}_{10}$    &   10    & \xmark & 75.83 & 73.54 &     69.56     \\
-			$\mathcal{N}_{10}^\star$ &   10    & \cmark & 92.93 & 93.48 &     74.58     \\		
-			%		$\mathcal{N}_{10}^\star$ &   10    & \cmark & 88.33 {\small $\pm$7.49} & 90.67{\small $\pm$2.81} & 92.14 {\small $\pm$ 2.59} \\
-			$\mathcal{N}_{20}$    &   20    & \xmark & 75.45 & 71.14 &     61.87     \\
-			$\mathcal{N}_{20}^\star$ &   20    & \cmark & 92.87 & 94.4  &     78.32     \\ \hline
-		\end{tabular}
-		\caption{The accuracy values of the neural nets. (See excel-spreadsheet)}
-		\label{tab: nn-accuracy}
-		% mittels deep-learning-vector-average
-	\end{figure}
-
-	%TODO: recompute
-	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be too small to have sufficient training.\\
-
-	\section{Observations} \label{sec: observation}
-	%	TODO (1 page)
-	%	TODO: restructure. Have like 3 parts
-	%	The greater the population, the more the borders between the stratas blur. 
-	As an overall result we get, that we can not determine the strata based on the information we have.
-	Using various datasets (c.f.
-\Cref{sec: proprocessing}), we witnessed that having equally distributed datasets lead to an overall higher accuracy.
-This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f.
-\Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
-	During clustering we observed, that using a different distance measure formula would not lead to a huge difference.
-Also, we detected that reducing the numbers of clusters leads to better results, which reduces the significance of result that could be stated.\\
-	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
-	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
-Therefore we see that the more data we have the higher the variance within the stratas themselves is.
-We cannot draw some kind of lines to distinguish between different entries.
-Datapoints, which were outliers in the smaller sets, are now not considered to be outliers, because of large numbers having the same characteristics.
-So the lines, we were able to draw for smaller datasets, are blurring and create some kind of transition phase.\\ %TODO: letzter satz doof
-
-	\section{Discussion} \label{sec: discussion}
-	%TODO: discussion
-%	TODO (2 page)
-%	\begin{itemize}
-%		\item repr. working day
-%		\item over/under std's
-%		\item indiv. of lifestyle
-%		\item inv. behaviour in same strata
-%	\end{itemize}
-	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we though about various aspects having an impact on the movement.
-We state 4 main influencing aspects, explaining the variance throughout the data.
-
-	\subsection{Representativity of the day}
-	The day the data was collected on (c.f.
-\Cref{sec: introduction}) is not mentioned.
-Therefore we can not infer that it is representative.
-The people asked to enter their movement could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
-\\
-	Also the day itself is not mentioned, so we don't have information if it was even a working day or a weekend. This influences the behavior drastically.
-
-	\subsection{People living over/under standards}
-	%TODO
-
-	\subsection{Individuality of lifestyle}
-	Obviously, people are individual in their way of life.
-So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply don't like driving.
-On the other end of the spectrum, some people, who have little money still own and drive a car daily in order to go to work, since they can't afford an other apartment.\\
-	One can think of multitude scenarios, of people behaving different, cause by their individuality.
-
-	\subsection{Inverse behavior in a strata}
-	As an example, take a look at the strata 6, which denotes the richest people considered. Their we have inter alia two groups: 
-	\begin{enumerate}
-		\setlength{\itemindent}{1cm}
-		\item[1. Group]
-		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
-		\item[2. Group]
-		People, who are rich just by birth, which don't have the necessity to work or move at all. They could possibly stay home and don't have to leave at all.
-	\end{enumerate} 
-	Those two groups are completely opposite, but still belong to the same strata.
-Now also considering strata one, we can find the exact same movement behavior in there as well.
-Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work.
-Also there are people, who don't move at all, because of the same reason.
-
-	So we can see that based on the information we have it is very unlikely to have a precise correlation between the given data and the strata.
-However given more information about the previously mentioned aspects and more entries in general there could be some kind correlation to be found.\\
-	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
-Also one could use different network visualization tools in order to infer their own patterns.
-An example would be to take the origin (destination) sectors as nodes, an edge if there is a movement, directed or undirected, and a different thickness or color gradient based on the number of this edge being taken.\\
-
-	Overall we can conclude that with the data we got we are not able to find a correlation.
-
-	\begin{thebibliography}{8}
-		\bibitem{rich_do_not_rise_early}
-		Lotero, Laura, et al. "Rich do not rise early: spatio-temporal patterns in the mobility networks of different socio-economic classes." Royal Society open science 3.10 (2016): 150654.
-
-		\bibitem{strata_def}
-		Hudson, Rex A. Colombia: A country study. Government Printing Office, 2010.
-
-		\bibitem{gower1971general}
-		Gower, John C,
-		\emph{A general coefficient of similarity and some of its properties},
-		Biometrics, pp. 857-871,
-		1971.
-
-		\bibitem{data_mining}
-		Han, Jiawei and Pei, Jian and Kamber, Micheline,
-		\emph{Data mining: concepts and techniques},
-		Elsevier, pp. 444-454,
-		2011.
-	\end{thebibliography}
-\end{document}

### 2018-07-12 dokumentation/submission.tex
OVERHAUL

--- 2018-07-11 dokumentation/submission.tex
@@
 \usepackage{listings}
-
+\setlength{\parindent}{0em}
 \newcommand{\cmark}{\ding{51}}%
@@
 	%
-	\begin{abstract} %TODO
-		We analyze movements in the urban environment of the columbian city Medellín.
+	\begin{abstract}
+		We analyze movements in the urban environment of the Columbian city Medellín.
 Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
-Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
+		% Marías sentence
+		The objective of the paper is twofold: try to find groups of movement patterns and see, whether they correspond to socio-economic status, and then, since we have the economic status of each person in the data set, we will apply supervised learning to classify patterns into socio-economic status.
+%		Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
 Due to bad results we introduce our preprocessing steps and switch to supervised learning.
-Decision trees and neural networks did neither match our performance expectations which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
+		Decision trees and neural networks did neither match our performance expectations, which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
 
-		\keywords{Data Mining \and Clustering \and Rapidminer \and Cluster \and Neural Nets} %TODO: maybe more
+		\keywords{Data Mining \and Clustering \and Rapidminer \and Cluster \and Neural Nets}
 	\end{abstract}
@@
 	%	
-	\section{Introduction} \label{sec: introduction}
-	%	TODO (1 page)
-	% could at more in depth 
-	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the data collected?
-In particular is it possible to determine the wealth of a person, only given movements of a single day?
-	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
-There we have a set of 124979 rows of movement data from various persons from a Columbian town, called Medellín.
-All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
-	The data entries consist of data about the movement, like endpoints or length, and also some meta parameters like gender, age or the so-called strata of the person.
-The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.\\
-	Our goal is to ascertain if there is a correlation between the movements and the strata.
+	\input{introduction.tex}
 
-	\section{Theoretical backgrounds}
+	\input{theoreticalBackground.tex}
 
-	\subsection{Clustering the data}
+%	\input{preprocessing.tex}
 
-	In the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} is a process of grouping data objects from a dataset into multiple groups or clusters.
-The essential criterion for the quality of the clustering is a certain \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.
+	\section{Results}
 
-	In the scope of this work, we decided to use well-known partitioning methods such as k-means or k-medoids.
-In general, given $n$ data objects, partitioning methods distribute the data object into $k$ clusters with $k<n$, using a distance measure to evaluate the repsective similarity.
-Partitioning methods form exclusive clusters by ensuring that each cluster contains at least one object.
-Note that the number $k$ of clusters has to be chosen manually prior to the partitioning process.
+	\subsection{Natural Clusters} \label{subsec: classification}
 
-	\subsubsection{k-Means}
-
-	is a \textit{centroid based technique}, which means that each cluster is represented by a data point that also is the center of the cluster.
-A distance measure is then used to assigns every remaining data object from the data set to the best fitting cluster according to it's similarity to the center of this cluster and it's dissimilarity to the centers of any other cluster.
-
-	The data objects within a dataset are considered to reside in a euclidean space.
-Thus, the euclidean distance is used to calculate a score for the similarity of two data points.
-When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared erros between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
-This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows:
-	\begin{equation}
-		E = \sum_{i=1}^{k} \sum_{p \in C_i} dist(p, c_i)^2,
-	\end{equation}
-	Given $k$, the first step of k-Means is to select $k$ random objects in the data set as initial representatives for the cluster centers.
-After that, each remaining data object is assigned to the best fitting cluster in terms of similarity, based on a similarity score that is determined by the euclidean distance between the data point and every cluster center.
-The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation.
-In order to achieve this, each cluster centroid is reassigned as the mean of all objects within that cluster.
-By considering the updated centroids, every data point is redistributed to the now best fitting cluster.
-This iterative process will continue, until no better clustering can be found, which means that the latest clustering equals the previous distribution.
-
-	\subsection{Gower Distance}\label{gower}
-
-	When clustering data with algorithms such as \textit{k-means}, the distance between datapoints is usually calculated by measures such as the Euclidean or Manhatten norm.
-These distance measures come with constraints though, since they are only defined for numerical variables.
-In the scope of this work, we are facing data points with mixed variable types, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points.
-The Gower distance measure distinguishes between three types of variables:\\
-	\\ \textbf{Binary} variables, where two variables of value 0 are \textit{not} considered a match \cite{gower1971general}.
-In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}.
-Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values.
-Hence, we consider all variables, that are not explicitly numerical, as categorical variables.\\
-	\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.\\	
-	\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.
-
-	\subsubsection{Calculating the distances.}
-
-	Given two data points $x$ and $y$, that each form a tuple of $v$ variables of arbitrary type, the similarity coefficient between the two points is given by
-	\begin{equation}\label{eq1}
-	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
-	\end{equation} 
-	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at position $k$ in the data points $x$ and $y$.
-Note, that the definition of the score dependends on the type of the variable, as defined below.
-In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, when the two variables are comparable and to 0 when not.
-For example, variables are not comparable, if values are undefined in the data points or the variable types do not match at index $k$, among other reasons.
-Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$, the number of considered variables in each data point, for all $x, y$ in the data set.
-Thus, the similarity coefficient in (\ref{eq1}) can be interpreted as the average value of all similarity scores.
-	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
-	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
-As pointed above, 0 values are not considered a match and even further, not considered to be comparable.
-Hence, the values result as in the table
-		\begin{center}
-			\begin{tabular}{l | c c c r}
-				i & 1 & 1 & 0 & 0 \\
-				j & 1 & 0 & 1 & 0 \\
-				\hline
-				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
-				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
-			\end{tabular}
-		\end{center}	
-	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$ and 0, if they differ.\\
-	\\ \textbf{Numerical:} for numerical variables, the similarity score is calculated by
-		\begin{equation*}
-		s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
-		\end{equation*}
-		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
-This can be a global range of acceptable values for variable $k$ or chosen on the basis of the dataset.
-
-	\section{Preprocessing}\label{sec: proprocessing}
-	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
-	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
-	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets considered by our clustering and neural net approaches.\\
-	We observe the following distribution over the whole dataset:\\
-	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
-			abs   & 6963 & 52265 & 49404 & 8772 & 5536 & 2038 &  124978  \\
-			\%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
-		\end{tabular}
-		\label{table: distribution normal}
-	\end{figure}
-	We observe that there is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
-	Furthermore we have to make sure that  two different data points, which belong to the very same person, are assigned to the same cluster.
-To do so we compute the value \texttt{ID} which identifies each person and can be used to combine movements that are considered to be from the same person, i.
-e.
-two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender.
-This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.
-	\begin{figure}[H]
-		\centering
-		\setlength\tabcolsep{.2cm}
-		\begin{tabular}{c|ccccccc}
-			strata &  1   &   2   &   3   &  4   &  5   &  6  & $\Sigma$ \\ \hline
-			abs   & 3153 & 23367 & 21418 & 3497 & 2083 & 595 &  54113   \\
-			\%   & 5.83 & 43.18 & 39.58 & 6.46 & 3.85 & 1.1 &   100
-		\end{tabular}
-	\end{figure}
-	In Section \Cref{subsec: person vector data} we introduce vectors representing single persons.
-Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.
-
-	\subsection{Stratified Person Data}\label{subsec: person vector data}
-	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person %TODO: ref zu code basics
-	, and all movements from that person.\\
-	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:
-	\begin{align*}
-	\underbrace{\#o_1, \dots, \#o_{413}, \#d_1, \dots, \#d_{413}}_{2\cdot 413} ,
-	\underbrace{\mathit{AM}, \mathit{MD}, \mathit{PM}, \mathit{MN}}_{4}, 
-	\underbrace{\#r_1, \dots, \#r_7}_{7}, \\
-	\underbrace{\#\mathit{MoT}_1, \dots, \#\mathit{MoT}_7}_{7}, \underbrace{\mathit{S_{Dest}}, \mathit{S_{Dist}}, \mathit{G}, \mathit{A} ,\mathit{strata}, \mathit{strataGrouped}}_{6}
-	\end{align*}
-	with the following abbreviations ($1 \le i \le 413$, $1 \le j \le 7$):
-	\begin{multicols}{2}
-		\begin{itemize}
-			\setlength{\itemindent}{.4cm}
-			\item[$o_i$:]  the $i$-th origin data point
-			\item[$d_i$:]  the $i$-th destination data point
-			\item[$\mathit{AM}$:] movements at time stamp AM
-			\item[$\mathit{MD}$:] movements at time stamp MD
-			\item[$\mathit{PM}$:] movements at time stamp PM
-			\item[$\mathit{MN}$:] movements at time stamp MN
-			\item[$r_j$:] the $j$-th reason
-			\item[$\mathit{MoT}_j$:] the $j$-th mean of transportation
-			\item[$\mathit{S_{Dest}}$:] sum of all durations
-			\item[$\mathit{S_{Dist}}$:] sum of all distances
-			\item[$\mathit{G}$:] the gender
-			\item[$\mathit{A}$:] the age
-			\item[$strata$:] the strata (used for comparison)
-			\item[$strataGrouped$:] the aggregated stratas
-		\end{itemize}
-	\end{multicols}
-	\section{Predicting}
-
-	\subsection{Distance Measures} \label{subsec: distance measures}
-	%TODO: distance measures
-	%TODO (1 page)
-
-	\subsection{Classification} \label{subsec: classification}
-	%TODO: clustering
-	%TODO (5 pages)	
-	{}\input{MiriamClustering}
+	\input{MiriamClustering}
 
 	\input{DecisionTree}
-	%\subsection{Neural Net}
-	%TODO (5 pages)	
-	\subsection{Neural Net} \label{subsec: neural net}
-	For all the neural net computations we considered person vector data sets of different sizes (c.f. \Cref{subsec: person vector data}).\\
 
-	We do this, because results on the normal datasets had an unacceptable performance, since only single movements and not complete paths of individuals are considered.
-An example training and performance measure is given in \Cref{fig: NN without vector} where unprocessed data is used.
-The performance is measured using 10-fold cross validation, i.
-e.
-the data is split into 10 subsets where in each iteration exactly one data set is used as test set and the other 9 as training set.
-The average value of the accuracy values lead
-	to the total accuracy value of the neural net.
-	\begin{figure}[H]
-		\centering
-		\includegraphics[scale = 0.4]{src/pic/NN_without_vector.png}
-		\caption{An example of a neural net trained without person vector data.}
-		\label{fig: NN without vector}
-	\end{figure}
+	\input{neuralnet.tex}
 
-	In the following we consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
-	As an example of other strata aggregation we combine the stratas 1--2, 3--4 and 5--6 together and call them $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$.
-This builds a superset of the original stratas and since the stratas themselves are logically connected this task should be easier to fulfill.\\
+	\input{observation.tex}
 
-	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata which are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
-For every neural net and every set size we performed 5 independent runs and calculated the average over those accuracy values in order to have a sophisticated, comparable statements.
-	\setlength\tabcolsep{.2cm}
-	\begin{figure}[H]
-		\centering
-		\begin{tabular}{|c|c|c|c|c|c|}
-			\hline
-			&   \#    &        & \multicolumn{3}{c|}{Set size} \\
-			Name           & Neurons &   AG   &  100  &  200  &      595      \\ \hline
-			$\mathcal{N}_5$      &    5    & \xmark & 60.03 & 59.92 &     60.18     \\
-			$\mathcal{N}_5^\star$   &    5    & \cmark & 87.6  & 89.7  &     71.05     \\
-			$\mathcal{N}_{10}$    &   10    & \xmark & 75.83 & 73.54 &     69.56     \\
-			$\mathcal{N}_{10}^\star$ &   10    & \cmark & 92.93 & 93.48 &     74.58     \\		
-			%		$\mathcal{N}_{10}^\star$ &   10    & \cmark & 88.33 {\small $\pm$7.49} & 90.67{\small $\pm$2.81} & 92.14 {\small $\pm$ 2.59} \\
-			$\mathcal{N}_{20}$    &   20    & \xmark & 75.45 & 71.14 &     61.87     \\
-			$\mathcal{N}_{20}^\star$ &   20    & \cmark & 92.87 & 94.4  &     78.32     \\ \hline
-		\end{tabular}
-		\caption{The accuracy values of the neural nets. (See excel-spreadsheet)}
-		\label{tab: nn-accuracy}
-		% mittels deep-learning-vector-average
-	\end{figure}
-
-	%TODO: recompute
-	The size of larger nets in terms of neurons is counter-productive, since if we take 50 neurons per layer we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input data set would be too small to have sufficient training.\\
-
-	\section{Observations} \label{sec: observation}
-	%	TODO (1 page)
-	%	TODO: restructure. Have like 3 parts
-	%	The greater the population, the more the borders between the stratas blur. 
-	As an overall result we get, that we can not determine the strata based on the information we have.
-	Using various datasets (c.f.
-\Cref{sec: proprocessing}), we witnessed that having equally distributed datasets lead to an overall higher accuracy.
-This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f.
-\Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
-	During clustering we observed, that using a different distance measure formula would not lead to a huge difference.
-Also, we detected that reducing the numbers of clusters leads to better results, which reduces the significance of result that could be stated.\\
-	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
-	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
-Therefore we see that the more data we have the higher the variance within the stratas themselves is.
-We cannot draw some kind of lines to distinguish between different entries.
-Datapoints, which were outliers in the smaller sets, are now not considered to be outliers, because of large numbers having the same characteristics.
-So the lines, we were able to draw for smaller datasets, are blurring and create some kind of transition phase.\\ %TODO: letzter satz doof
-
-	\section{Discussion} \label{sec: discussion}
-	%TODO: discussion
-%	TODO (2 page)
-%	\begin{itemize}
-%		\item repr. working day
-%		\item over/under std's
-%		\item indiv. of lifestyle
-%		\item inv. behaviour in same strata
-%	\end{itemize}
-	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we though about various aspects having an impact on the movement.
-We state 4 main influencing aspects, explaining the variance throughout the data.
-
-	\subsection{Representativity of the day}
-	The day the data was collected on (c.f.
-\Cref{sec: introduction}) is not mentioned.
-Therefore we can not infer that it is representative.
-The people asked to enter their movement could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
-\\
-	Also the day itself is not mentioned, so we don't have information if it was even a working day or a weekend. This influences the behavior drastically.
-
-	\subsection{People living over/under standards}
-	%TODO
-
-	\subsection{Individuality of lifestyle}
-	Obviously, people are individual in their way of life.
-So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply don't like driving.
-On the other end of the spectrum, some people, who have little money still own and drive a car daily in order to go to work, since they can't afford an other apartment.\\
-	One can think of multitude scenarios, of people behaving different, cause by their individuality.
-
-	\subsection{Inverse behavior in a strata}
-	As an example, take a look at the strata 6, which denotes the richest people considered. Their we have inter alia two groups: 
-	\begin{enumerate}
-		\setlength{\itemindent}{1cm}
-		\item[1. Group]
-		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
-		\item[2. Group]
-		People, who are rich just by birth, which don't have the necessity to work or move at all. They could possibly stay home and don't have to leave at all.
-	\end{enumerate} 
-	Those two groups are completely opposite, but still belong to the same strata.
-Now also considering strata one, we can find the exact same movement behavior in there as well.
-Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work.
-Also there are people, who don't move at all, because of the same reason.
-
-	So we can see that based on the information we have it is very unlikely to have a precise correlation between the given data and the strata.
-However given more information about the previously mentioned aspects and more entries in general there could be some kind correlation to be found.\\
-	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
-Also one could use different network visualization tools in order to infer their own patterns.
-An example would be to take the origin (destination) sectors as nodes, an edge if there is a movement, directed or undirected, and a different thickness or color gradient based on the number of this edge being taken.\\
-
-	Overall we can conclude that with the data we got we are not able to find a correlation.
+	\input{conclusion}
 
@@
 		2011.
+
+		\bibitem{population_number}
+		Estimates and projections of the total national, departmental and municipal population by area 1985-2020 (XLS). NADS. Retrieved 1 September 2014.
+
+		\bibitem{neuralNet}
+		Nasrabadi, Nasser M. "Pattern recognition and machine learning." Journal of electronic imaging 16.4 (2007): 049901.
 	\end{thebibliography}

### 2018-07-12 dokumentation/theoreticalBackground.tex
OVERHAUL

\section{Methods}

	\input{preprocessing.tex}

	\subsection{Unsupervised learning}
	\subsubsection{Clustering the data}

	is in the field of \textit{Data Mining}, \textit{Cluster Analysis} or \textit{Clustering} a process of grouping data objects from a dataset into multiple groups. The essential criterion, for the quality of the clustering, is \textit{similarity}, such that data objects are similar to other objects in the same cluster and dissimilar to objects from other clusters.

	In the scope of this work, we decided to use the well-known partitioning methods k-means. In general, given $n$ data objects partitioning methods distribute the data objects into $k$ clusters with $k\le n$, using a distance measure to evaluate the respective similarity. 
%	Those methods form exclusive clusters by ensuring that each cluster contains at least one object.
	Note that the number $k$ of clusters has to be chosen manually a-priori and given to the partitioning process.

	\subsubsection{k-Means}

	is a \textit{centroid based technique}, which means that each cluster is represented by a data point, that also is the centre of the cluster. A distance measure is then used to assign every remaining data object from the data set to the best fitting cluster. This is done according to its similarity to the centre of this cluster and its dissimilarity to the centres of any other cluster.

	The data objects within a dataset are considered to reside in a euclidean space. Thus, the euclidean distance is used to calculate a score for the similarity of two data points. When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared errors between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster. This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows: 
	\begin{equation}
		E = \sum_{i=1}^{k} \sum_{p \in C_i} dist(p, c_i)^2,
	\end{equation}
	Given $k$, the first step of k-Means is to select $k$ random points as cluster centres. Those do not have to be actual data points.
%	objects in the data set as initial representatives for the cluster centres.
	After that, each remaining data object is assigned to the best fitting cluster based on a similarity score between the data point and every cluster centre. The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation. In order to achieve this, each cluster centroid is redefined as the mean of all objects within that cluster. By considering the updated centroids, every data point is reassigned to the now best fitting cluster. This iterative process will continue until no better clustering can be found.\\
%	\subsubsection{Clustering with RapidMiner}	
	As a tool we considered RapidMiner, since it has many modules already efficiently implemented and is easy to adapt.
	\vspace*{-2em}
	\begin{figure}[!htbp]
		\centering
		\includegraphics[width=0.9\textwidth]{ClusteringRapid.PNG}
		\caption{Process of k-means clustering}
		\label{fig: kclust}
		\vspace*{-2em}
	\end{figure}

	The Process, \Cref{fig: kclust}, contains the following steps:\\
	\begin{tabular}{r c l}
		\textbf{Retrieve} & : &  gives the data into the process \\
		\textbf{Generate ID} & : & creates an ID such that we can make the comparison\\
		 & & step at the end through joining the sets\\
		\textbf{Multiply} & : & creates two identical data sets\\
		\textbf{Select Attributes} & : &  troughs away the strata before the clustering step,\\
		& & everything except cluster and id after the clustering \\
		& & and just keeps id and strata for the join step\\
		\textbf{Clustering} & : &  runs the k-means clustering algorithm. The number of \\
		& & Clusters has to be fixed \\
		\textbf{Join} & : & For comparing the clustering result and the strata we \\
		& & join the two filtered data sets by the id \\
	\end{tabular}

	In the clustering block, we can choose between different distance measures and maximal step numbers. We keep the default configuration and choose the mixed euclidean distance measure.
%	We decide to concentrate on almost everywhere basic configurations and chose the squared euclidean distance in the mixed version.
\vspace*{-1em}
	\subsubsection{Gower Distance}\label{gower}, in contrast to the popular distance measures, can also handle the mixed data within the given dataset. The Gower distance measure distinguishes between three types of variables: binary, categorical and numerical.
%	When clustering data with algorithms such as \textit{k-means}, the distance between data points is usually calculated by measures such as the Euclidean or Manhattan norm.
%	These distance measures come with constraints though, since they are only defined for numerical variables. 
	%In the scope of this work, we are facing data points with mixed variable types.
	%, therefore we investigated the \textbf{Gower distance measure} as an appropriate measure to calculate an overall \textit{similarity} between mixed data points. 
%	\\ \textbf{Binary} variables, where two variables of value 0 are \textit{not} considered as match \cite{gower1971general}. In the scope of this work, the only considerable candidate for a binary variable would have been \textit{gender}. Though, during our research process, we concluded that the similarity measure for binary variables was not a suitable choice, as the distance does not match 0 values. Hence, we consider all variables, that are not explicitly numerical, as categorical variables.\\
%	\textbf{Categorical} variables form a set of unordered values and are comparable to ENUMs in programming languages.\\	
%	\textbf{Numerical} variables hold ordered numerical values that support arithmetic operations.
%	\subsubsection{Calculating the distances}
	The distance can be calculated as follows:

	Given two data points $x$ and $y$, each form a tuple of $v$ variables of arbitrary type, the similarity coefficient is given by
	\begin{equation}\label{eq1}
	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
	\end{equation} 
	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at the $k$-th entry in the data points $x$ and $y$. Note, that the definition of the score depends on the
	%Timo nervt
	type of the variable, as defined below. In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, if the two variables are comparable and to 0 if not. %PASS AUF
	For example, variables are not comparable, if values are undefined in the data points or the variable types do not match. Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$. Thus, the similarity coefficient in \hyperref[eq1]{(2)} can be interpreted as the average value of all similarity scores. 
	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation. As pointed above, 0 values are not considered as a match and even further, not considered to be comparable. Hence, the values result as in the table
		\begin{center}
			\begin{tabular}{l | c c c r}
				i & 1 & 1 & 0 & 0 \\
				j & 1 & 0 & 1 & 0 \\
				\hline
				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
			\end{tabular}
		\end{center}	
	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$, and 0, if they differ.

	\textbf{Numerical:} For numerical variables, the similarity score is calculated by
		\begin{equation*}
		s_{xy,k} = 1 - \frac{\vert x_k - y_k \vert}{range(k)}
		\end{equation*}
		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept. This can be a global maximum of acceptable values for variable $k$, or chosen on the basis of the dataset.

	\subsection{Supervised learning}

	\subsubsection{Decision Trees} are a good manner to figure out, which parts of the data set have the most influence on the decision. Therefore labelled data is needed and we have given the strata. We used again RapidMiner for building trees based on different data sets.

	\begin{figure}[!htbp]
		\centering
		\includegraphics[width = 0.9\textwidth]{DecisionTreeRapidModel.PNG}
		\caption{Process for decision trees in RapidMiner}
		\label{fig: RapDec}
	\end{figure}

	RapidMiner does the following steps, illustrated in \Cref{fig: RapDec}:

	\begin{tabular}{r c l}
		\textbf{Retrieve} & : & includes the dataset\\
		\textbf{Select Attributes} & : & makes it possible to or have a look at grouped\\
		&& strata or normal strata\\
		\textbf{Set Role} & : & gives strata the label role, so that the decision tree\\
		&& has those as leafs\\
		\textbf{Multiply} & : & clones the data set \\
		\textbf{Decision Tree} &: & creates the decision tree\\
		\textbf{Apply Model} & : & creates the labelled data set for the \textbf{Performance} step\\
		\textbf{Performance} & : & gives the performance result of the created model\\
	\end{tabular}

	Furthermore we choose information gain as splitting criterion (minimal gain 0.1) and a confidence of 0.25. Other configurations do not show different results.

	\subsubsection{Neural Net}
	% TODO
	% like k-means
	are based on biology. In detail based on the neurons and synapses within the human brain. There a neuron fires to any further connected neurons, if a weighted sum minus a bias value is above a certain threshold.
%	The individual weights and biases are learned by the net.
	Through iteration-wise computing the result and comparing it to the expected outcome, called label and the use of various machine learning techniques the net is able to adjust the weights and biases \cite{neuralNet}.

	\begin{figure}[H]
		\centering
		\includegraphics[width = 0.9\textwidth]{RapidNN.PNG}
	\end{figure}
		\begin{tabular}{r c l}
		\textbf{Retrieve} & : & includes the dataset\\
		\textbf{Replace Missing Values} & : & ensure applicable data\\
		\textbf{Discretize} & : & translates numerical to nominal data \\
		\textbf{Set Role} & : & gives strata the label role, so that the decision tree\\
		&& has those as leafs\\		
		\textbf{Cross Validation} & : & models the neural network \\
	\end{tabular}

### 2018-07-13 dokumentation/DecisionTree.tex
Schlaf ist nur für schwache Menschen...

@@
 \centering
-\begin{subfigure}{0.9\textwidth}
-\includegraphics[width = \linewidth]{Dec200eqPrec.PNG}
-\caption{200 in every strata}
+\includegraphics[width = 0.9\textwidth]{Dec200eqPrec.PNG}
+\caption{Performance for stratified person data with 200 in every strata}
 \label{fig:decvec200}
-\end{subfigure}
-\begin{subfigure}{0.9\textwidth}
-\includegraphics[width= \linewidth]{decvec585pre.PNG}
-\caption{585 in every strata}
-\label{fig:DecVec585}
-\end{subfigure}
-\caption{Performance for stratified person data}
-\label{fig:DecVec}
 \end{figure}
 
-In \Cref{fig:DecVec} the best and worst outcome can be seen for 6 clusters.
+\begin{figure}[H]
+\includegraphics[width= 0.9\textwidth]{decvec585pre.PNG}
+\caption{Performance for stratified person data with 585 in every strata}
+\label{fig:DecVec585}
+\end{figure}
+
+In \Cref{fig:DecVec200} the best and in \Cref{fig:DecVec200} theworst outcome can be seen for 6 clusters.
 For all other configurations the outcome is similar, so just small data sets have an acceptable result and larger datasets have no sufficient results for our question.\\

### 2018-07-13 dokumentation/MiriamClustering.tex
Schlaf ist nur für schwache Menschen...

@@
 
-\begin{figure}[!htbp]
+\begin{figure}[H]
 \centering
@@
 %PCA?
+
+As an improvent of time behaviour and as a further we apply the PCA process on the \textit{original data} without strata and a variance
+threshold of 0.95. We get two Attributes, so we can assume, that the data is strongly corrleated. As expected the computating time decreases a lot for the whole process of clustering.
+
+\begin{figure}[H]
+\vspace*{-1em}
+\centering
+\begin{subfigure}{.3\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidStrata.PNG}
+  \caption{Strata}
+  \label{fig:PCAOrgSt}
+\end{subfigure}%
+\begin{subfigure}{.3\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster.PNG}
+  \caption{Cluster}
+  \label{fig:PCAOrgCl}
+\end{subfigure}
+\begin{subfigure}{.3\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster2Cluster.PNG}
+  \caption{3Cluster}
+  \label{fig:PCAOrgCl3}
+\end{subfigure}
+\caption{Distribution of original data}
+\label{fig:PCAOrgDist}
+\vspace*{-2em}
+\end{figure}
+
+For a clustering of 6 clusters we get a distribution, which can be seen in \Cref{fig:PCAOrgDist}, similar to the one we got from the k-means process with the same comfigurations. 
+For a clustering of 3 clusters we get again results, that are not clear, visible in \Cref{fig:PCAOrgCl3}.
+
+Using different dataset results in comparable unrewarding result.
+\vspace*{-2em}

### 2018-07-13 dokumentation/theoreticalBackground.tex
Schlaf ist nur für schwache Menschen...

@@
 %	\subsubsection{Clustering with RapidMiner}
+
 	As a tool we considered RapidMiner, since it has many modules already efficiently implemented and is easy to adapt.
+
 	\vspace*{-2em}
-	\begin{figure}[!htbp]
+	\begin{figure}[H]
 		\centering
@@
 		\label{fig: kclust}
-		\vspace*{-2em}
+		%\vspace*{-2em}
 	\end{figure}
@@
 	The Process, \Cref{fig: kclust}, contains the following steps:\\
+
 	\begin{tabular}{r c l}
@@
 
+\subsubsection{Principal component analysis}
+
+\textit{PCA} uses an orthogonal transformation to convert a dataset with correlated datas into a dataset without correlations and decreased dimension.
+This is a helpful technique to have a better understanding of the data complexity as well as an improvment of time consumption in further steps.
+
+The first component for the PCA transformation has to satisfy the following:
+
+\vspace*{-1em}
+\begin{align*}
+w_1 &= \text{argmax}_{||w|| = 1} \{\sum_i (t_1)_i^2\}\\
+	&= \text{argmax}_{||w|| = 1} \{\sum_i (x_iw)^2\}
+	\vspace*{-2em}
+\end{align*}
+
+Which is achieved by the corresponding eigenvector.
+
+The next components need to fullfill:
+\begin{align*}
+\hat{X}_k &= X - \sum_{s=1}^{k-1} Xw_sw_s^T\\
+w_k &= \text{argmax}_{||w|| = 1} \{||\hat{X}_kw||^2\}\\
+	&= \text{argmax}_{||w|| = 1} \{\frac{w^T\hat{X}_k^T\hat{X}_kw}{w^Tw}\}
+	\vspace*{-1em}
+\end{align*}
+
+Based on the 'weight matrix' a mapping is possible from the original space to a new space in which the dataset is uncorrelated.
+
+	\vspace*{-1em}
+	\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.9\textwidth]{PCAClustering}
+	\caption{RapidMiner process for PCA and k-means}
+	\label{fig:PCA}
+	\vspace*{-2em}
+	\end{figure}
+
+The Process does the following steps, like in \Cref{fig:PCA} visible:
+
+\begin{tabular}{r c l}
+	\textbf{Retrieve} & : & gives the data in the process.\\
+	\textbf{Generate Id} & : & creates an idea for the join later\\
+	\textbf{Select Attribute} & : & just gives the chosen attributes back\\
+	\textbf{Nominal to Numerical} & : & changes the nominal data to numerical data, so \\
+	&&that we can apply PCA in the next step \\
+	\textbf{PCA} & : & applys the PCA reduction to the data\\
+	&& threshold $0.95$ \\
+	\textbf{Clustering} & : &  clusters the given data \\
+	\textbf{Join} & : & joins the clustering outcome and the strata by the\\
+	&& ID\\
+\end{tabular}
+
 	\subsection{Supervised learning}
@@
 
-	\begin{figure}[!htbp]
+	\vspace*{-2em}
+	\begin{figure}[H]
 		\centering
@@
 		\label{fig: RapDec}
+		\vspace*{-1.5em}
 	\end{figure}
@@
 	Furthermore we choose information gain as splitting criterion (minimal gain 0.1) and a confidence of 0.25. Other configurations do not show different results.
+	\vspace*{-1em}
 
@@
 
+	\vspace*{-2em}
 	\begin{figure}[H]
@@
 		\includegraphics[width = 0.9\textwidth]{RapidNN.PNG}
+		\caption{RapidMiner process for neural nets}
+		\vspace*{-1em}
 	\end{figure}
+
 	\begin{tabular}{r c l}

### 2018-07-15 dokumentation/MiriamClustering.tex
koen hat korrigiert

--- 2018-07-12 dokumentation/MiriamClustering.tex
@@
 
-In \Cref{fig:OrgDist} is the result to see of the first try.
+\Cref{fig:OrgDist} is the result of the first try.
 \Cref{fig:OrgSt} shows the strata distribution as pie chart and \Cref{fig:OrgCl} the resulted cluster distribution.
@@
 
-The next idea is: are 6 cluster too fine? Based on the grouped strata we look at 3 clusters.
+The next idea is: are 6 cluster too specific? Based on the grouped strata we look at 3 clusters.
 
@@
 
-\Cref{fig:OrgDist3Cl} looks promising, so we had a look at the cluster distribution in the 3 grouped stratas, visualized \Cref{fig:Groupdist}.
+\Cref{fig:OrgDist3Cl} looks promising, so we had a look at the cluster distribution in the 3 grouped stratas, visualized in \Cref{fig:Groupdist}.
 It can be seen, that all clusters appear in all grouped strata.
@@
 
-\Cref{fig:1000vect3} shows clearly, that again no correlation can be found. Applying the procedure to the different datasets we generated creates similar results.
+\Cref{fig:1000vect3} shows clearly, that again no correlation can be found. Applying the procedure to the different datasets we generate similar results.
 
 %PCA?
+To save time we apply the PCA process on the \textit{original data} without strata and a variance
+threshold of 0.95. We get two Attributes, so we can assume, that the data is strongly correlated. As expected the computing time decreases a lot for the whole process of clustering.
+
+\begin{figure}[H]
+\vspace*{-1em}
+\centering
+\begin{subfigure}{.3\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidStrata.PNG}
+  \caption{Strata}
+  \label{fig:PCAOrgSt}
+\end{subfigure}%
+\begin{subfigure}{.3\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster.PNG}
+  \caption{Cluster}
+  \label{fig:PCAOrgCl}
+\end{subfigure}
+\begin{subfigure}{.3\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster2Cluster.PNG}
+  \caption{3Cluster}
+  \label{fig:PCAOrgCl3}
+\end{subfigure}
+\caption{Distribution of original data}
+\label{fig:PCAOrgDist}
+\vspace*{-2em}
+\end{figure}
+
+For a clustering of 6 clusters we get a distribution, which can be seen in \Cref{fig:PCAOrgDist}, similar to the one we get from the k-means process with the same configurations. 
+For a clustering of 3 clusters we again get results, that are unclear(\Cref{fig:PCAOrgCl3}).
+
+Using different dataset results in comparable unrewarding results.
+\vspace*{-2em}

### 2018-07-15 dokumentation/conclusion.tex
koen hat korrigiert

@@
 %	\end{itemize}
-	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we though about various aspects having an impact on the movement.
+	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we thought about various aspects having an impact on the movement.
 We state 4 main influencing aspects, explaining the variance throughout the data.
@@
 \Cref{sec: introduction}) is not mentioned.
-Therefore we cannot infer, that it is representative.
-The people, asked to enter their movement, could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
+Therefore we cannot know, if it is representative.
+The people, asked to enter their movement, could had an exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
 \\
-	Also the day itself is not mentioned, so we do not have information, if it was even a working day or a weekend. This influences the behavior drastically.
+	Also the day itself is not mentioned, so we cannot determine, if it was even a working day or a weekendday. This influences the behaviour drastically.
 
@@
 	Obviously, people are individual in their way of life.
-So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply do not like driving.
+So there are people, with enough money to buy for example a car, but refuse to do so in order to reduce CO$_2$ emission, or simplydislike driving.
 On the other end of the spectrum, some people, who have little money, still own and drive a car daily in order to go to work, since they cannot afford an other apartment.\\
@@
 		\item[2. Group]
-		People, who are rich just by birth, which do not have the necessity to work or move at all. They could possibly stay home and do not have to leave at all.
+		People, who are rich by birth, which do not have the necessity to work or move at all. They could possibly stay home and do not have to leave at all.
 	\end{enumerate} 
 	Those two groups are completely opposite, but still belong to the same strata.
-Now also considering strata one, we can find the exact same movement behavior in there as well.
+Now also considering strata 1, we can find the exact same movement behaviour in there as well.
 Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work.
@@
 	So we can see that based on the information we have, it is very unlikely to have a clear correlation between the given data and the strata.
-However, given more information about the previously mentioned aspects and more entries in general, there could be some kind of correlation to be found.\\
-	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
+However, given more information about the previously mentioned aspects and more entries in general, a kind of correlation could be found.\\
+	Future research could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
 Also one could use different network visualization tools in order to infer patterns.

### 2018-07-15 dokumentation/introduction.tex
koen hat korrigiert

@@
 	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the collected data?
-In particular is it possible to determine the wealth of a person, only given movements of a single day?
+In particular is it possible to determine the wealth of a person, only given movements of one single day?
 	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
@@
 	First, by clustering using the k-means algorithm, considering different distance measures, with optionally including the principal component analysis (PCA).\\
-	And second, a decision tree and neural net were trained and tested.
+	And second, a decision tree and neural net by training and testing.

### 2018-07-15 dokumentation/observation.tex
koen hat korrigiert

@@
 Also, we detected that reducing the numbers of clusters leads to better results, but decreases the significance of the statement, that could be made.\\
-	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
+	Using a different neural net architecture will most likely not chance the accuracy value drastically.
+As stated, more complex nets need more training data, which is restricted for reasons mentioned above.\\
 	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
-Therefore we see, the more data we have, the higher is the variance within the stratas themselves.
-We cannot draw some kind of lines to distinguish between different stratas.
+Therefore we see, the more data we have, the higher the variance within the stratas.
+We cannot draw lines to distinguish between different stratas.
 Data points, which were outliers in the smaller sets, are now not considered to be outliers, because many others have the same characteristics.

### 2018-07-15 dokumentation/preprocessing.tex
Schlaf ist nur für schwache Menschen...
koen hat korrigiert

@@
 	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
-	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
+	We observe that even though we have 124979 individual lines defining a movement, there is 1 line defining a \texttt{NotANumber}-exception and therefore gets excluded from further usage.	\\
 	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets used by our clustering and neural net approaches.\\
@@
 	There is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
-	Furthermore we have to make sure that two different data points, which belong to the very same person, are assigned to the same cluster.
+	Furthermore, we have to make sure that two different data points, which belong to the very same person, are assigned to the same cluster.
 To do so, we compute the value \texttt{ID}, which identifies each person and can be used to combine movements that are considered to be from the same person.

### 2018-07-15 dokumentation/submission.tex
koen hat korrigiert

@@
 		We analyze movements in the urban environment of the Columbian city Medellín.
-Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
+Each movement is given as spatiotemporal pattern with additional information about the reason of movment, means of transportation and the corresponding person concern the socio-economic status (strata), the age and gender.
 		% Marías sentence
-		The objective of the paper is twofold: try to find groups of movement patterns and see, whether they correspond to socio-economic status, and then, since we have the economic status of each person in the data set, we will apply supervised learning to classify patterns into socio-economic status.
+		The objective of the paper is twofold: first we try to find groups of movement patterns and see, whether they correspond to socio-economic status, with this information, since we have the economic status of each person in the data set, we will apply supervised learning to classify patterns into socio-economic status.
 %		Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
 Due to bad results we introduce our preprocessing steps and switch to supervised learning.
-		Decision trees and neural networks did neither match our performance expectations, which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
+		Neither decision trees nor neural networks match our performance expectations, which leads to the conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
 

### 2018-07-15 dokumentation/theoreticalBackground.tex
koen hat korrigiert

@@
 
-	is a \textit{centroid based technique}, which means that each cluster is represented by a data point, that also is the centre of the cluster.
+	is a \textit{centroid based technique}, which means that each cluster is represented by a data point, which at the same time is the centre of the cluster.
 A distance measure is then used to assign every remaining data object from the data set to the best fitting cluster.
@@
 		\textbf{Retrieve} & : &  gives the data into the process \\
-		\textbf{Generate ID} & : & creates an ID such that we can make the comparison\\
+		\textbf{Generate ID} & : & creates an ID so we can make the comparison\\
 		 & & step at the end through joining the sets\\
 		\textbf{Multiply} & : & creates two identical data sets\\
-		\textbf{Select Attributes} & : &  troughs away the strata before the clustering step,\\
-		& & everything except cluster and id after the clustering \\
+		\textbf{Select Attributes} & : &  throws away the strata before the clustering step,\\
+		& & everything except cluster and ID after the clustering \\
 		& & and just keeps id and strata for the join step\\
@@
 		\textbf{Join} & : & For comparing the clustering result and the strata we \\
-		& & join the two filtered data sets by the id \\
+		& & join the two filtered data sets by the ID \\
 	\end{tabular}
@@
 	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
-As pointed above, 0 values are not considered as a match and even further, not considered to be comparable.
+As pointed out above, 0 values are not considered as a match such not even considered to be comparable.
 Hence, the values result as in the table
@@
 \textit{PCA} uses an orthogonal transformation to convert a dataset with correlated datas into a dataset without correlations and decreased dimension.
-This is a helpful technique to have a better understanding of the data complexity as well as an improvment of time consumption in further steps.
+This is a helpful technique to get a better understanding of the data complexity as well as decreasing the amount of time consumption in further steps.
 
@@
 
-The Process does the following steps, like in \Cref{fig:PCA} visible:
+The Process does the following steps, visible in \Cref{fig:PCA}:
 
 \begin{tabular}{r c l}
-	\textbf{Retrieve} & : & gives the data in the process.\\
+	\textbf{Retrieve} & : & imports the data in the process.\\
 	\textbf{Generate Id} & : & creates an idea for the join later\\
-	\textbf{Select Attribute} & : & just gives the chosen attributes back\\
+	\textbf{Select Attribute} & : & gives the chosen attributes back\\
 	\textbf{Nominal to Numerical} & : & changes the nominal data to numerical data, so \\
 	&&that we can apply PCA in the next step \\
-	\textbf{PCA} & : & applys the PCA reduction to the data\\
+	\textbf{PCA} & : & applies the PCA reduction to the data\\
 	&& threshold $0.95$ \\
@@
 	\subsubsection{Decision Trees} are a good manner to figure out, which parts of the data set have the most influence on the decision.
-Therefore labelled data is needed and we have given the strata.
-We used again RapidMiner for building trees based on different data sets.
+Therefore, labelled data is needed and we have given the strata.
+We again used RapidMiner for building trees based on different data sets.
 
@@
 		\textbf{Retrieve} & : & includes the dataset\\
-		\textbf{Select Attributes} & : & makes it possible to or have a look at grouped\\
+		\textbf{Select Attributes} & : & makes it possible to have a look at grouped\\
 		&& strata or normal strata\\

### 2018-07-16 dokumentation/MiriamClustering.tex
Miriam hat keine Lust mehr

@@
 %oder doch eilaina
-In the first step we try to cluster the \textit{Original Data} in \textit{6 Cluster}. Therefore we retrieve the original data set in RapidMiner and choose k as 6.
+In the first step we try to cluster the \textit{Original Data} in \textit{6 cluster}. Therefore we retrieve the original data set in RapidMiner and choose $k$ as 6.
 
@@
 
-\Cref{fig:OrgDist} is the result of the first try.
-\Cref{fig:OrgSt} shows the strata distribution as pie chart and \Cref{fig:OrgCl} the resulted cluster distribution.
+\Cref{fig:OrgDist} is the result of the first try represented as pie charts.
+\Cref{fig:OrgSt} shows the strata distribution and \Cref{fig:OrgCl} the resulted cluster distribution.
 It can already been seen, that the distributions are not similar.
@@
 \includegraphics[width=0.5\textwidth]{ClusterOrigRapidDistribution2Cluster.PNG}
-\caption{Distribution of the clusters in between the grouped strata}
+\caption{Distribution of the clusters within the grouped strata}
 \label{fig:Groupdist}
@@
 \end{figure}
+\vspace*{-3em}
 
@@
 \end{figure}
+\vspace*{-2em}
 
 Different dataset sizes and equal distribution of the stratas do not effect the result.
-In  \Cref{fig:2038_Clust} two clustering results can be seen.
+In  \Cref{fig:2038_Clust} the clustering results for the maximal sized dataset can be seen.
+Once aggregated and once normal.
 Again we could not observe a significant correlation.
 
-After those not convincing results we applied the process to the \textit{stratified person data}.
+After those not convincing results, we apply the process to the \textit{stratified person data}.
 
-The results for the whole data set without equalization are shown in \Cref{fig:2038_Clust}.
+The results for the whole data set without equalization are shown in \Cref{fig:VecDist}.
 We change the number of steps to 1000 for comparison and the result, illustrated in \Cref{fig:1000vect}, lets us assume, that 3 clusters would fit better.
@@
 \centering
-\includegraphics[width=0.69\textwidth]{vectorClustering31000.PNG}
+\includegraphics[width=0.5\textwidth]{vectorClustering31000.PNG}
 \caption{1000 steps clustering in 3 clusters on maximal stratified person dataset}
@@
 
-\Cref{fig:1000vect3} shows clearly, that again no correlation can be found. Applying the procedure to the different datasets we generate similar results.
+\Cref{fig:1000vect3} shows clearly, that again no correlation can be found. Applying the procedure to the different datasets, we generate similar results.
 
 %PCA?
-To save time we apply the PCA process on the \textit{original data} without strata and a variance
-threshold of 0.95. We get two Attributes, so we can assume, that the data is strongly correlated. As expected the computing time decreases a lot for the whole process of clustering.
+To save time, we apply the PCA process on the \textit{original data} without strata and a variance
+threshold of 0.95. We get two attributes, so we can assume, that the data is strongly correlated. As expected the computing time decreases a lot for the whole process of clustering.
 
@@
   \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster.PNG}
-  \caption{Cluster}
+  \caption{6 cluster}
   \label{fig:PCAOrgCl}
@@
   \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster2Cluster.PNG}
-  \caption{3Cluster}
+  \caption{3 cluster}
   \label{fig:PCAOrgCl3}
 \end{subfigure}
-\caption{Distribution of original data}
+\caption{Distributions of the original data. Clustering with the PCA process}
 \label{fig:PCAOrgDist}
@@
 
-For a clustering of 6 clusters we get a distribution, which can be seen in \Cref{fig:PCAOrgDist}, similar to the one we get from the k-means process with the same configurations. 
-For a clustering of 3 clusters we again get results, that are unclear(\Cref{fig:PCAOrgCl3}).
+For the clustering with the preprocessing step PCA, we get distributions, which can be seen in \Cref{fig:PCAOrgDist}, similar to the distributions we get from the k-means process with the same configurations.
+For a clustering of 3 clusters we again get results, that are not meaningful (\Cref{fig:PCAOrgCl3}).
 

### 2018-07-16 dokumentation/conclusion.tex
Miriam hat keine Lust mehr

@@
 %	\end{itemize}
-	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we thought about various aspects having an impact on the movement.
+	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we come uo with various aspects having an impact on the data.
 We state 4 main influencing aspects, explaining the variance throughout the data.
@@
 \Cref{sec: introduction}) is not mentioned.
-Therefore we cannot know, if it is representative.
-The people, asked to enter their movement, could had an exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day.
+Therefore we do not know, if it is representative.
+The people, asked to enter their movement, could had an exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from an usual day.
 \\
-	Also the day itself is not mentioned, so we cannot determine, if it was even a working day or a weekendday. This influences the behaviour drastically.
+	Also the day itself is not mentioned, so we cannot determine, if it was even a working day or a weekend day. This influences the behaviour drastically.
 
@@
 	Obviously, people are individual in their way of life.
-So there are people, with enough money to buy for example a car, but refuse to do so in order to reduce CO$_2$ emission, or simplydislike driving.
+So there are people, with enough money to buy for example a car, but refuse to do so in order to reduce CO$_2$ emission, or simply dislike driving.
 On the other end of the spectrum, some people, who have little money, still own and drive a car daily in order to go to work, since they cannot afford an other apartment.\\
@@
 		\setlength{\itemindent}{1cm}
-		\item[1. Group]
-		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
-		\item[2. Group]
+		\item[1. Group:]
+		Hard working people, laboring 60+ hours per week to earn money. One can imagine that they have to move quite a lot since they are always busy.
+		\item[2. Group:]
 		People, who are rich by birth, which do not have the necessity to work or move at all. They could possibly stay home and do not have to leave at all.
@@
 	So we can see that based on the information we have, it is very unlikely to have a clear correlation between the given data and the strata.
-However, given more information about the previously mentioned aspects and more entries in general, a kind of correlation could be found.\\
+However, given more information about the previously mentioned aspects and more entries a correlation can not be ruled out in general.\\
 	Future research could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering.
 Also one could use different network visualization tools in order to infer patterns.
-An example would be to take the origin (destination) sectors as nodes, an edge, if there is a movement between, directed or undirected, and a different thickness or color gradient, based on the number of this edge being taken.\\
+An example would be to take the origin (destination) sectors as nodes, an edge, if there is a movement between, directed or undirected, and a different thickness or color gradient, based on the number of times this edge being taken.\\
 
-	Overall we can conclude that with the data we got, we are not able to find a correlation.
+	Overall we can conclude that with the data we get, we are not able to find a correlation.

### 2018-07-16 dokumentation/introduction.tex
Miriam hat keine Lust mehr

@@
 	% could add more in depth 
-	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the collected data?
-In particular is it possible to determine the wealth of a person, only given movements of one single day?
+	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the collected data.
+In particular, is it possible to determine the wealth of a person, only given movements of one single day?
 	For this we considered the dataset stated in \cite{rich_do_not_rise_early}.
@@
 The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs.
-Those costs are defined by Colombians laws, which classifies households to regulate the access to public utility services, having as a result six socio-economic strata \cite{rich_do_not_rise_early}.\\
-	Our goal is to ascertain, if there is a correlation between the movements and the strata, in order to be able to predict the strata based on the movements. This is done in two steps:\\
+Those costs are defined by Colombians laws, which classifies households to regulate the access to public utility services, having as a result six socio-economic stratas \cite{rich_do_not_rise_early}.\\
+	Our goal is to ascertain, if there is a correlation between the movements and the strata, in order to be able to reproduce or predict the strata based on the movements. This is done in two steps:\\
 	First, by clustering using the k-means algorithm, considering different distance measures, with optionally including the principal component analysis (PCA).\\

### 2018-07-16 dokumentation/neuralnet.tex
Miriam hat keine Lust mehr

@@
 	In this section we want to further improve the accuracy of predicting the strata of one person by using neural networks.
-Like in the previous section, where we used decision trees, we need ground truth, in this case the strata, to be able to train the neural networks.
-For all the neural net computations we considered person vector data sets of different sizes (c.f.
+Like in the previous section, where we used decision trees, we need ground truth to be able to train the neural networks.
+For all the neural net computations we consider person vector data sets of different sizes (c.f.
 \Cref{sec: proprocessing}).\\
@@
 the data is split into 10 subsets, where in each iteration exactly one data set is used as test set and the other 9 as training set.
-The average value of the accuracy values leads to the total accuracy of the neural net.
+The average value of those accuracy values leads to the total accuracy of the neural net.
+
+	In the following we consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
+	We call the aggregated strata sets $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$ denoting the number of neurons.
+This builds a superset of the original stratas and since the stratas themselves are logically connected, this task should be easier to fulfill.\\
 		\begin{figure}[H]
@@
 
-	In the following we consider 3 neural nets $\mathcal{N}_1,\mathcal{N}_2$ and $\mathcal{N}_3$, all having 4 hidden layers, 50 epochs and 10 iterations.	
-	As an example of other strata aggregation we combine the stratas 1--2, 3--4 and 5--6 together and call them $\mathcal{N}_i^\star$, for $i \in \{5,10,20\}$.
-This builds a superset of the original stratas and since the stratas themselves are logically connected this task should be easier to fulfill.\\
-
-	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata which are provided by the \texttt{testDataGenerator} from \Cref{sec: proprocessing}.
-For every neural net and every set size we perform 5 independent runs and calculate the average over those accuracy values in order to have a sophisticated, comparable statement.
+	For each neural net we are using equally distributed data sets with 100, 200 and the maximal amount of 595 individuals per strata.
+For every neural net and every set size, we perform 5 independent runs and calculate the average over those accuracy values in order to have a sophisticated, comparable statement.
 	\setlength\tabcolsep{.2cm}
@@
 		\end{tabular}
-		\caption{The accuracy values of the neural nets. (See excel-spreadsheet)}
+		\caption{The accuracy values of the neural nets}
 		\label{tab: nn-accuracy}
@@
 	%TODO: recompute
-	The size of larger nets in terms of neurons is counter-productive, since, if we take 50 neurons per layer, we have $14 \cdot 50^4 \cdot 6 \approxeq 525.000.000$ synapses for which the input dataset would be too small to perform sufficient training.\\
+	The size of larger nets, in terms of neurons, is counter-productive, since, if we take 50 neurons per layer, we have $14 \cdot 50^4 \cdot 6 < 849 \cdot 50^4 \cdot 6 \approxeq 31.837.500.000$ synapses for which the input dataset would be too small to perform sufficient training.\\

### 2018-07-16 dokumentation/observation.tex
Miriam hat keine Lust mehr

@@
 	%	The greater the population, the more the borders between the stratas blur. 
-	As an overall result we get, that we can not determine the strata based on the information we have.
-	Using various datasets (c.f.
-\Cref{sec: proprocessing}), we witnessed, that having equally distributed datasets leads to an overall higher accuracy.
+	As an overall result we observe, that we can not determine the strata based on the information we have.
+	Using various datasets, we witness, that having equally distributed datasets leads to an overall higher accuracy.
 This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f.
 \Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
-	During clustering we observed, that using a different distance measure formula would not lead to a huge difference.
-Also, we detected that reducing the numbers of clusters leads to better results, but decreases the significance of the statement, that could be made.\\
+	During clustering we observe, that using a different distance measure formula would not lead to a huge difference.
+Also, we detect that reducing the number of clusters leads to better results, but decreases the significance of the statement, that could be made.\\
 	Using a different neural net architecture will most likely not chance the accuracy value drastically.
-As stated, more complex nets need more training data, which is restricted for reasons mentioned above.\\
+As stated, more complex nets need more training data, which is restricted for the reasons mentioned above.\\
 	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions.
-Therefore we see, the more data we have, the higher the variance within the stratas.
+Therefore we see: the more data we have, the higher the variance within the stratas is.
 We cannot draw lines to distinguish between different stratas.

### 2018-07-16 dokumentation/preprocessing.tex
Miriam hat keine Lust mehr

@@
 	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
-	We observe that even though we have 124979 individual lines defining a movement, there is 1 line defining a \texttt{NotANumber}-exception and therefore gets excluded from further usage.	\\
+	We observe that even though we have 124979 individual lines defining a movement, there is one line encounter a \texttt{NotANumber}-exception and therefore gets excluded from further usage.	\\
 	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets used by our clustering and neural net approaches.\\
@@
 	Following, we introduce vectors representing single persons.
-Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset where each data point coincides with one person.\\
+Since strata 6 is the smallest strata with 595 persons, it limits the size of an equally distributed dataset, where each data point coincides with one person.\\
 
-	\textbf{Stratified Person Data}\label{subsubsec: person vector data}\\
-	As stated before, instead of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
-This class stores the ID, the parameters defining a person (c.f.
-\Cref{sec: proprocessing}), and all movements from that person.\\
+	Moreover, we define aggregated strata as the combination of strata 1 and 2 to "strata\_1", strata 3 and 4 to "strata\_2" and strata 5 and 6 to "strata\_3".
+
+	\textbf{Stratified person data}\label{subsubsec: person vector data}\\
+	As stated before, additionally of simple IDs for every person we expand the parsing by using a data encapsulating in a class called \texttt{Person}.
+This class stores the ID, the parameters defining a person, and all movements from that person.\\
 	Then we are able to compute the following vector, with 850 entries, for further usage, that combines all movements of the person:

### 2018-07-16 dokumentation/submission.tex
Miriam hat keine Lust mehr

--- 2018-07-12 dokumentation/submission.tex
@@
 		We analyze movements in the urban environment of the Columbian city Medellín.
-Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender.
+Each movement is given as spatiotemporal pattern with additional information about the reason of movement, means of transportation and the corresponding person concerning the socio-economic status (strata), the age and gender.
 		% Marías sentence
-		The objective of the paper is twofold: try to find groups of movement patterns and see, whether they correspond to socio-economic status, and then, since we have the economic status of each person in the data set, we will apply supervised learning to classify patterns into socio-economic status.
+		The objective of the paper is twofold: first we try to find groups of movement patterns and see, whether they correspond to the socio-economic status, with this information, since we have the economic status of each person in the data set, we will apply supervised learning to classify patterns into socio-economic status.
 %		Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters.
 Due to bad results we introduce our preprocessing steps and switch to supervised learning.
-		Decision trees and neural networks did neither match our performance expectations, which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.
+		Neither decision trees nor neural networks match our performance expectations, which leads to the conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately, if even possible.
 
@@
 
-	\subsection{Natural Clusters} \label{subsec: classification}
+	\subsection{Natural clusters} \label{subsec: classification}
 

### 2018-07-16 dokumentation/theoreticalBackground.tex
Miriam hat keine Lust mehr

@@
 
-	In the scope of this work, we decided to use the well-known partitioning methods k-means.
-In general, given $n$ data objects partitioning methods distribute the data objects into $k$ clusters with $k\le n$, using a distance measure to evaluate the respective similarity.
+	In the scope of this work, we decided to use the well-known partitioning method k-means.
+In general, given $n$ data objects, partitioning methods distribute the data objects into $k\in\mathbb{N}$ clusters with $k\le n$, using a distance measure to evaluate the respective similarity.
 %	Those methods form exclusive clusters by ensuring that each cluster contains at least one object.
@@
 
-	\subsubsection{k-Means}
+	\subsubsection{k-means}
 
-	is a \textit{centroid based technique}, which means that each cluster is represented by a data point, which at the same time is the centre of the cluster.
+	is a \textit{centroid based technique}, meaning that each cluster is represented by a data point, which at the same time is the centre of the cluster.
 A distance measure is then used to assign every remaining data object from the data set to the best fitting cluster.
@@
 
-	The data objects within a dataset are considered to reside in a euclidean space.
+	The data objects within a dataset are considered to reside in an euclidean space.
 Thus, the euclidean distance is used to calculate a score for the similarity of two data points.
-When using k-Means, the quality of a cluster $C_i$ can be evaluated by computing the sum of squared errors between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
+When using k-means, the quality of a cluster $C_i$, $i\in [k]$, can be evaluated by computing the sum of squared distances between all data points $p$ in the object space and the centroids $c_i \in C_i$ of every cluster.
 This method is known as \textit{within-cluster variation} \cite{data_mining} and defined as follows:
@@
 %	objects in the data set as initial representatives for the cluster centres.
-	After that, each remaining data object is assigned to the best fitting cluster based on a similarity score between the data point and every cluster centre.
-The overall goal of k-Means clustering is to iteratively optimize the within-cluster variation.
 	In order to achieve this, each cluster centroid is redefined as the mean of all objects within that cluster.
 By considering the updated centroids, every data point is reassigned to the now best fitting cluster.
-This iterative process will continue until no better clustering can be found.\\
+This iterative process will continue until no better clustering can be found or a maximal chosen step size is reached.\\
 %	\subsubsection{Clustering with RapidMiner}
 
-	As a tool we considered RapidMiner, since it has many modules already efficiently implemented and is easy to adapt.
+	As a tool we consider RapidMiner, since it has many modules already efficiently implemented and is easy to adapt.
 
@@
 
-	The Process, \Cref{fig: kclust}, contains the following steps:\\
+	The Process (c.f. \Cref{fig: kclust}) contains the following steps:\\
 
 	\begin{tabular}{r c l}
-		\textbf{Retrieve} & : &  gives the data into the process \\
+		\textbf{Retrieve} & : &  imports the data into the process \\
 		\textbf{Generate ID} & : & creates an ID so we can make the comparison\\
 		 & & step at the end through joining the sets\\
+		 && based on the ID\\
 		\textbf{Multiply} & : & creates two identical data sets\\
-		\textbf{Select Attributes} & : &  throws away the strata before the clustering step,\\
+		\textbf{Select Attributes} & : &  removes the strata before the clustering step,\\
 		& & everything except cluster and ID after the clustering \\
-		& & and just keeps id and strata for the join step\\
+		& & and just keeps ID and strata for the join step\\
 		\textbf{Clustering} & : &  runs the k-means clustering algorithm. The number of \\
-		& & Clusters has to be fixed \\
-		\textbf{Join} & : & For comparing the clustering result and the strata we \\
-		& & join the two filtered data sets by the ID \\
+		& & clusters has to be fixed \\
+		\textbf{Join} & : & For comparing the clustering result and the strata, we \\
+		& & join the two filtered data sets on the ID \\
 	\end{tabular}
@@
 \vspace*{-1em}
-	\subsubsection{Gower Distance}\label{gower}, in contrast to the popular distance measures, can also handle the mixed data within the given dataset.
+	\subsubsection{Gower distance{\normalfont,}}\label{gower} in contrast to the popular distance measures, can also handle the mixed data within the given dataset.
 The Gower distance measure distinguishes between three types of variables: binary, categorical and numerical.
@@
 
-	Given two data points $x$ and $y$, each form a tuple of $v$ variables of arbitrary type, the similarity coefficient is given by
+	Given two data points $x$ and $y$, each form a tuple of $v$ variables of arbitrary type, the similarity coefficient is given by:
 	\begin{equation}\label{eq1}
@@
 Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$.
-Thus, the similarity coefficient in \hyperref[eq1]{(2)} can be interpreted as the average value of all similarity scores.
+Thus, the similarity coefficient in equation \hyperref[eq1]{(2)} can be interpreted as the average value of all similarity scores.
 	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
-	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation.
-As pointed out above, 0 values are not considered as a match such not even considered to be comparable.
+	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical conjunction operation.
+As pointed out above, 0 values are considered to be not a match, thus not even considered to be comparable.
 Hence, the values result as in the table
@@
 		\end{equation*}
-		where $range(k)$ is the total range of values, that the numerical variable at index $k$ can accept.
-This can be a global maximum of acceptable values for variable $k$, or chosen on the basis of the dataset.
+		where $range(k)$ is the total range of values, that the numerical attribute at index $k$ can accept.
+This can be a global maximum of acceptable values for attribute $k$, or chosen on the basis of the dataset.
 
@@
 
-\textit{PCA} uses an orthogonal transformation to convert a dataset with correlated datas into a dataset without correlations and decreased dimension.
-This is a helpful technique to get a better understanding of the data complexity as well as decreasing the amount of time consumption in further steps.
+(PCA) uses an orthogonal transformation to convert a dataset, with correlated data, into a dataset without correlations and decreased dimension.
+This is a helpful technique to get a better understanding of the data complexity, as well as decreasing the time consumption in further steps.
 
-The first component for the PCA transformation has to satisfy the following:
+The first component for the PCA transformation, the weight matrix, has to satisfy the following:
 
@@
 \begin{align*}
-w_1 &= \text{argmax}_{||w|| = 1} \{\sum_i (t_1)_i^2\}\\
-	&= \text{argmax}_{||w|| = 1} \{\sum_i (x_iw)^2\}
+w_1 &= \text{argmax}_{\|w\| = 1} \{\sum_i (t_1)_i^2\}\\
+	&= \text{argmax}_{\|w\| = 1} \{\sum_i (x_iw)^2\}
 	\vspace*{-2em}
@@
 
-Which is achieved by the corresponding eigenvector.
+which is achieved by the corresponding eigenvector.
 
@@
 \hat{X}_k &= X - \sum_{s=1}^{k-1} Xw_sw_s^T\\
-w_k &= \text{argmax}_{||w|| = 1} \{||\hat{X}_kw||^2\}\\
-	&= \text{argmax}_{||w|| = 1} \{\frac{w^T\hat{X}_k^T\hat{X}_kw}{w^Tw}\}
+w_k &= \text{argmax}_{\|w\| = 1} \{\|\hat{X}_kw\|^2\}\\
+	&= \text{argmax}_{\|w\| = 1} \{\frac{w^T\hat{X}_k^T\hat{X}_kw}{w^Tw}\}
 	\vspace*{-1em}
@@
 
-Based on the 'weight matrix' a mapping is possible from the original space to a new space in which the dataset is uncorrelated.
+Based on the weight matrix a mapping is possible from the original space to a new space in which the dataset is uncorrelated.
 
@@
 	\textbf{Retrieve} & : & imports the data in the process.\\
-	\textbf{Generate Id} & : & creates an idea for the join later\\
+	\textbf{Generate ID} & : & creates an idea for the join later\\
 	\textbf{Select Attribute} & : & gives the chosen attributes back\\
@@
 	\subsubsection{Decision Trees} are a good manner to figure out, which parts of the data set have the most influence on the decision.
-Therefore, labelled data is needed and we have given the strata.
-We again used RapidMiner for building trees based on different data sets.
+Therefore, labelled data is needed, which we have given with the strata.
+We again used RapidMiner for building trees, based on different data sets.
 
@@
 
-	Furthermore we choose information gain as splitting criterion (minimal gain 0.1) and a confidence of 0.25. Other configurations do not show different results.
+	We tried several configurations regarding the splitting criterion and  confidence threshold, but the results were similar.
 	\vspace*{-1em}
@@
 	are based on biology.
-In detail based on the neurons and synapses within the human brain.
+In detail, based on the neurons and synapses within the human brain.
 There a neuron fires to any further connected neurons, if a weighted sum minus a bias value is above a certain threshold.
 %	The individual weights and biases are learned by the net.
-	Through iteration-wise computing the result and comparing it to the expected outcome, called label and the use of various machine learning techniques the net is able to adjust the weights and biases \cite{neuralNet}.
+	Through iteration-wise computing the result and comparing it to the expected outcome, called label, and the use of various machine learning techniques the net is able to adjust the weights and biases \cite{neuralNet}.
 

### 2018-07-17 statement.tex
Statement - Miriam stinkt :D
Hab meins grob geschrieben. Werde da wsk noch was ändern

\documentclass{article}

\begin{document}

\section{Miriam}
My part of the practical part was mostly the RapidMiner part. I did the clustering and decision tree  mining with help of this tool. I evaluated the outcome and adapted the system. Also I gave input about the requirements of the testdata generator, especially the person vector. In the second step, together with Timo, we built the presentation. We merged the different parts and prepared the presentation. In the last step I firstly wrote down my own results and together with Timo joined all the different parts, corrected with Timo the resulting paper and we made the fine tuning of it at the end.

\section{Martin}
At beginning of the practicum I was mainly focused on experimenting with the Rapidminer toolkit to find ways to read, process and evaluate the raw data, especially using k-means without PCA (neglecting the given stratification) and using neural networks for classification using the given ground truth. I also was looking for clusters inside the whole dataset by generating different charts that visualize a subset of attributes of all data points with Rapidminer. This was part of the final presentation but was later cut out of the final paper. The testing of performance of the neural networks (which was labor intensive and error prone) was done with Timo to enable double checks, with whom I also wrote the corresponding part in the final paper. Additionally I wrote the abstract.

\section{Walter}

\section{Moritz}

\section{Timo}	
My main focus at the start of the practicum was to create a sophisticated generator for different testsets, with the sizes and properties like different sizes and distribution, for which I wrote the \texttt{testDataGenerator.py}. All this includes an analysis of the data with the whole team. Further, I implemented the vector computation of mapping movements to persons and computing the vector format proposed by Miriam. All those different options are choose-able through commandline flags. For the whole script I wrote comments and documentation in order to provide a reusable base.\\
Later through the project my focus was to aid during various tasks. Therefore, Martin and me build different neural net architectures and tested them under various conditions and wrote the results down for the final paper.\\
Also Miriam and I tried several different clusterings, which she build in RapidMiner. We also prepared and practiced the presentation, which included the tasks of thinking about topics, which were not included of the technical part of our team. Like, a precise but short introduction or reasons for the problem, that we cannot reproduce or find clusters, we faced.\\
After the presentation we merged the different written parts of our team members and made additions or replacements if necessary. After the feedback from María, also Miriam and I, restructured and basically (re-)wrote certain aspects. This 
was finished by proofreading and fine tuning through the two of us.
\end{document}

### 2018-07-17 statement.tex
jetzt steht da nicht mehr Mara..... Ach und miriam geht jetzt pennen

@@
 \documentclass{article}
+\usepackage[utf8]{inputenc}
 
@@
 \section{Miriam}
-My part of the practical part was mostly the RapidMiner part.
+My part of the practical part, the beginning of the practicum, was mostly the RapidMiner.
 I did the clustering and decision tree  mining with help of this tool.
-I evaluated the outcome and adapted the system.
+I evaluated the outcome and adapted the system, such that I tried different configurations and all sets generated.
 Also I gave input about the requirements of the testdata generator, especially the person vector.

### 2018-07-17 statement.tex
Timo stinkt viel mehr!!!!

@@
 \section{Timo}	
+%#Streber
+%#PizzaFresser
+%#WohntMitFranzosenZusammen
+%#Sonnenbrand
 My main focus at the start of the practicum was to create a sophisticated generator for different testsets, with the sizes and properties like different sizes and distribution, for which I wrote the \texttt{testDataGenerator.py}.

### 2018-07-17 statement.tex
ihr riecht beide irgendwie seltsam

@@
 
-\section{Walter}
+\section{Walter \& Moritz}
 
-\section{Moritz}
+The two of us grouped up for our contribution to the practicum work.
+Our task was to research the theoretical backgrounds on the topics "clustering algorithms" and "distance measures".
+We implemented the gower distance, provided presentation slides and did some coding to test approaches with the generated data from other group members.
+The latter, among other things, did not result as part of the final submission and we shortened some of our contributions to match the submission requirements.
+The tasks were equally distributed among us.
 
